### 2015-12-22 ms.tex
Initial import and setup

\documentclass[a4paper,11pt]{article}
\usepackage[osf]{mathpazo}
\usepackage{ms}
\usepackage[]{natbib}
\raggedright

\newcommand{\smurl}[1]{{\footnotesize\url{#1}}}
\usepackage{graphicx}

\title{Toward dynamic models of combat or lack thereof in animal mating}
\author{
* John Wilshire$^1$, Will Cornwell$^1$ , Daniel Falster$^2$, Michael Kasumovic$^1$, \\
Daniel Noble$^1$,$^1$ Loic Thibaut$^1$}
\affiliation{
*final list and order undecided\\
$^1$ University of NSW\\
$^2$ Macquarie University\\
}
\date{}

\bibliographystyle{mee}

\usepackage[title,titletoc,toc]{appendix}

\mstype{Research Article}
\runninghead{A new framework for fighting}
\keywords{}

\begin{document}
\mstitlepage
\noindent
% \doublespacing
\linenumbers

\section{Summary}
The diversity of animal mating systems is astounding. In some of these
systems, very costly combat behaviour -- among males, among females, or
both -- is a feature of the mating process.  In other systems, resources
are divided among individuals in an entirely pacific process.  Can we
understand why? Animal combat strategies likely emerge from trade-offs
in investment in growth, mate seeking, and information gathering.
Willingness to engage in combat is a trait that evolves based on the
fitness landscape, which itself changes depending on both the
environment and the strategies of other individuals.  Using recently
developed methods for modelling dynamic fitness landscapes, we examine:
(1) why combat behaviours arise, (2) under what conditions combat
behaviours are evolutionarily stable, and (3) when different combat
strategies co-exist.  We hypothesize that the reliability and
"public-ness" of information is an important feature driving combat or
lack thereof in many animal systems.

\subsection{Introduction}

Evolution of animal personalities: \citep{Wolf-2007}

Individual-based models of natural selection :\citep{MGonigle-2012}

\subsection{Methods}

\bibliography{refs}

\end{document}

### 2015-12-22 ms.tex
Add references, subsections
figure explanations and Triats

@@
 
-\subsection{Introduction}
+\section{Introduction}
 
@@
 
-\subsection{Methods}
+\section{Methods}
+
+\subsection{Population dynamics}
+
+\subsection{Encounters}
+
+\subsection{Evolutionary dynamics}
+
+\subsection{Numerical analysis}
+
+\section{Figures}
+Figure 1. Event Algorithm
+Population state generates events, events are either:
+\begin{enumerate}
+    \item A new individual x has matured and entered the contest (based off of a trait(?))
+    \item Discovery event wherin both a nest RR (repoductive resource) and an individual x are drawn from a pool, that nest is either:
+\begin{enumerate}
+    \item un-occupied: in which x takes ownership of RR.
+    \item occupied by x' in which a constest occurs (see fig 2)
+\end{enumerate}
+\end{enumerate}
+After the contest Energy costs are deducted from both the winner and the loser.
+The winner also has a chance to abandon the RR (based off a trait).
+If a death occurs (Energy(x) < 0) then x is removed from the individual pool.
+
+Figure 2. Contest Algorithm
+for an occupier x' vs an invader x, there are 3 stages:
+\begin{enumerate}
+    \item Display 1
+    \item Display 2
+    \item Fight
+\end{enumerate}
+In the display phases the probability of escalation is dependant on the visible traits of x and x'
+The rate of events increases as time t -> t1
+
+Traits:
+\begin{enumerate}
+    \item Visible to opponent:
+\begin{enumerate}
+    \item Size
+\end{enumerate}
+    \item invisible to opponent:
+\begin{enumerate}
+    \item Energy levels
+    \item Abandon Rate
+    \item Aggression
+    \item Encounter Rate
+    \item Experience (?)
+\end{enumerate}
+\end{enumerate}
 

### 2015-12-22 ms.tex
Expand text (population dynamics)

@@
 
-Evolution of animal personalities: \citep{Wolf-2007}
+Evolution of animal personalities: \citep{Wolf-2007,Wolf-2012} show can have
+coexistence of risky, explorative strategies and risk-averse strategies.
 
-Individual-based models of natural selection :\citep{MGonigle-2012}
+Animal personalities linked to other life history traits: \citep{Biro-2008}
+
+Individual-based models of natural selection: \citep{MGonigle-2012}
 
@@
 \subsection{Population dynamics}
+
+We consider a population of males competing for mates.
+The population is one with non-overlapping generations, and within each generation there is a predefined annual cycle.
+In spring, individuals hatch from eggs, grow throughout spring and summer to increase size, and then within a short period, the entire population mates, females lay eggs and everyone dies.
+The short duration of mating period is such that each male and female only mates once.
+The population then re-establishes from eggs the following spring.
+
+The reproductive success of males in the population is determined via their ability to compete for and hold nests.
+Denote $N$ to be the total number of available nests, $R_i$ to be the reproductive value of nest $i$ (number of offspring per nest per year), $\bar{R}$ to be the average reproductive value of all nests (number of offspring per nest per year), and $p_{\rm n}(R)$ to be the probability density distribution of nest quality with respect to $R$.
+The distribution of nest qulatiies is then set according to the distribution
+\begin{equation} \label{eq:pdf_R}
+Pr(R_i = x) =\bar(R) \, p_{\rm n}(R).
+\end{equation}
+
+The male population is divided into two groups: immature and mature males.
+Immature males devote all their efforts towards feeding, increasing their size.
+Males mature with a probability defined by their size and a maturation trait.
+Once they enter the mature population, males stop feeding and focus solely on obtaining nests.
+Nests can be vacant or occupied.
+Males can differ in the rate at which they search, and thus the rate at which they encounter nests.
+Empty nests can be immediately occupied with no minimal additional cost.
+Filled nests can only be obtained by entering into and winning a contest.
+Once occupied, males can choose to occupy a nest or continue searching.
+At the end of each season, each male has a fitness defined by the reproductive value of the nest it holds, $R_i$.
+
+Within the immature phase, we assume males grow with rate
+\begin{equation} \label{eq:growth}
+\frac{{\rm d} m}{{\rm d} t} = a m^{b}
+\end{equation}
+where $a$ and $b$ are metabolic constants. Following this equation, the size of individuals at time $t$, having started growing at time $0$ is then
+\begin{equation} \label{eq:growth}
+m(t) = \left(m(0)^{1-b} + t a(1-b)\right)^{\frac1{1-b}}.
+\end{equation}
 
@@
 
+\section{Figures}
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=15cm,height=15cm,keepaspectratio]{figures/event_algorithm}
+\caption{Algorithm for modelling dynamics within the ``mature'' male population.}
+\label{fig:events}
+\end{figure}
+\clearpage
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=15cm,height=15cm,keepaspectratio]{figures/contest_algorithm}
+\caption{Algorithm for modelling contests behaviour within the ``mature'' male population.}
+\label{fig:events}
+\end{figure}
+\clearpage
+
 \bibliography{refs}

### 2015-12-22 ms.tex
Minor text change
changes to traits

@@
 \section{Figures}
-Figure 1. Event Algorithm
+\subsection {Figure 1. Event Algorithm}
 Population state generates events, events are either:
 \begin{enumerate}
-    \item A new individual x has matured and entered the contest (based off of a trait(?))
-    \item Discovery event wherin both a nest RR (repoductive resource) and an individual x are drawn from a pool, that nest is either:
+    \item A new individual x has matured and entered the pool (based off of a trait(?))
+    \item Discovery event wherin both a nest RR (repoductive resource) and an individual x are drawn from a pool, the metabolic costs of searching are deducted from the time differences between events, that nest is either:
 \begin{enumerate}
@@
 The winner also has a chance to abandon the RR (based off a trait).
-If a death occurs (Energy(x) < 0) then x is removed from the individual pool.
+If a death occurs (Energy(x) less than 0) then x is removed from the individual pool.
 
-Figure 2. Contest Algorithm
+\subsection{Figure 2. Contest Algorithm}
 for an occupier x' vs an invader x, there are 3 stages:
@@
 In the display phases the probability of escalation is dependant on the visible traits of x and x'
-The rate of events increases as time t -> t1
+The rate of events increases as time t approaches t1
 
@@
 \begin{enumerate}
-    \item Size
+    \item Size, influenced by time entered maturity
 \end{enumerate}
@@
 \begin{enumerate}
-    \item Energy levels
-    \item Abandon Rate
-    \item Aggression
-    \item Encounter Rate
-    \item Experience (?)
+    \item Energy level, influenced by size and past actions
+    \item Abandon Rate, influences RR abandonment rate
+    \item Aggressiveness, influences escalation in contests
+    \item Encounter Rate, influences chance of being selected for an event
 \end{enumerate}

### 2015-12-22 ms.tex
Change growth model to exponential; density of maturing individuals
expoential much easier to solve! Little biological difference at this stage

@@
 % \doublespacing
-\linenumbers
+% \linenumbers
 
@@
 The reproductive success of males in the population is determined via their ability to compete for and hold nests.
-Denote $N$ to be the total number of available nests, $R_i$ to be the reproductive value of nest $i$ (number of offspring per nest per year), $\bar{R}$ to be the average reproductive value of all nests (number of offspring per nest per year), and $p_{\rm n}(R)$ to be the probability density distribution of nest quality with respect to $R$.
+Denote $N$ to be the total number of available nests, $R_i$ to be the reproductive value of nest $i$ (number of offspring per nest per year), $\bar{R}$ to be the average reproductive value of all nests (number of offspring per nest per year), and $p(R)$ to be the probability density distribution of nest quality with respect to $R$.
 The distribution of nest qulatiies is then set according to the distribution
 \begin{equation} \label{eq:pdf_R}
-Pr(R_i = x) =\bar(R) \, p_{\rm n}(R).
+Pr(R_i = x) =\bar{R} \, p(R).
 \end{equation}
@@
 
-Within the immature phase, we assume males grow with rate
+Within the immature phase, we assume males grow exponentially with rate
 \begin{equation} \label{eq:growth}
-\frac{{\rm d} m}{{\rm d} t} = a m^{b}
+\frac{{\rm d} m}{{\rm d} t} = a \, m
 \end{equation}
-where $a$ and $b$ are metabolic constants. Following this equation, the size of individuals at time $t$, having started growing at time $0$ is then
+where $a$ is mass-based rate of growth. Integrating eq. \ref{eq:growth}, the size of individuals at time $t$, having started growing at time $0$ is given by
 \begin{equation} \label{eq:growth}
-m(t) = \left(m(0)^{1-b} + t a(1-b)\right)^{\frac1{1-b}}.
+m(t) = m(0)\exp(a \, t).
+\end{equation}
+If we assume a constant mortality rate $k_i$ during the immature phase, the probability of males surviving to time $t$ is given by
+\begin{equation} \label{eq:surv_immature}
+S(t) = \exp(-k_i \, t).
+\end{equation}
+The time $\tau$ taken for a male to grow from $m(0) \rightarrow m$ is then
+\begin{equation} \label{eq:tau}
+\tau(m_0, m) = \frac1{a}\ln\left(\frac{m}{m_0}\right).
+\end{equation}
+Assuming all remained in the immature phase, the density $n(m,t)$ of males with size $m$ at time $t$, would be given by \cite{DeRoos-1997}:
+\begin{equation} \label{eq:n1}
+n(m,t | {\rm immature}) = \frac{B(t - \tau(m_0, m))}{g(m)} \exp\left(-\int_{m_0}^m \frac{k_i}{g(m^\prime)} {\rm d}m^\prime \right),
+\end{equation}
+where $B(t)$ is the population birth rate at time $t$.
+Substituting from above, we obtain
+\begin{equation} \label{eq:n1}
+n(m, t \, | {\rm immature}) = \frac{B(t - \tau(m_0, m))}{a\, m} \left(\frac{m}{m_0}\right)^{\frac{k_i}{a}}.
+\end{equation}
+
+Finally, we assume males transition from immature $\rightarrow$ mature according to logistic probability function
+\begin{equation}\label{eq:allocation}
+Pr({\rm mature} \, | m) = \frac1{1 + \exp\left(\alpha \left(1 - m / m_{\rm mat}\right)\right)},
+\end{equation}
+where $m_{\rm mat}$ is the size at which 50\% of the population is mature and $\alpha$ determines the width of the curve.
+
+Thus we can calculate the flux of males size $m$ into the reproductive pool at time $t$ as
+\begin{equation}\label{eq:flux}
+n(m, t\, | {\rm immature}) \, \frac{{\rm d}}{{\rm d} m} Pr({\rm mature}\,  | m).
 \end{equation}
@@
 
-\subsection{Numerical analysis}
+\subsection{Algorithms}
 
-\section{Figures}
-\subsection {Figure 1. Event Algorithm}
-Population state generates events, events are either:
+\subsection{Event Algorithm}
+See Fig. \ref{fig:events}. Population state generates events, events are either:
 \begin{enumerate}
@@
 
-\subsection{Figure 2. Contest Algorithm}
-for an occupier x' vs an invader x, there are 3 stages:
+\subsection{Contest Algorithm}
+See Fig. \ref{fig:contests}.
+For an occupier x' vs an invader x, there are 3 stages:
 \begin{enumerate}
@@
 \caption{Algorithm for modelling contests behaviour within the ``mature'' male population.}
-\label{fig:events}
+\label{fig:contests}
 \end{figure}

### 2015-12-22 ms.tex
More on population dynamics
- revert decision about expoential growth as don't need analyitcal solution for density
- details about Gillespie's minimal process method

@@
 
-\subsection{Population dynamics}
+% Verbal description of model
 
@@
 The population is one with non-overlapping generations, and within each generation there is a predefined annual cycle.
-In spring, individuals hatch from eggs, grow throughout spring and summer to increase size, and then within a short period, the entire population mates, females lay eggs and everyone dies.
+In spring, up to $K$individuals hatch from eggs, grow throughout spring and summer to increase size, and then within a short period, the entire population mates, females lay eggs and everyone dies.
 The short duration of mating period is such that each male and female only mates once.
@@
 
-The reproductive success of males in the population is determined via their ability to compete for and hold nests.
-Denote $N$ to be the total number of available nests, $R_i$ to be the reproductive value of nest $i$ (number of offspring per nest per year), $\bar{R}$ to be the average reproductive value of all nests (number of offspring per nest per year), and $p(R)$ to be the probability density distribution of nest quality with respect to $R$.
-The distribution of nest qulatiies is then set according to the distribution
-\begin{equation} \label{eq:pdf_R}
-Pr(R_i = x) =\bar{R} \, p(R).
-\end{equation}
-
+The reproductive success of males in the population is determined via their ability to compete for and hold $N$ nests.
+Nests differ in quality (potential number of offspring per nest per year).
 The male population is divided into two groups: immature and mature males.
@@
 Once they enter the mature population, males stop feeding and focus solely on obtaining nests.
-Nests can be vacant or occupied.
-Males can differ in the rate at which they search, and thus the rate at which they encounter nests.
-Empty nests can be immediately occupied with no minimal additional cost.
+Nests can be vacant or occupied and males are either ``searching'' or ``occupying'' a nest.
+Males differ in the rate at which they search, and thus the rate at which they encounter nests.
+Empty nests can be immediately occupied with no additional cost.
 Filled nests can only be obtained by entering into and winning a contest.
@@
 
-Within the immature phase, we assume males grow exponentially with rate
+\subsection{Habitat}
+
+Denote $N$ to be the total number of available nests, $R_i$ to be the reproductive value of nest $i$ (number of offspring per nest per year), $\bar{R}$ to be the average reproductive value of all nests (number of offspring per nest per year), and $p(R)$ to be the probability density distribution of nest quality with respect to $R$.
+The distribution of nest qualities is then set according to the distribution
+\begin{equation} \label{eq:pdf_R}
+Pr(R_i = x) =\bar{R} \, p(R).
+\end{equation}
+The total number of male offspring produced in each generation is then
+\begin{equation} \label{eq:pdf_R}
+K = \bar{R} N.
+\end{equation}
+
+\subsection{Immature phase}
+
+Within the immature phase, we assume males grow with rate
 \begin{equation} \label{eq:growth}
-\frac{{\rm d} m}{{\rm d} t} = a \, m
+\frac{{\rm d} m}{{\rm d} t} = a \, m^ b
 \end{equation}
@@
 \begin{equation} \label{eq:growth}
-m(t) = m(0)\exp(a \, t).
+m(t) = \left(m(0)^{1-b} + t a(1-b)\right)^{\frac1{1-b}},
 \end{equation}
+while the he time $\tau$ taken for a male to grow from $m(0) \rightarrow m$ is then
+\begin{equation} \label{eq:tau}
+\tau(m_0, m) = \frac1{a(1-b)}\left(m ^{1-b} - m_0 ^{1-b}\right).
+\end{equation}
+
 If we assume a constant mortality rate $k_i$ during the immature phase, the probability of males surviving to time $t$ is given by
@@
 S(t) = \exp(-k_i \, t).
-\end{equation}
-The time $\tau$ taken for a male to grow from $m(0) \rightarrow m$ is then
-\begin{equation} \label{eq:tau}
-\tau(m_0, m) = \frac1{a}\ln\left(\frac{m}{m_0}\right).
-\end{equation}
-Assuming all remained in the immature phase, the density $n(m,t)$ of males with size $m$ at time $t$, would be given by \cite{DeRoos-1997}:
-\begin{equation} \label{eq:n1}
-n(m,t | {\rm immature}) = \frac{B(t - \tau(m_0, m))}{g(m)} \exp\left(-\int_{m_0}^m \frac{k_i}{g(m^\prime)} {\rm d}m^\prime \right),
-\end{equation}
-where $B(t)$ is the population birth rate at time $t$.
-Substituting from above, we obtain
-\begin{equation} \label{eq:n1}
-n(m, t \, | {\rm immature}) = \frac{B(t - \tau(m_0, m))}{a\, m} \left(\frac{m}{m_0}\right)^{\frac{k_i}{a}}.
 \end{equation}
@@
 
-Thus we can calculate the flux of males size $m$ into the reproductive pool at time $t$ as
-\begin{equation}\label{eq:flux}
-n(m, t\, | {\rm immature}) \, \frac{{\rm d}}{{\rm d} m} Pr({\rm mature}\,  | m).
-\end{equation}
+\subsection{Mature phase}
 
-\subsection{Encounters}
+Need equations for
+
+\begin{itemize}
+    \item Populations of ``searching'' males and  ``occupying'' males
+    \item Rate at which ``searching'' males encounter empty nests
+    \item Rate at which ``searching'' males encounter occupied nests
+    \item Rate at which ``occupying'' males abandon an occupied nest
+    \item Outcomes of contest between two males
+\end{itemize}
 
@@
 
+What types of mutations?
+
 \subsection{Algorithms}
+Simulating the population dynamics requires accounting for four types of events:
+\begin{itemize}
+    \item Immature individuals become mature
+    \item Searching males encounter an empty nest
+    \item Searching males encounter an occupied nest and contest it
+    \item Occupying males abandon an occupied nest
+\end{itemize}
+
+An efficient method for stepping the population is to step between successive events.
+So rather than stepping the system over a given time interval and considering any transitions that may occur in that time period, the system is stepped between successive events by drawing the next event and the time at which the event occurs from appropriate distributions.
+We can identify time to next event using Gillespie minimal process method \citep{Gillespie-1976}, which indicates that with a given event rate, the distribution of waiting times follows an exponential distribution.
 

### 2015-12-24 ms.tex
variable table

@@
 \end{equation}
-while the he time $\tau$ taken for a male to grow from $m(0) \rightarrow m$ is then
+while the time $\tau$ taken for a male to grow from $m(0) \rightarrow m$ is then
 \begin{equation} \label{eq:tau}
@@
 \end{enumerate}
+\clearpage
+
+I am not sure if this is the way to do this, 
+or if there is a way to do it with refs (?)
+\begin{table}[h!]
+    \caption{Variable names and definitions in the model.}
+    \centering
+    \begin{tabular}{c | c | l }
+        \hline
+        Symbol & unit & Description\\
+        \hline
+        \hline
+        $K$ & & number of individual males \\
+        $N$ & & number of nests \\
+        $R_i$ & & repoductive value of nest of nest $i$\\
+        $\bar{R}$ & & average repoductive value of nest\\
+        \hline
+        $m$ & $kg$ & mass\\
+        \hline
+    \end{tabular}
+\end{table}
 

### 2015-12-27 ms.tex
some mature equations and table updated.

@@
 
-Need equations for
+/* this is just for consideration I am not sure if it will be needed or useful */
+Number of mature males at time $t$:
+\begin{equation} \label{eq:num_mature}
+    K_{mature}(t) = \lfloor{K_{initial}(Pr({\rm mature} \, | m(t)) - S(t))}\rfloor
+\end{equation}
+A searching individual could encounter nests at a rate equal to its ``exploration trait'' $\lambda_i$:
+\begin{equation} \label{eq:encounter rate}
+    {\rm Encounters} = \lfloor\lambda_it_{searching}\rfloor
+\end{equation}
+/* end */
 
+Need equations for:
 \begin{itemize}
@@
         \hline
+        $t_m$ & & the time at which the females begin to mature\\
+        $S(t)$ = exp($-k_it$) & & The probability of an individual surviving to time t.\\
+        \hline
         $m$ & $kg$ & mass\\
+        $m_mat$ & $kg$ & The size at which 50\% of the population is mature.\\ 
         \hline

### 2016-01-06 ms.tex
abandon, encounter, equations and exploratory python script looking at encounter paramters

@@
 \end{equation}
-A searching individual could encounter nests at a rate equal to its ``exploration trait'' $\lambda_i$:
+A searching individual has a chance to encounter a nest equal to its ``exploration trait'' $\lambda_i$:
+\begin{equation} \label{eq:encounter probability}
+    Pr({\rm Encounter}) = \lambda_i
+\end{equation}
+The number of encounters that a searching individual has engaged in is:
 \begin{equation} \label{eq:encounter rate}
-    {\rm Encounters} = \lfloor\lambda_it_{searching}\rfloor
+    {\rm Encounters} = \lfloor\lambda_i \cdot t_{searching}\rfloor
+\end{equation}
+the average exploration trait across a cohort is $\bar\lambda$.
+
+similarly for males abandoning nests males the probability of abandoning a nest $j$, is a function of the  the reproductive value of the nest $R_j$ and the abandon trait of the individual $\psi_i$:
+
+assumption: $1 < R_j$, $0 < \psi_i < 1$
+
+\begin{equation} \label{eq:abandon}
+    Pr({\rm abandon} | R_j ) = \frac{1}{1 + \psi_i \cdot R_j}
 \end{equation}

### 2016-01-11 ms.tex
small change to abandon

@@
 
-/* this is just for consideration I am not sure if it will be needed or useful */
+/* this is just for consideration I am not sure if it will be needed or useful */\\ \\
 Number of mature males at time $t$:
@@
 
-similarly for males abandoning nests males the probability of abandoning a nest $j$, is a function of the  the reproductive value of the nest $R_j$ and the abandon trait of the individual $\psi_i$:
+Similarly for males abandoning nests males the probability of abandoning a nest $j$, is a function of the  the reproductive value of the nest $R_j$ and the abandon trait of the individual $\psi_i$:
 
-assumption: $1 < R_j$, $0 < \psi_i < 1$
+assumption: $1 < R_j$,  $1 < \psi_i$
 

### 2016-01-11 ms.tex
tex formatting error
escalation probabilities

@@
 
-/* this is just for consideration I am not sure if it will be needed or useful */\\ \\
+/* this is just for consideration I am not sure if it will be needed or useful */\\
 Number of mature males at time $t$:
@@
 See Fig. \ref{fig:contests}.
-For an occupier x' vs an invader x, there are 3 stages:
+For a defender $x_{def}$ vs an attacker $x_{atk}$, there are 3 stages:
+each individual has an aggression trait $a$ and a mass $m$.\\
+\begin{equation}\label{eq:logit}
+    logit^{-1}(\alpha) = \frac{exp(\alpha)}{1 + exp(\alpha)}
+\end{equation}
+
 \begin{enumerate}
     \item Display 1
-    \item Display 2
+        Outcomes:
+        \begin{itemize}
+            \item $x_{atk}$ escalates: 
+                $Pr(\rm{escalation}) =  logit^{-1}(a_{atk} \cdot m_{atk} - m_{def})$
+                or leaves.
+            \item $x_{def}$ escalates:
+                $Pr(\rm{escalation}) =  logit^{-1}(a_{def} \cdot m_{def} - m_{atk})$ 
+                or leaves.
+            \item if the attacker leaves the defender keeps the nest
+            \item if the defender leaves the attacker keeps the nest
+            \item the attacker is evaluated first
+            \item if they both escalate then the next stage occurs
+        \end{itemize}
+    \item Display 2:
+        similar to stage 1, this could involve the $R$ value of the nest, or include the oponents aggression levels or energy reseves(?).
     \item Fight
 \end{enumerate}
-In the display phases the probability of escalation is dependant on the visible traits of x and x'
+The cost of the fighting stage is much greater than the cost of the display stage.
+
+In the display phases the probability of escalation is dependent on the visible traits of x and x'
 The rate of events increases as time t approaches t1

### 2016-01-11 ms.tex
searching and occupying odes

@@
 
+Searching individuals:
+\begin{equation} \label{eq:searching}
+    \frac{dS}{dt} = aS -bS^2 - \frac{cS}{1+O} - eS - gO + fO
+\end{equation}
+Occupying individuals
+\begin{equation} \label{eq:occupying}
+    \frac{dO}{dt} = \frac{cS}{1+O} - eO - fO
+\end{equation}
+
+$aS - bS^2$ is individuals becoming mature.\\
+$\frac{cS}{1+O}$ is searching individuals becoming occupiers through finding nests that are unoccuied.\\
+$e$ is metabolic death rate\\
+$f$ is the abandon rate\\
+$g$ is the rate at which occupiers die in a contest\\
+
 Need equations for:
@@
 For a defender $x_{def}$ vs an attacker $x_{atk}$, there are 3 stages:
-each individual has an aggression trait $a$ and a mass $m$.\\
+each individual has an aggression trait $a$ and a mass $m$.
 \begin{equation}\label{eq:logit}
@@
 
-I am not sure if this is the way to do this, 
-or if there is a way to do it with refs (?)
+I am not sure if this is the best way to do this, the tables in plant are more aesthetic. 
 \begin{table}[h!]

### 2016-01-11 ms.tex
small changes to coeffs in the S/O DE's

@@
 \begin{equation} \label{eq:searching}
-    \frac{dS}{dt} = aS -bS^2 - \frac{cS}{1+O} - eS - gO + fO
+    \frac{dS}{dt} = aS - bS^2 - \frac{cS}{1+eO} - fS - hO + gO
 \end{equation}
-Occupying individuals
+Occupying individuals:
 \begin{equation} \label{eq:occupying}
-    \frac{dO}{dt} = \frac{cS}{1+O} - eO - fO
+    \frac{dO}{dt} = \frac{cS}{1+eO} - fO - gO
 \end{equation}
@@
 $aS - bS^2$ is individuals becoming mature.\\
-$\frac{cS}{1+O}$ is searching individuals becoming occupiers through finding nests that are unoccuied.\\
-$e$ is metabolic death rate\\
-$f$ is the abandon rate\\
-$g$ is the rate at which occupiers die in a contest\\
+$\frac{cS}{1+eO}$ is searching individuals becoming occupiers through finding nests that are unoccuied.\\
+$f$ is the metabolic death rate\.\
+$g$ is the abandon rate.\\
+$h$ is the rate at which occupiers die in a contest.\\
 

### 2016-01-14 python/generation.py
very basic contests, fixed bug with death counts

@@ -65,7 +65,10 @@ class Generation:
         self.searching.append(self.immature.pop(0))
 
         # until the females mature: 
+        i = 0
         while (self.time < f_time):
+            i += 1
+
             # check next mature male
             if self.immature: # check not null
                 if self.immature[0].maturation_time <= self.time:
@@ -80,7 +83,7 @@ class Generation:
                     if not n.occupier.is_alive(): # remove the dead
                         m = n.eject()
                         self.killed += 1
-                        print "male %s in nest %s has died at t = %s" % (
+                        print "%s, male %s in nest %s has died at t = %s" % ( i,
                             m.id,
                             n.id,
                             self.time)
@@ -89,9 +92,6 @@ class Generation:
             # iterate over searching males
             self.searching_step(dt)
 
-            # remove occupying males
-            self.searching = filter(lambda m: not m.occupying, self.searching)
-            
             self.log_cohort()
 
             self.time += dt
@@ -103,10 +103,14 @@ class Generation:
 
     # itereates over all searching males
     def searching_step(self, dt):
-        for m in self.searching:
-            if m.search(dt):
+        for m in self.searching[:]:
+            if not m.is_alive():
+                print "male %s has died at %s" % (m.id, self.time)
+                self.searching.remove(m)
+                self.killed += 1
+            elif m.search(dt):
+                # select a nest at randoms
                 index = int(uni.rvs() * len(self.nests))
-                # select a nest at random
                 print "male %s has discovered nest %s at %s" % (
                     m.id,
                     index,
@@ -116,23 +120,28 @@ class Generation:
                     print "\tnest %s is occupied by %s, contest" % (
                         index, 
                         nest.occupier.id)
-                    print "\tno change"
+                    loser = nest.contest(m)
+                    if loser.id != m.id:
+                        self.searching.insert(0,loser)
+                        self.searching.remove(m)
                     self.contests += 1
+                    print "\tnest %s is occupied by %s" % (
+                        index, 
+                        nest.occupier.id)
+
                 else:
-                    m.occupying = True
                     nest.occupy(m)
+                    self.searching.remove(m)
                     print "\tnest %s is now occupied by %s" % (
                         index, 
                         m.id)
 
-            if not m.is_alive():
-                print "male %s has died at %s" % (m.id, self.time)
-
-
-            # remove dead searching males
-            self.killed += len(self.searching)
[…]

### 2016-01-14 python/male.py

@@ -1,7 +1,7 @@
 #!/usr/bin/python
 
 import math
-from scipy.stats import norm, logistic
+from scipy.stats import norm, logistic, uniform
 
 # a male individual in the population
 class Male(object):
@@ -36,12 +36,15 @@ class Male(object):
     # constructor
     def __init__(self, params, id, mom = None, dad = None):
         self.id = id
-        if not mom and not dad: # first gen no breeding
+        if not mom and not dad: # first gen: no breeding
 
             # pull the exploration trait from the normal distrobution
             self.exploration = abs(norm.rvs(
                 params["exploration_mean"],
                 params["exploration_sd"]))
+
+            # pull the aggression from the normal distrobution
+            self.aggro = params["aggression_max"]*uniform.rvs()
 
             # countdown timer to next event
             self.tt_event = self.exploration

### 2016-01-14 python/nest.py

@@ -1,25 +1,104 @@
 #!/usr/bin/python
 import sys
-
+from scipy.stats import logistic, uniform
 """ a nest has a RR and female genetics and a male occupier"""
 class Nest(object):
 
     def __init__(self, params, id):
         self.id = id
         self.rr = 0
+        self.fight_cost = params["fight_cost"]
+        self.display_1_cost = params["display_1_cost"]
         self.occupier = None
 
     def occupied(self):
         return bool(self.occupier)
 
+    def contest(self, attacker):
+        defender = self.occupier
+
+        # energy costs associated with this contest
+        costs = 0
+        loser = None
+
+        if not defender.is_alive():
+            loser = defender
+
+        if not loser:
+            loser = self.display_1(attacker, defender)
+            costs += self.display_1_cost
+
+        # fight
+        if not loser:
+            loser = self.fight(attacker, defender)
+            costs += self.fight_cost
+        
+        attacker.energy -= costs
+        defender.energy -= costs
+
+        if loser.id == defender.id:
+            self.eject()
+            self.occupy(attacker)
+        #raw_input()
+        return loser
+
+    # the first display phase, the loser is returned or None
+    def display_1(self, attacker, defender):
+        ## attacker goes first
+        # prob the attacker will escalate
+        print "fight!"
+        print "attacker mass = %s, attacker aggro = %s" % (attacker.mass, attacker.aggro)
+        print "defender mass = %s, defender aggro = %s" % (defender.mass, defender.aggro)
+        
+        atk_escalation = logistic.cdf(attacker.mass * attacker.aggro - defender.mass)
+        rng = uniform.rvs()
+        
+        if rng > atk_escalation :
+            # attacker decides to escalate
+            print "%s > %s" % (rng, atk_escalation)
+            print "\tattacker backs down"
+            return attacker
+        else: 
+            print "%s <= %s" % (rng, atk_escalation)
+            print "\tattacker escalates"
+        
+        def_escalation = logistic.cdf(defender.mass * defender.aggro - attacker.mass)
+        rng = uniform.rvs()
+        if rng > def_escalation:
+            print "%s > %s" % (rng, def_escalation)
+            print "\tdefender backs down"
+            return defender
+        else:
+            print "%s <= %s" % (rng, def_escalation)
+            print "\tdefender escalates"
+
+        return None
+
+    def fight(self, attacker, defender):
+        print "fight"
+        defender_wins = logistic.cdf(defender.mass - attacker.mass)
[…]

### 2016-01-14 ms.tex
info re: SSA -> ms.tex

@@
 \end{enumerate}
-After the contest Energy costs are deducted from both the winner and the loser.
-The winner also has a chance to abandon the RR (based off a trait).
-If a death occurs (Energy(x) less than 0) then x is removed from the individual pool.
+
+In the SSA algorithm \citep{Gillespie-1976} there are 6 different \"species\", representing different states:
+\begin{enumerate}
+    \item $S_1 = $ immature male
+    \item $S_2 = $ mature searching male
+    \item $S_3 = $ unoccupied nest
+    \item $S_4 = $ occupied nest
+    \item $S_5 = $ recently defeated male
+    \item $S_6 = $ recently changed nest
+\end{enumerate}
+
+these species have $R_\mu$ different interactions
+\begin{table}[h!]
+    \caption{state changes}
+    \centering
+    \begin{tabular}{c | c | c | c | l }
+        \hline
+        Reaction num & reactants  & & products & Description\\
+        \hline
+        \hline
+            $R_1$       &   $S_1$       & $--c_1-->$       &   $ S_2$      & maturation\\
+            $R_2$       &   $S_2 + S_3$ & $--c_2-->$      &   $S_4 $      & occupation\\
+            $R_3$       &   $S_2 + S_4$ & $--c_3-->$      &   $S_4 + S_5$ & contest\\
+            $R_4$       &   $S_5$       & $--c_4-->$      &   $S_2$       & recovery\\
+            $R_5$       &   $S_2 + S_4$ & $--c_5-->$      &   $S_4$       & contest with death\\
+            $R_6$       &   $S_4$       & $--c_6-->$      &   $S_2 + S_3$ & abandon event\\
+            $R_7$       &   $S_1$       & $--c_7-->$      &   $0$         & metab0lic / predation event\\
+            $R_8$       &   $S_2$       & $--c_8-->$      &   $0$         & predation event\\
+            $R_9$       &   $S_4$       & $--c_9-->$      &   $S_3$       & metab0lic / predation event\\
+            $R_{10}$    &   $S_6$       & $--c_{10}-->$   &   $S_4$       & nest handover\\
+        \hline
+    \end{tabular}
+\end{table}
+
+$c_n$ is the stochatic reaction constant
 
@@
 
+$m$ is mass
+$a$ is aggression trait
 \begin{enumerate}
@@
     \item Fight
+        So far I have something like
+        \begin{itemize}
+            \item $x_{def}$ wins:
+            $Pr(\rm{def}) = logit{-1}(m_{def} - m_{atk})$
+        \end{itemize}
 \end{enumerate}
@@
 In the display phases the probability of escalation is dependent on the visible traits of x and x'
-The rate of events increases as time t approaches t1
+
+After the contest Energy costs are deducted from both the winner and the loser.
+The winner also has a chance to abandon the RR (based off a trait).
+If a death occurs (Energy(x) less than 0) then x is removed from the individual pool.
 

### 2016-01-14 python/generation.py
more plots :D

@@ -197,15 +197,22 @@ class Generation:
         plt.plot(times, mean_energy_searching, label = "searching")
         plt.plot(times, mean_energy_occupying, label = "occupying")
         plt.legend(loc = 2)
-
-        plt.title("trait values in winners")
-        plt.xlabel("trait values")
-        plt.ylabel("number")
         plt.show()
+
 
         occupying_males = [ n.occupier for n in self.nests if n.occupied()]
         occupying_exploration_trait = [m.exploration for m in occupying_males]
-        plt.hist(occupying_exploration_trait, 15, label = "exploration")
-
+        occupying_aggro_trait = [m.aggro / self.params["aggression_max"] for m in occupying_males]
+        
+        bins = np.linspace(-10, 10, )
+        plt.hist(occupying_exploration_trait, self.params["trait_bins"], alpha = 0.5, label = "exploration")
+        plt.hist(occupying_aggro_trait, self.params["trait_bins"], alpha = 0.75, label = "aggression")
         plt.legend(loc = 2)
+        plt.show()
+        
+
+        plt.plot(occupying_exploration_trait, occupying_aggro_trait, 'ro')
+        plt.title("trait values of individuals")
+        plt.xlabel("exploration")
+        plt.ylabel("aggression")
         plt.show()

### 2016-01-14 python/male.py

@@ -37,17 +37,14 @@ class Male(object):
     def __init__(self, params, id, mom = None, dad = None):
         self.id = id
         if not mom and not dad: # first gen: no breeding
-
-            # pull the exploration trait from the normal distrobution
-            self.exploration = abs(norm.rvs(
-                params["exploration_mean"],
-                params["exploration_sd"]))
+            
+            self.exploration = uniform.rvs()
 
             # pull the aggression from the normal distrobution
             self.aggro = params["aggression_max"]*uniform.rvs()
 
             # countdown timer to next event
-            self.tt_event = self.exploration
+            self.tt_event = 1.0
             
             self.metabolic_cost_search = params["metabolic_cost_search"]
             self.metabolic_cost_occupy = params["metabolic_cost_occupy"]
@@ -84,10 +81,10 @@ class Male(object):
         self.tt_event -= dt
         self.energy -= dt * self.metabolic_cost_search
         if self.tt_event <= 0:
-            self.tt_event = self.exploration
-            return True
-        else:
-            return False
+            self.tt_event = 1.0
+            if uniform.rvs() < self.exploration:
+                return True
+        return False
 
 
     # the outcome of a contest is decided here

### 2016-01-14 python/generation.py
some optims some plots, debug mode

@@ -4,7 +4,7 @@ from nest import Nest
 
 from scipy.stats import uniform as uni
 import numpy as np
-
+import sys
 import matplotlib.pyplot as plt
 
 # a generation in the minimal model
@@ -15,6 +15,7 @@ import matplotlib.pyplot as plt
 # TODO: 
 # step directly to next event not through time_steps
 # nest collisions (this shouldn't be an issue for small deltas)
+# move logs into a new class
 
 class Generation:
 
@@ -27,6 +28,7 @@ class Generation:
         self.contests = 0
         self.num_matured = 0
         self.killed = 0
+        self.debug = params["debug"]
 
         self.logs = []
 
@@ -55,38 +57,47 @@ class Generation:
         f_time = self.params["time_female_maturity"]
 
         # print the cohort
-        for x in self.immature:
-            print x.to_string() 
+        if self.debug:
+            for x in self.immature:
+                print x.to_string() 
         
         # start the generation when the first male matures:
         self.time = self.immature[0].maturation_time
         self.num_matured += 1
-        print "start time\t", self.time
+        if self.debug:
+            print "start time\t", self.time
+
         self.searching.append(self.immature.pop(0))
-
+        step = 0
         # until the females mature: 
-        i = 0
         while (self.time < f_time):
-            i += 1
-
-            # check next mature male
-            if self.immature: # check not null
-                if self.immature[0].maturation_time <= self.time:
-                    self.searching.append(self.immature.pop(0))
-                    self.num_matured += 1
+            step += 1
+            # check the next mature male
+            maturing = filter(lambda m: m.maturation_time <= self.time, self.immature)
+            for m in maturing:
+                self.searching.append(m)
+                self.immature.remove(m)
+                if self.debug:
+                    print "male matured at %s" % self.time
+                self.num_matured += 1
             
             
-            # iterate over nests subtracting metabolic costs from occupying males
+            # iterate over nests subtracting metabolic costs from occupying
+            # males
+            self.occupying = 0
             for n in self.nests:
                 if n.occupied():
                     n.occupier.occupy(dt)
-                    if not n.occupier.is_alive(): # remove the dead
+                    if not n.occupier.is_alive(): # remove the dead males
                         m = n.eject()
                         self.killed += 1
-                        print "%s, male %s in nest %s has died at t = %s" % ( i,
-                            m.id,
-                            n.id,
-                            self.time)
+                        if self.debug:
[…]

### 2016-01-14 python/male.py

@@ -14,16 +14,16 @@ class Male(object):
         
         heriditable traits:
             @exploration 
-            psi
-            aggression
+            psi / abandon
+            @aggression
 
         costs:
-            metabolic
+            @metabolic
             predation
 
         events:
-            contests
-            searching
+            @contests
+            @searching
             abandoning
 
         genetics:

### 2016-01-14 python/nest.py

@@ -2,6 +2,10 @@
 import sys
 from scipy.stats import logistic, uniform
 """ a nest has a RR and female genetics and a male occupier"""
+
+""" TODO
+    abandoning nests
+"""
 class Nest(object):
 
     def __init__(self, params, id):
@@ -10,6 +14,7 @@ class Nest(object):
         self.fight_cost = params["fight_cost"]
         self.display_1_cost = params["display_1_cost"]
         self.occupier = None
+        self.debug = params["debug"]
 
     def occupied(self):
         return bool(self.occupier)
@@ -46,45 +51,53 @@ class Nest(object):
     def display_1(self, attacker, defender):
         ## attacker goes first
         # prob the attacker will escalate
-        print "fight!"
-        print "attacker mass = %s, attacker aggro = %s" % (attacker.mass, attacker.aggro)
-        print "defender mass = %s, defender aggro = %s" % (defender.mass, defender.aggro)
+        if self.debug:
+            print "fight!"
+            print "attacker mass = %s, attacker aggro = %s" % (attacker.mass, attacker.aggro)
+            print "defender mass = %s, defender aggro = %s" % (defender.mass, defender.aggro)
         
         atk_escalation = logistic.cdf(attacker.mass * attacker.aggro - defender.mass)
         rng = uniform.rvs()
         
         if rng > atk_escalation :
             # attacker decides to escalate
-            print "%s > %s" % (rng, atk_escalation)
-            print "\tattacker backs down"
+            if self.debug:
+                print "%s > %s" % (rng, atk_escalation)
+                print "\tattacker backs down"
             return attacker
-        else: 
-            print "%s <= %s" % (rng, atk_escalation)
-            print "\tattacker escalates"
+        else:
+            if self.debug:
+                print "%s <= %s" % (rng, atk_escalation)
+                print "\tattacker escalates"
         
         def_escalation = logistic.cdf(defender.mass * defender.aggro - attacker.mass)
         rng = uniform.rvs()
         if rng > def_escalation:
-            print "%s > %s" % (rng, def_escalation)
-            print "\tdefender backs down"
+            if self.debug:
+                print "%s > %s" % (rng, def_escalation)
+                print "\tdefender backs down"
             return defender
         else:
-            print "%s <= %s" % (rng, def_escalation)
-            print "\tdefender escalates"
+            if self.debug:
+                print "%s <= %s" % (rng, def_escalation)
+                print "\tdefender escalates"
 
         return None
 
[…]

### 2016-01-14 python/simulation.py

@@ -2,7 +2,6 @@
 
 import json
 import numpy as np
-
 from generation import Generation
 
 
@@ -28,9 +27,11 @@ class Simulation:
 
 def main():
     # read in parameters from file
+
     params = json.loads(open("parameters.json").read())
     np.random.seed(seed=params["random_seed"])
     # create a new simulation object
+
     sim = Simulation(params)
     sim.start()
 

### 2016-01-14 python/generation.py
logger added

@@ -1,11 +1,9 @@
-##!/usr/bin/env 
+#!/usr/bin/env python
 from male import Male
 from nest import Nest
+from logger import Logger
 
-from scipy.stats import uniform as uni
-import numpy as np
-import sys
-import matplotlib.pyplot as plt
+from scipy.stats import uniform
 
 # a generation in the minimal model
 # aggregates males
@@ -14,10 +12,9 @@ import matplotlib.pyplot as plt
 
 # TODO: 
 # step directly to next event not through time_steps
-# nest collisions (this shouldn't be an issue for small deltas)
 # move logs into a new class
 
-class Generation:
+class Generation (object):
 
     # constructor
     # params is the parameter dict
@@ -30,7 +27,7 @@ class Generation:
         self.killed = 0
         self.debug = params["debug"]
 
-        self.logs = []
+        self.logger = Logger(self)
 
         if not prev_gen: # no genetics
             # initialise some lists
@@ -48,9 +45,6 @@ class Generation:
             pass
 
         self.run()
-        # for each male increment exploration by their lambda
-        # when exploration reaches some value
-        # pick a nest and occupy or contest
 
     def run(self):
         dt = self.params["time_step"]
@@ -103,15 +97,12 @@ class Generation:
             # iterate over searching males
             self.searching_step(dt)
 
-            self.log_cohort()
+            self.logger.log_cohort()
 
             self.time += dt
         
         if self.params["generation_plot"]:
-            self.plot_cohort()
-
-
-            # occupying
+            self.logger.plot_cohort()
 
     # itereates over all searching males
     def searching_step(self, dt):
@@ -126,7 +117,7 @@ class Generation:
 
             elif m.search(dt):
                 # select a nest at random
-                index = int(uni.rvs() * len(self.nests))
+                index = int(uniform.rvs() * len(self.nests))
                 if self.debug:
                     print "male %s has discovered nest %s at %s" % (
                         m.id,
@@ -147,7 +138,7 @@ class Generation:
                         print "\tnest %s is occupied by %s" % (
                             index, 
                             nest.occupier.id)
-
+                # un occupied nests are taken over by the searching male
                 else:
                     nest.occupy(m)
                     self.occupying += 1
@@ -157,84 +148,3 @@ class Generation:
                             index, 
                             m.id)
 
-    # logs stats about the cohort to a list
-    def log_cohort(self):
-        row = {}
[…]

### 2016-01-14 python/logger.py

@@ -1,0 +1,128 @@
+# a class to hold data logs of generations:
+
+import numpy as np
+import sys
+import matplotlib.pyplot as plt
+
+class Logger(object):
+    """docstring for logs"""
+    def __init__(self, generation):
+        self.generation = generation
+        self.data = []
+
+    # this function should be called each time step
+    # to log the generation to the data list
+    def log_cohort(self):
+        row = {}
+        # average energy of searching males
+        row["time"] = self.generation.time
+        row["searching"] = len(self.generation.searching)
+        row["occupying"] = self.generation.occupying
+        row["contests"] = self.generation.contests 
+        row["num_matured"] = self.generation.num_matured
+        row["killed"] = self.generation.killed
+
+        row["energy_searching"] = np.mean([m.energy for m in self.generation.searching])
+        row["energy_occupying"] = np.mean([n.occupier.energy 
+                                            for n in self.generation.nests
+                                                if n.occupied()])
+        self.data.append(row)
+
+    def get_col(self, col_name):
+        return [ row[col_name] for row in self.data ]
+
+    def plot_cohort(self):
+        self.plot_time_series()
+        self.plot_E_time_series()
+        self.plot_trait_hist()
+        self.plot_trait_scatter()
+
+    def plot_time_series(self):
+        times = self.get_col("time")
+        searching = self.get_col("searching")
+        occupying = self.get_col("occupying")
+        contests = self.get_col("contests")
+        num_matured = self.get_col("num_matured")
+        killed = self.get_col("killed")
+
+        plt.title("various metrics through time. delta = %s" 
+                % self.generation.params["time_step"])
+        plt.xlabel("time steps")
+
+        plt.plot(times, searching, label = "searching males")
+        plt.plot(times, occupying, label = "occupying males")
+        plt.plot(times, contests, label = "total contests")
+        plt.plot(times, num_matured, label = "total matured")
+        plt.plot(times, killed, label = "total_deaths")
+
+        plt.legend(loc = 2)
+        plt.show()
+
+    def plot_E_time_series(self):
+        times = self.get_col("time")
+        mean_energy_searching = self.get_col("energy_searching")
+        mean_energy_occupying = self.get_col("energy_occupying")
+
+        plt.title("Mean cohort energy levels through time. delta = %s" 
+                % self.generation.params["time_step"])
+        plt.xlabel("time steps")
[…]

### 2016-01-14 python/simulation.py

@@ -8,7 +8,7 @@ from generation import Generation
 # the simulation class aggregates generation objects
 class Simulation:
 
-    #constructor
+    # constructor
     def __init__(self, params):
         self.params = params
         self.generations = list()

### 2016-01-14 python/logger.py
added extinction catch

@@ -78,6 +78,8 @@ class Logger(object):
             for n in self.generation.nests 
             if n.occupied()
         ]
+        if len(occupying_males) == 0:
+            print "extintion event"
 
         occupying_exploration_trait = [
             m.exploration 

### 2016-01-18 ms.tex
more plots :D
fixed latex syntax error
fixed major bug with new generations

@@
 $\frac{cS}{1+eO}$ is searching individuals becoming occupiers through finding nests that are unoccuied.\\
-$f$ is the metabolic death rate\.\
+$f$ is the metabolic death rate.\\
 $g$ is the abandon rate.\\
@@
 
-$c_n$ is the stochatic reaction constant
+$c_n$ is the stochatic reaction constant\\
+$c_n \cdot dt = $ the average probability that a particular combination of $R_n$, reactant molecules will react accordingly in the next infinitesimal time interval $dt$.
+\citep{Gillespie-1976}
+
+we generate the reactions then step through them with a population.
 

### 2016-01-18 python/generation.py

@@ -12,7 +12,7 @@ from scipy.stats import uniform
 
 # TODO: 
 # step directly to next event not through time_steps
-# move logs into a new class
+# mature based off mass logistic pdf???
 
 class Generation (object):
 
@@ -24,6 +24,7 @@ class Generation (object):
         self.id = id
         # counters:
         self.contests = 0
+        self.take_overs = 0
         self.num_matured = 0
         self.killed = 0
         self.debug = params["debug"]
@@ -135,6 +136,7 @@ class Generation (object):
                         self.time)
                 nest = self.nests[index]
                 if nest.occupied():
+                    self.contests += 1
                     if self.debug:
                         print "\tnest %s is occupied by %s, contest" % (
                             index, 
@@ -143,7 +145,7 @@ class Generation (object):
                     if loser.id != m.id:
                         self.searching.insert(0,loser)
                         self.searching.remove(m)
-                    self.contests += 1
+                        self.take_overs += 1
                     if self.debug:
                         print "\tnest %s is occupied by %s" % (
                             index, 

### 2016-01-18 python/logger.py

@@ -25,6 +25,7 @@ class Logger(object):
         row["contests"] = self.generation.contests 
         row["num_matured"] = self.generation.num_matured
         row["killed"] = self.generation.killed
+        row["take_overs"] = self.generation.take_overs
 
         row["energy_searching"] = np.mean([m.energy
                                             for m in self.generation.searching])
@@ -49,6 +50,7 @@ class Logger(object):
         contests = self.get_col("contests")
         num_matured = self.get_col("num_matured")
         killed = self.get_col("killed")
+        take_overs = self.get_col("take_overs")
 
         plt.title("gen %s Various generation metrics through time. dt = %s" 
                 % ( self.generation.id, self.generation.params["time_step"]))
@@ -59,6 +61,7 @@ class Logger(object):
         plt.plot(times, contests, label = "total contests")
         plt.plot(times, num_matured, label = "total matured males")
         plt.plot(times, killed, label = "total deaths")
+        plt.plot(times, take_overs, label = "take overs")
 
         plt.legend(loc = 2)
         plt.show()
@@ -69,7 +72,7 @@ class Logger(object):
         mean_energy_occupying = self.get_col("energy_occupying")
 
         plt.title("gen %s: Mean energy levels through time. delta = %s" 
-                % (self.generation.params["time_step"], self.generation.id))
+                % (self.generation.id, self.generation.params["time_step"]))
         plt.xlabel("time steps")
         plt.ylabel("mean Energy values (J)")
         plt.plot(times, mean_energy_searching, label = "searching")
@@ -113,7 +116,7 @@ class Logger(object):
         plt.title("gen %s: trait distribution of winners" % self.generation.id)
         plt.show()
 
-    def plot_trait_scatter(self):
+    def plot_trait_scatter(self, savefig = False):
         occupying_males = [
             n.occupier 
             for n in self.generation.nests 
@@ -129,8 +132,12 @@ class Logger(object):
             for m in occupying_males
         ]
         plt.plot(occupying_exploration_trait, occupying_aggro_trait, 'ro')
-        plt.title("gen %s: trait values of winners" % self.generation.id)
-        plt.xlabel("exploration trait")
+        plt.title("gen %s: trait values of %s winners" %( self.generation.id, len(occupying_exploration_trait)))
+        plt.axis([0,1,0,1])
+        plt.xlabel("exploration prob")
         plt.ylabel("aggression trait")
[…]

### 2016-01-18 python/male.py

@@ -27,7 +27,7 @@ class Male(object):
             abandoning
 
         genetics:
-            maybe make the exploration probs logistic 
+            @maybe make the exploration probs logistic 
             ?????????
             ?????????
             ?????????
@@ -74,7 +74,16 @@ class Male(object):
             if uniform.rvs() < mutation_rate:
                 self.exploration_trait += norm.rvs(0, mutation_sd)
 
-        self.exploration = logistic.cdf(self.exploration_trait)
+            if uniform.rvs() < mutation_rate:
+                self.aggro += norm.rvs(0, mutation_sd)
+
+            if self.aggro < 0:
+                self.aggro = 0
+
+        self.exploration = logistic.cdf(
+            self.exploration_trait / params["exploration_prob_scale"]
+        )
+
             
     
     # mass from 0 to time t

### 2016-01-18 python/nest.py

@@ -45,7 +45,6 @@ class Nest(object):
         if loser.id == defender.id:
             self.eject()
             self.occupy(attacker)
-        #raw_input()
         return loser
 
     # the first display phase, the loser is returned or None

### 2016-01-18 python/simulation.py

@@ -14,36 +14,36 @@ class Simulation(object):
         self.generations = list()
         self.current_gen = 0
 
-    # makes the first generation
-    def start(self):
-        self.current_gen += 1
-        self.generations.append(Generation(self.params))
-
     # creates the next generation from the previous one
     def step(self):
+        if self.current_gen > 0:
+            self.generations.append(Generation(
+                self.params,
+                prev_gen = self.generations[-1],
+                id = self.current_gen))
+            # remove the second last generation from memory so we dont blow out
+            del self.generations[-2]
+        else:
+            self.generations.append(Generation(
+                self.params,
+                id = self.current_gen))
         self.current_gen += 1
-        self.generations.append(Generation(
-            self.params,
-            prev_gen = self.generations[-1],
-            id = self.current_gen))
-        # remove the second last generation from memory soo we dont blow out
-        del self.generations[-2]
+
 
 def main():
     # read in parameters from file
 
     params = json.loads(open("parameters.json").read())
     np.random.seed(seed=params["random_seed"])
-    # create a new simulation object
-
-    sim = Simulation(params)
-    sim.start()
-    if params["initial_plot"]:
-        sim.generations[0].logger.plot_cohort()
     
-    for i in range(1, params["generations"]):
-        sim.step()
+    sim = Simulation(params)
+    for i in range(params["generations"]):
+        print ""
         print "gen", i
+        sim.step()
+        if i % params["save_every"] == 0 and params["save_pngs"]:
+            sim.generations[-1].logger.plot_trait_scatter(True)
+        print "num winners: ",len(sim.generations[-1].winners)
 
     if params["final_plot"]:
         sim.generations[-1].logger.plot_cohort()

### 2016-01-18 ms.tex
added naive event algorithm to ms

@@
 \subsection{Algorithms}
+\subsection{Event Algorithms}
+So far we are considering two algorithms for stepping through events one is a stochastic simulation algorithm and a simpler (although more compuationally expensive) stepping algorithm.
+\subsection{SSA}
 Simulating the population dynamics requires accounting for four types of events:
@@
 
-\subsection{Event Algorithm}
 See Fig. \ref{fig:events}. Population state generates events, events are either:
@@
     \end{tabular}
+
 \end{table}
 
-$c_n$ is the stochatic reaction constant\\
-$c_n \cdot dt = $ the average probability that a particular combination of $R_n$, reactant molecules will react accordingly in the next infinitesimal time interval $dt$.
+$c_n$ is called the stochatic reaction constant\\
+$c_n \cdot dt = $ the average probability that a particular combination of $R_n$, reactant molecules will react accordingly in the next infinitesimal time interval $dt$
+
 \citep{Gillespie-1976}
 
-we generate the reactions then step through them with a population.
+We generate the reaction events then step through them with a population.
+\clearpage
+\subsection{Naive Event Algorithm}
+we have 3 lists
+\begin{verbatim}
+    immature males
+    searching males
+    nests
+\end{verbatim}
+
+every full time unit that the searching male is alive for it has a chance to discover a nest based off of its exploration trait.
+
+\begin{verbatim}
+time = 0
+while time < time when the females mature:
+    check which males have matured
+        add them to the list of searching males
+    for each occupied nest:
+        deduct metabolic costs from the occupier
+        remove dead males
+
+    for each searching male m:
+        deduct metabolic costs from m
+        remove dead males
+        reduce time until next event for m
+        if time until next event <= 0:
+            reset the time until next event
+            if random > exploration_prob:
+                choose a nest at random
+                if it is occupied:
+                    contest the occupying male for control of the nest
+                    if m has won the contest:
+                        m occupy's n
+                        the loser is added to the list of searching males
+                else:
+                    occupy the nest
+    t += delta_time
+\end{verbatim}
 
@@
 See Fig. \ref{fig:contests}.
-For a defender $x_{def}$ vs an attacker $x_{atk}$, there are 3 stages:
+For a defender $x_{def}$ vs an attacker $x_{atk}$, there are 3 (currently 2) stages:
 each individual has an aggression trait $a$ and a mass $m$.
@@
             \item $x_{def}$ wins:
-            $Pr(\rm{def}) = logit{-1}(m_{def} - m_{atk})$
+            $Pr(\rm{def}) = logit^{-1}(m_{def} - m_{atk})$
         \end{itemize}

### 2016-01-18 ms.tex
small ms.tex changes

@@
 \citep{Gillespie-1976}
+I think we would generate the reaction events then step through them with a population.
 
-We generate the reaction events then step through them with a population.
 \clearpage
@@
 
+\clearpage
 \subsection{Contest Algorithm}
@@
 
-In the display phases the probability of escalation is dependent on the visible traits of x and x'
+After the contest Energy costs are deducted from both the winner and the loser.
 
-After the contest Energy costs are deducted from both the winner and the loser.
-The winner also has a chance to abandon the RR (based off a trait).
+The winner also has a chance to abandon the RR (based off a trait).\\
+
 If a death occurs (Energy(x) less than 0) then x is removed from the individual pool.
@@
     \begin{enumerate}
-        \item Size, influenced by time entered maturity
+        \item Size/Mass, influenced by time entered maturity
     \end{enumerate}
@@
     \begin{enumerate}
-        \item Energy level, influenced by size and past actions
         \item Abandon Rate, influences RR abandonment rate
@@
         \hline
-        $m$ & $kg$ & mass\\
-        $m_mat$ & $kg$ & The size at which 50\% of the population is mature.\\ 
+        $m$ & $g$ & mass\\
+        $m_mat$ & $g$ & The size at which 50\% of the population is mature.\\
+        $E$ & kJ & kiloJoules\\
         \hline

### 2016-01-21 python/generation.py
optimised logger a little bit

@@ -31,6 +31,7 @@ class Generation (object):
 
         self.searching = []
 
+
         if not prev_gen: # no genetics
             # create males
             self.immature = [Male(params, self.logger, i) for i in range(params["K"])]
@@ -46,7 +47,8 @@ class Generation (object):
                 print "next generation has %s individuals" % len(self.immature)
 
         # sort males by when they mature
-        self.immature.sort(key = lambda x : x.maturation_time)  
+        self.immature.sort(key = lambda x : x.maturation_time)
+        self.winners = []
         self.run()
 
     def run(self):

### 2016-01-21 python/logger.py

@@ -18,33 +18,36 @@ class Logger(object):
         self.num_matured = 0
         self.occupying = 0
         self.contests = 0
-        self.take_overs = 0
-        
-        # total energy expenditure
+        self.take_overs = 0 
         self.search_energy = 0
         self.contest_energy = 0
         self.occupying_energy = 0
 
-        self.data = []
+        self.data = {
+            "time":[],
+            "searching":[],
+            "killed":[],
+            "num_matured":[],
+            "occupying":[],
+            "contests":[],
+            "take_overs":[],
+            "search_energy":[],
+            "contest_energy":[],
+            "occupying_energy":[]
+        }
 
     # this function should be called each time step
-    # to log the generation to the data list
     def log_cohort(self):
-        row = {}
-        # average energy of searching males
-        row["time"] = self.generation.time
-        row["searching"] = self.searching
-        row["occupying"] = self.occupying
-        row["contests"] = self.contests 
-        row["num_matured"] = self.num_matured
-        row["killed"] = self.killed
-        row["take_overs"] = self.take_overs
-
-        row["search_energy"] = self.search_energy
-        row["contest_energy"] = self.contest_energy
-        row["occupying_energy"] = self.occupying_energy
-
-        self.data.append(row)
+        self.data["time"].append(self.generation.time)
+        self.data["searching"].append(self.searching)
+        self.data["occupying"].append(self.occupying)
+        self.data["contests"].append(self.contests) 
+        self.data["num_matured"].append(self.num_matured)
+        self.data["killed"].append(self.killed)
+        self.data["take_overs"].append(self.take_overs)        
+        self.data["search_energy"].append(self.search_energy)      
+        self.data["contest_energy"].append(self.contest_energy)        
+        self.data["occupying_energy"].append(self.occupying_energy)
 
     def get_col(self, col_name):
         return [ row[col_name] for row in self.data ]
@@ -56,13 +59,13 @@ class Logger(object):
         self.plot_trait_scatter()
 
     def plot_time_series(self):
-        times = self.get_col("time")
-        searching = self.get_col("searching")
-        occupying = self.get_col("occupying")
-        contests = self.get_col("contests")
-        num_matured = self.get_col("num_matured")
-        killed = self.get_col("killed")
-        take_overs = self.get_col("take_overs")
+        times = self.data["time"]
+        searching = self.data["searching"]
+        occupying = self.data["occupying"]
[…]

### 2016-01-21 ms.tex
ms.pdf clean up, removed de's added searching area

@@
 \subsection{Mature phase}
+When a male matures it is given an energy budget linearly proportional to its mass, this is a parameter into the model.\\
 
-/* this is just for consideration I am not sure if it will be needed or useful */\\
-Number of mature males at time $t$:
-\begin{equation} \label{eq:num_mature}
-    K_{mature}(t) = \lfloor{K_{initial}(Pr({\rm mature} \, | m(t)) - S(t))}\rfloor
+\subsection{Searching Males}
+When males mature they begin searching for a females nest
+
+The area that a male $m$ with speed trait $v$ and radius trait $r$ can search in time step $dt$ is
+\begin{equation}
+    A_{searched} = 2r \cdot v \cdot dt
 \end{equation}
-A searching individual has a chance to encounter a nest equal to its ``exploration trait'' $\lambda_i$:
-\begin{equation} \label{eq:encounter probability}
-    Pr({\rm Encounter}) = \lambda_i
+See Fig. \ref{fig:area searched}.
+
+We let the total area enclosing all males and all nests be a parameter, $A_{total}$.
+
+The probability of a male that covers $A_{searched}$ area in time period $dt$ encountering any nest is:
+\begin{equation}
+    P(\rm {discovery}) = \frac {N \cdot A_{searched}} {A_{total}}
 \end{equation}
-The number of encounters that a searching individual has engaged in is:
-\begin{equation} \label{eq:encounter rate}
-    {\rm Encounters} = \lfloor\lambda_i \cdot t_{searching}\rfloor
-\end{equation}
-the average exploration trait across a cohort is $\bar\lambda$.
+\citep{Gillespie-1976}
+\\
+TODO metabolic cost of searching with r and v ?\\
 
-Similarly for males abandoning nests males the probability of abandoning a nest $j$, is a function of the  the reproductive value of the nest $R_j$ and the abandon trait of the individual $\psi_i$:
+\subsection{Occupying males}
+The probability of a male abandoning a nest $j$, is a function of the  the reproductive value of the nest $R_j$ and the abandon trait of the individual $\psi_i$:
 
@@
 \end{equation}
-/* end */
-
-Searching individuals:
-\begin{equation} \label{eq:searching}
-    \frac{dS}{dt} = aS - bS^2 - \frac{cS}{1+eO} - fS - hO + gO
-\end{equation}
-Occupying individuals:
-\begin{equation} \label{eq:occupying}
-    \frac{dO}{dt} = \frac{cS}{1+eO} - fO - gO
-\end{equation}
-
-$aS - bS^2$ is individuals becoming mature.\\
-$\frac{cS}{1+eO}$ is searching individuals becoming occupiers through finding nests that are unoccuied.\\
-$f$ is the metabolic death rate.\\
-$g$ is the abandon rate.\\
-$h$ is the rate at which occupiers die in a contest.\\
-
-Need equations for:
-\begin{itemize}
-    \item Populations of ``searching'' males and  ``occupying'' males
-    \item Rate at which ``searching'' males encounter empty nests
-    \item Rate at which ``searching'' males encounter occupied nests
-    \item Rate at which ``occupying'' males abandon an occupied nest
-    \item Outcomes of contest between two males
-\end{itemize}
 
@@
 
-What types of mutations?
+A new individual will inherit the fathers genes,
+There is a global mutation rate parameter which is the chance that a trait will not be inherited perfectly but instead will be mutated.
+This means adding some normally distributed noise with $\mu = 0$ and $\sigma$ varying from trait to trait.
+
+\begin{enumerate}
+    \item Abandon Rate, influences RR abandonment rate (not yet implemented)
+    Trade offs:
+    \begin{itemize}
+        \item more likely to abandon a nest and not discover another before the females mature
+    \end{itemize}
+
+    \item Aggressiveness, influences escalation in contests
+    Trade offs:
+    \begin{itemize}
+        \item more likely to engage in fights with a low probability of winning
+    \end{itemize}
+
+    \item search radius, the radius of a searching male (can be thought of as line of sight), See Fig. \ref{fig:area searched}.
+    Trade offs:
+    \begin{itemize}
+        \item higher metabolic cost of search
+    \end{itemize}
+
+    \item search speed, the distance that a searching male will cover in one time step, see Fig. \ref{fig:area searched}.
+    Trade offs:
+        \begin{itemize}
+            \item higher metabolic cost of search
+            \item higher predation rate
+        \end{itemize}
+\end{enumerate}
 
@@
 Simulating the population dynamics requires accounting for four types of events:
-\begin{itemize}
-    \item Immature individuals become mature
-    \item Searching males encounter an empty nest
-    \item Searching males encounter an occupied nest and contest it
-    \item Occupying males abandon an occupied nest
-\end{itemize}
 
@@
 
-See Fig. \ref{fig:events}. Population state generates events, events are either:
-\begin{enumerate}
-    \item A new individual x has matured and entered the pool (based off of a trait(?))
-    \item Discovery event wherin both a nest RR (repoductive resource) and an individual x are drawn from a pool, the metabolic costs of searching are deducted from the time differences between events, that nest is either:
-    \begin{enumerate}
-        \item un-occupied: in which x takes ownership of RR.
-        \item occupied by x' in which a constest occurs (see fig 2)
-    \end{enumerate}
-\end{enumerate}
-
-In the SSA algorithm \citep{Gillespie-1976} there are 6 different \"species\", representing different states:
+In the SSA algorithm \citep{Gillespie-1976} there are 6 different states
 \begin{enumerate}
@@
 
-these species have $R_\mu$ different interactions
+these states have $R_\mu$ different interactions
 \begin{table}[h!]
@@
         \hline
-        Reaction num & reactants  & & products & Description\\
+            Event Number & reactants  & & products & Description\\
         \hline
@@
             $R_6$       &   $S_4$       & $--c_6-->$      &   $S_2 + S_3$ & abandon event\\
-            $R_7$       &   $S_1$       & $--c_7-->$      &   $0$         & metab0lic / predation event\\
+            $R_7$       &   $S_1$       & $--c_7-->$      &   $0$         & metabolic / predation event\\
             $R_8$       &   $S_2$       & $--c_8-->$      &   $0$         & predation event\\
-            $R_9$       &   $S_4$       & $--c_9-->$      &   $S_3$       & metab0lic / predation event\\
+            $R_9$       &   $S_4$       & $--c_9-->$      &   $S_3$       & metabolic / predation event\\
             $R_{10}$    &   $S_6$       & $--c_{10}-->$   &   $S_4$       & nest handover\\
@@
 
-$c_n$ is called the stochatic reaction constant\\
-$c_n \cdot dt = $ the average probability that a particular combination of $R_n$, reactant molecules will react accordingly in the next infinitesimal time interval $dt$
+$c_n$ is called the stochastic reaction constant\\
+$c_n \cdot dt = $ the average probability that a particular combination of $R_n$, reactant molecules will react accordingly in the next infinitesimal time interval $dt$\\
 
 \citep{Gillespie-1976}
-I think we would generate the reaction events then step through them with a population.
 
@@
 \end{verbatim}
-
-every full time unit that the searching male is alive for it has a chance to discover a nest based off of its exploration trait.
 
@@
     \item Fight
-        So far I have something like
         \begin{itemize}
@@
 
-Traits:
-\begin{enumerate}
-    \item Visible to opponent:
-    \begin{enumerate}
-        \item Size/Mass, influenced by time entered maturity
-    \end{enumerate}
-    \item invisible to opponent:
-    \begin{enumerate}
-        \item Abandon Rate, influences RR abandonment rate
-        \item Aggressiveness, influences escalation in contests
-        \item Encounter Rate, influences chance of being selected for an event
-    \end{enumerate}
-\end{enumerate}
 \clearpage
 
-I am not sure if this is the best way to do this, the tables in plant are more aesthetic. 
+I am not sure if this is the best way to do this, the tables in the plant paper are more aesthetic.\\
+
 \begin{table}[h!]
@@
 \centering
-\includegraphics[width=15cm,height=15cm,keepaspectratio]{figures/event_algorithm}
+\includegraphics[width=10cm,height=10cm,keepaspectratio]{figures/event_algorithm}
 \caption{Algorithm for modelling dynamics within the ``mature'' male population.}
@@
 
+\begin{figure}[h!]
+\centering
+\includegraphics[width=15cm,height=15cm,keepaspectratio]{figures/area_searched}
+\caption{the area searched my a male in interval $dt$}
+\label{fig:area searched}
+\end{figure}
+\clearpage
+
 \bibliography{refs}

### 2016-01-21 nest.py

@@ -1,10 +1,0 @@
-#!/usr/bin/python
-
-""" a nest has a RR and female genetics and a male occupier"""
-class Nest(object):
-
-    def __init__(self, params):
-        self.rr = 0
-        self.female = []
-        self.male = []
-        self.occupied = False

### 2016-01-21 python/generation.py

@@ -71,37 +71,12 @@ class Generation (object):
         # until the females mature: 
         while (self.time < f_time):
             # check the next mature male
-            maturing = filter(lambda m: m.maturation_time <= self.time, self.immature)
-            for m in maturing:
-                self.searching.append(m)
-                self.immature.remove(m)
-                if self.debug:
-                    print "male matured at %s" % self.time
-                self.logger.inc_num_matured()
-            
-            
-            # iterate over nests subtracting metabolic costs from occupying
-            # males
-            for n in self.nests:
-                if n.occupied():
-                    n.occupier.occupy(dt)
-                    if not n.occupier.is_alive(): # remove the dead males
-                        m = n.eject()
-                        self.logger.inc_killed()
-                        self.logger.dec_occupying()
-
-                        if self.debug:
-                            print "male %s in nest %s has died at t = %s" % (
-                                m.id,
-                                n.id,
-                                self.time)
-
-
+            self.maturation_step()
+            # iterate over occupying males
+            self.occupation_step(dt)
             # iterate over searching males
             self.searching_step(dt)
-
             self.logger.log_cohort()
-
             self.time += dt
         
         if self.params["generation_plot"]:
@@ -161,3 +136,29 @@ class Generation (object):
                             index, 
                             m.id)
 
+    # allows males to mature
+    def maturation_step(self):
+        maturing = filter(lambda m: m.maturation_time <= self.time, self.immature)
+        for m in maturing:
+            self.searching.append(m)
+            self.immature.remove(m)
+            if self.debug:
+                print "male matured at %s" % self.time
+            self.logger.inc_num_matured()
+
+    # iterate over nests subtracting metabolic costs from occupying
+    # males
+    def occupation_step(self, dt):
+        for n in self.nests:
+            if n.occupied():
+                n.occupier.occupy(dt)
+                if not n.occupier.is_alive(): # remove the dead males
+                    m = n.eject()
+                    self.logger.inc_killed()
+                    self.logger.dec_occupying()
+
+                    if self.debug:
+                        print "male %s in nest %s has died at t = %s" % (
+                            m.id,
+                            n.id,
+                            self.time)

### 2016-01-21 python/logger.py

@@ -67,6 +67,10 @@ class Logger(object):
         killed = self.data["killed"]
         take_overs = self.data["take_overs"]
 
+        occupying_males = self.get_occupying_males()
+
+        winners_matured = np.mean([m.maturation_time for m in occupying_males])
+                
         plt.title("Gen %s Various generation metrics through time. dt = %s" 
                 % ( self.generation.id, self.generation.params["time_step"]))
         plt.xlabel("time steps")
@@ -77,6 +81,8 @@ class Logger(object):
         plt.plot(times, num_matured, label = "total matured males")
         plt.plot(times, killed, label = "total deaths")
         plt.plot(times, take_overs, label = "take overs")
+        plt.axvline(winners_matured, color='k', linestyle='--')
+        plt.text(winners_matured + 0.1,1400,'average winner matured',rotation=90)
 
         plt.legend(loc = 2)
         plt.show()
@@ -99,10 +105,12 @@ class Logger(object):
             occupying_energy,
             contest_energy)
 
+        # plot the legend
         p1 = plt.Rectangle((0, 0), 1, 1, fc="red")
         p2 = plt.Rectangle((0, 0), 1, 1, fc="green")
         p3 = plt.Rectangle((0, 0), 1, 1, fc="blue")
         plt.legend([p1, p2, p3], ["search", "occupying", "contest"])
+
         if savefig:
             plt.savefig("plots/gen_%03d_total_energy_expenditure.png" % self.generation.id)
             plt.close()
@@ -110,11 +118,7 @@ class Logger(object):
         plt.show()
 
     def plot_trait_hist(self):
-        occupying_males = [
-            n.occupier 
-            for n in self.generation.nests 
-            if n.occupied()
-        ]
+        occupying_males = self.get_occupying_males()
         if len(occupying_males) == 0:
             print "extintion event"
 
@@ -146,13 +150,7 @@ class Logger(object):
         plt.show()
 
     def plot_trait_scatter(self, savefig = False):
-        occupying_males = [
-            n.occupier 
-            for n in self.generation.nests 
-            if n.occupied()
-        ]
-
-
+        occupying_males = self.get_occupying_males()
         occupying_exploration_trait = [
             m.exploration
             for m in occupying_males
@@ -204,3 +202,10 @@ class Logger(object):
     
     def inc_occupying_energy(self, j):
         self.occupying_energy += j
+
+    def get_occupying_males(self):
+        return [
+            n.occupier 
+            for n in self.generation.nests 
+            if n.occupied()
+        ]

### 2016-01-21 python/generation.py
generation logger and simulation logger schism

@@ -1,13 +1,14 @@
 #!/usr/bin/env python
 from male import Male
 from nest import Nest
-from logger import Logger
+from generation_logger import GenerationLogger
 
 from scipy.stats import uniform
+import numpy as np
 
 # a generation in the minimal model
 # aggregates males
-# aggregates nests
+# aggregates nest
 # steps them through 1 generation of the simulation
 
 # TODO: 
@@ -24,7 +25,7 @@ class Generation (object):
         self.id = id
         # counters:
         self.debug = params["debug"]
-        self.logger = Logger(self)
+        self.logger = GenerationLogger(self)
         
         # make nests
         self.nests = [Nest(params, i) for i in range(params["N"])]
@@ -48,8 +49,11 @@ class Generation (object):
 
         # sort males by when they mature
         self.immature.sort(key = lambda x : x.maturation_time)
-        self.winners = []
         self.run()
+        if self.params["generation_plot"]:
+            self.logger.plot_cohort()
+
+        self.winners = [n for n in self.nests if n.occupied()]
 
     def run(self):
         dt = self.params["time_step"]
@@ -78,11 +82,7 @@ class Generation (object):
             self.searching_step(dt)
             self.logger.log_cohort()
             self.time += dt
-        
-        if self.params["generation_plot"]:
-            self.logger.plot_cohort()
 
-        self.winners = [n for n in self.nests if n.occupied()]
 
     # itereates over all searching males
     def searching_step(self, dt):

### 2016-01-21 python/generation_logger.py

@@ -1,14 +1,13 @@
-# a class to hold data logs of generations:
+# a class to hold data logs of a generation:
 
 import numpy as np
 import sys
 import matplotlib.pyplot as plt
 
-# TODO make 1 logger per simulation which aggregates generation data
 # TODO rename to generation logger
-# TODO make energy time series used a %
+# TODO make energy time series use a %
 
-class Logger(object):
+class GenerationLogger(object):
     """docstring for logs"""
     def __init__(self, generation):
         self.generation = generation
@@ -81,6 +80,7 @@ class Logger(object):
         plt.plot(times, num_matured, label = "total matured males")
         plt.plot(times, killed, label = "total deaths")
         plt.plot(times, take_overs, label = "take overs")
+        
         plt.axvline(winners_matured, color='k', linestyle='--')
         plt.text(winners_matured + 0.1,1400,'average winner matured',rotation=90)
 

### 2016-01-21 python/male.py

@@ -5,35 +5,6 @@ from scipy.stats import norm, logistic, uniform
 
 # a male individual in the population
 class Male(object):
-    #TODO
-    """
-        qualities:
-            @maturation time:
-            @energy at maturation
-            @mass at maturation
-        
-        heriditable traits:
-            @exploration 
-            psi / abandon
-            @aggression
-
-        costs:
-            @metabolic
-            predation
-
-        events:
-            @contests
-            @searching
-            abandoning
-
-        genetics:
-            @maybe make the exploration probs logistic 
-            ?????????
-            ?????????
-            ?????????
-    """
-
-
     # constructor
     def __init__(self, params, logger, id, mom = None, dad = None):
         self.id = id

### 2016-01-21 python/simulation.py

@@ -3,7 +3,7 @@
 import json
 import numpy as np
 from generation import Generation
-
+from simulation_logger import SimulationLogger
 
 # the simulation class aggregates generation objects
 class Simulation(object):
@@ -11,7 +11,8 @@ class Simulation(object):
     # constructor
     def __init__(self, params):
         self.params = params
-        self.generations = list()
+        self.generations = []
+        self.sim_logger = SimulationLogger(self, params)
         self.current_gen = 0
 
     # creates the next generation from the previous one
@@ -21,12 +22,17 @@ class Simulation(object):
                 self.params,
                 prev_gen = self.generations[-1],
                 id = self.current_gen))
-            # remove the second last generation from memory so we dont blow out
+
+            # remove the second last generation from memory 
+            # so we dont blow out
             del self.generations[-2]
         else:
             self.generations.append(Generation(
                 self.params,
                 id = self.current_gen))
+        
+        self.sim_logger.log(self.generations[-1])
+        
         self.current_gen += 1
 
 
@@ -48,6 +54,8 @@ def main():
             sim.generations[-1].logger.plot_e_time_series(True)
 
         print "num winners: ",len(sim.generations[-1].winners)
+
+    sim.sim_logger.plot()
 
     if params["final_plot"]:
         sim.generations[-1].logger.plot_cohort()

### 2016-01-21 python/simulation_logger.py

@@ -1,0 +1,53 @@
+import numpy as np
+import matplotlib.pyplot as plt
+"""
+    This class will hold summary information about many generations in a simulation
+"""
+class SimulationLogger (object):    
+    def __init__(self, simulation, params):
+        self.simulation = simulation
+        self.params = params
+        self.data = {
+            "generation": [],
+            "mean_exploration": [],
+            "std_exploration": [],
+            "mean_aggro": [],
+            "std_aggro": [],
+            "num_winners": []
+        }
+
+    def log(self, gen):
+        winning_males = [ n.occupier for n in gen.winners ]
+
+        exploration_traits = map(lambda x: x.exploration_trait, winning_males)
+
+        aggro_traits = map(lambda x: x.aggro, winning_males)
+
+        self.data["generation"].append(gen.id)
+        self.data["mean_exploration"].append(np.mean(exploration_traits))
+        self.data["std_exploration"].append(np.std(exploration_traits))
+        self.data["mean_aggro"].append(np.mean(aggro_traits))
+        self.data["std_aggro"].append(np.std(aggro_traits))
+        self.data["num_winners"].append(len(winning_males))
+
+    # TODO
+    def print_csv(self):
+        pass
+
+    # TODO I need some way to match the paramters with a plot
+    def plot(self):
+        gens = self.data["generation"]
+        
+        plt.plot(gens, self.data["mean_exploration"], label = "mean exploration")
+        plt.plot(gens, self.data["std_exploration"], label = "std exploration")
+        plt.plot(gens, self.data["mean_aggro"], label = "mean aggro")
+        plt.plot(gens, self.data["std_aggro"], label = "std aggro")
+
+        plt.legend(loc = 2)
+        plt.title("Trait values through Generations")
+        
+        plt.xlabel("generation")
+        plt.ylabel("value")
+
+        plt.show()
+

### 2016-01-21 python/generation_logger.py
boxplots and searching energy cost formula

@@ -80,7 +80,7 @@ class GenerationLogger(object):
         plt.plot(times, num_matured, label = "total matured males")
         plt.plot(times, killed, label = "total deaths")
         plt.plot(times, take_overs, label = "take overs")
-        
+
         plt.axvline(winners_matured, color='k', linestyle='--')
         plt.text(winners_matured + 0.1,1400,'average winner matured',rotation=90)
 
@@ -123,7 +123,7 @@ class GenerationLogger(object):
             print "extintion event"
 
         occupying_exploration_trait = [
-            m.exploration 
+            m.exploration
             for m in occupying_males
         ]
         occupying_aggro_trait = [
@@ -133,7 +133,7 @@ class GenerationLogger(object):
 
         bins = self.generation.params["trait_bins"]
         hist_range = (0,1)
-
+        """
         plt.hist(occupying_exploration_trait,
             range = hist_range,
             bins = bins,
@@ -145,6 +145,10 @@ class GenerationLogger(object):
             bins = bins,
             alpha = 0.3,
             label = "aggression")
+        """
+        plt.boxplot([occupying_aggro_trait, occupying_exploration_trait])
+        labels = ('aggression', 'exploration')
+        plt.xticks([1,2],labels)
         plt.legend()
         plt.title("gen %s: trait distribution of winners" % self.generation.id)
         plt.show()

### 2016-01-21 ms.tex
boxplots and searching energy cost formula
new search in the model

@@
 \citep{Gillespie-1976}
-\\
-TODO metabolic cost of searching with r and v ?\\
+The energy used by a male with speed $v$ and radius $r$ is 
+\begin{equation}
+    E(r,v) = k_{search} \cdot r \cdot v
+\end{equation}
+where $k_{search}$ is a constant.
 
@@
 \subsection{Event Algorithms}
-So far we are considering two algorithms for stepping through events one is a stochastic simulation algorithm and a simpler (although more compuationally expensive) stepping algorithm.
+So far we are considering two algorithms for stepping through events one is a stochastic simulation algorithm and a simpler (although more computationally expensive) naive stepping algorithm.
 \subsection{SSA}

### 2016-01-21 python/generation_logger.py

@@ -98,7 +98,6 @@ class GenerationLogger(object):
                 % (self.generation.id, self.generation.params["time_step"]))
         plt.xlabel("Time")
         plt.ylabel("Energy expenditure (J)")
-        plt.axis([times[0],times[-1],0,self.generation.params["energy_y_max"]])
 
         plt.stackplot(times,
             search_energy,
@@ -127,7 +126,7 @@ class GenerationLogger(object):
             for m in occupying_males
         ]
         occupying_aggro_trait = [
-            m.aggro / self.generation.params["aggression_max"]
+            m.aggro
             for m in occupying_males
         ]
 
@@ -160,12 +159,11 @@ class GenerationLogger(object):
             for m in occupying_males
         ]
         occupying_aggro_trait = [
-            m.aggro / self.generation.params["aggression_max"]
+            m.aggro
             for m in occupying_males
         ]
         plt.plot(occupying_exploration_trait, occupying_aggro_trait, 'ro',)
         plt.title("gen %s: trait values of %s winners" %( self.generation.id, len(occupying_exploration_trait)))
-        plt.axis([0,1,0,1])
         plt.xlabel("exploration prob")
         plt.ylabel("aggression trait")
         if savefig:

### 2016-01-21 python/male.py

@@ -11,12 +11,6 @@ class Male(object):
 
         self.logger = logger
 
-        # countdown timer to next event
-        self.tt_event = 1.0
-        
-        self.metabolic_cost_search = params["metabolic_cost_search"]
-        self.metabolic_cost_occupy = params["metabolic_cost_occupy"]
-
         self.occupying = False
         
         # pull the maturation time from the inverse logit
@@ -32,31 +26,40 @@ class Male(object):
 
         # trait values
         if not mom and not dad: # first gen: no breeding
-            self.exploration_trait = norm.rvs(params["exploration_mean"], params["exploration_sd"])
-
+            self.radius = abs(norm.rvs(params["radius_mean"], params["radius_sd"]))
+            self.speed = abs(norm.rvs(params["speed_mean"], params["speed_sd"]))
             # pull the aggression from the normal distribution
             self.aggro = params["aggression_max"] * uniform.rvs()
 
-        elif not mom: # asexual genetics
+        elif not mom: # genetic inheritance from the dads
             mutation_rate = params["mutation_rate"]
             mutation_sd = params["mutation_sd"]
             
-            self.exploration_trait = dad.exploration_trait
+            self.radius = dad.radius
+            self.speed = dad.speed
             self.aggro = dad.aggro
             
             if uniform.rvs() < mutation_rate:
-                self.exploration_trait += norm.rvs(0, mutation_sd)
+                self.radius += norm.rvs(0, mutation_sd)
+
+            if uniform.rvs() < mutation_rate:
+                self.speed += norm.rvs(0, mutation_sd)
 
             if uniform.rvs() < mutation_rate:
                 self.aggro += norm.rvs(0, mutation_sd)
 
+            self.aggro = 0 if self.aggro < 0 else self.aggro
+            self.speed = 0 if self.speed < 0 else self.speed
+            self.radius = 0 if self.radius < 0 else self.radius
+
             if self.aggro < 0:
                 self.aggro = 0
 
-        self.exploration = logistic.cdf(
-            self.exploration_trait / params["exploration_prob_scale"]
-        )
+        self.exploration = 2 * self.radius * self.speed
+        self.exploration_prob = (params["N"] * self.exploration * params["time_step"]) / params["patch_area"]
 
+        self.metabolic_cost_search = params["search_energy_coef"] * self.radius * self.speed
+        self.metabolic_cost_occupy = params["metabolic_cost_occupy"]
             
     
     # mass from 0 to time t
[…]

### 2016-01-21 python/simulation_logger.py

@@ -19,7 +19,7 @@ class SimulationLogger (object):
     def log(self, gen):
         winning_males = [ n.occupier for n in gen.winners ]
 
-        exploration_traits = map(lambda x: x.exploration_trait, winning_males)
+        exploration_traits = map(lambda x: x.exploration, winning_males)
 
         aggro_traits = map(lambda x: x.aggro, winning_males)
 

### 2016-01-22 python/generation.py
maturation trait

@@ -95,6 +95,7 @@ class Generation (object):
                     print "male %s has died at %s" % (m.id, self.time)
                 self.searching.remove(m)
                 self.logger.inc_killed()
+                self.logger.dec_searching()
             elif discovered:
                 # select a nest at random
                 index = int(uniform.rvs() * len(self.nests))
@@ -130,6 +131,7 @@ class Generation (object):
                     nest.occupy(m)
                     self.searching.remove(m)
                     self.logger.inc_occupying()
+                    self.logger.dec_searching()
 
                     if self.debug:
                         print "\tnest %s is now occupied by %s" % (
@@ -145,6 +147,7 @@ class Generation (object):
             if self.debug:
                 print "male matured at %s" % self.time
             self.logger.inc_num_matured()
+            self.logger.inc_searching()
 
     # iterate over nests subtracting metabolic costs from occupying
     # males
@@ -156,7 +159,6 @@ class Generation (object):
                     m = n.eject()
                     self.logger.inc_killed()
                     self.logger.dec_occupying()
-
                     if self.debug:
                         print "male %s in nest %s has died at t = %s" % (
                             m.id,

### 2016-01-22 python/generation_logger.py

@@ -57,7 +57,7 @@ class GenerationLogger(object):
         self.plot_trait_hist()
         self.plot_trait_scatter()
 
-    def plot_time_series(self):
+    def plot_time_series(self, savefig = False):
         times = self.data["time"]
         searching = self.data["searching"]
         occupying = self.data["occupying"]
@@ -85,6 +85,10 @@ class GenerationLogger(object):
         plt.text(winners_matured + 0.1,1400,'average winner matured',rotation=90)
 
         plt.legend(loc = 2)
+        if savefig:
+            plt.savefig("plots/gen_%03d_time_series.png" % self.generation.id)
+            plt.close()
+            return
         plt.show()
 
     def plot_e_time_series(self, savefig = False):
@@ -162,10 +166,13 @@ class GenerationLogger(object):
             m.aggro
             for m in occupying_males
         ]
+
         plt.plot(occupying_exploration_trait, occupying_aggro_trait, 'ro',)
+        plt.axis([0,40,0,10])
         plt.title("gen %s: trait values of %s winners" %( self.generation.id, len(occupying_exploration_trait)))
-        plt.xlabel("exploration prob")
-        plt.ylabel("aggression trait")
+        
+        plt.xlabel("2 * r * v")
+        plt.ylabel("aggression")
         if savefig:
             plt.savefig("plots/gen_%03d_winner_trait_values.png" % self.generation.id)
             plt.close()

### 2016-01-22 python/male.py

@@ -13,16 +13,8 @@ class Male(object):
 
         self.occupying = False
         
-        # pull the maturation time from the inverse logit
-        self.maturation_time = abs(logistic.rvs(
-            params["maturation_center"],
-            params["maturation_width"]))
-
-        #get mass at maturation:
-        self.grow(params)
-
-        # energy at maturation
-        self.energy = self.mass * params["mass_to_energy"]
+        # is maturation time an inheritable trait
+        inherit_time = params["inherit_maturation_time"]
 
         # trait values
         if not mom and not dad: # first gen: no breeding
@@ -30,7 +22,9 @@ class Male(object):
             self.speed = abs(norm.rvs(params["speed_mean"], params["speed_sd"]))
             # pull the aggression from the normal distribution
             self.aggro = params["aggression_max"] * uniform.rvs()
-
+            self.maturation_time = abs(logistic.rvs(
+                params["maturation_center"],
+                params["maturation_width"]))
         elif not mom: # genetic inheritance from the dads
             mutation_rate = params["mutation_rate"]
             mutation_sd = params["mutation_sd"]
@@ -38,7 +32,16 @@ class Male(object):
             self.radius = dad.radius
             self.speed = dad.speed
             self.aggro = dad.aggro
-            
+
+            if inherit_time:
+                self.maturation_time = dad.maturation_time
+                self.maturation_time += norm.rvs(0, params["maturation_noise"])
+            else:
+                self.maturation_time = abs(logistic.rvs(
+                    params["maturation_center"],
+                    params["maturation_width"]))
+
+            # mutations
             if uniform.rvs() < mutation_rate:
                 self.radius += norm.rvs(0, mutation_sd)
 
@@ -51,9 +54,13 @@ class Male(object):
             self.aggro = 0 if self.aggro < 0 else self.aggro
             self.speed = 0 if self.speed < 0 else self.speed
             self.radius = 0 if self.radius < 0 else self.radius
+            self.maturation_time = 0 if self.maturation_time < 0 else self.maturation_time
 
-            if self.aggro < 0:
-                self.aggro = 0
+        #get mass at maturation:
+        self.grow(params)
+
+        # energy at maturation
+        self.energy = self.mass * params["mass_to_energy"]
 
         self.exploration = 2 * self.radius * self.speed
         self.exploration_prob = (params["N"] * self.exploration * params["time_step"]) / params["patch_area"]

### 2016-01-22 python/simulation.py

@@ -47,17 +47,18 @@ def main():
         print ""
         print "gen", i
         sim.step()
-        if i % params["save_every"] == 0 and params["save_scatter_pngs"]:
-            sim.generations[-1].logger.plot_trait_scatter(True)
-        
-        if i % params["save_every"] == 0 and params["save_energy_pngs"]:
-            sim.generations[-1].logger.plot_e_time_series(True)
+        if i % params["save_every"] == 0:
+            if params["save_scatter_pngs"]:
+                sim.generations[-1].logger.plot_trait_scatter(True)
+            if params["save_energy_pngs"]:
+                sim.generations[-1].logger.plot_e_time_series(True)
+            if params["save_time_series_pngs"]:
+                sim.generations[-1].logger.plot_time_series(True)
 
         print "num winners: ",len(sim.generations[-1].winners)
 
-    sim.sim_logger.plot()
-
     if params["final_plot"]:
+        sim.sim_logger.plot(params["save_sim_trait_pngs"])
         sim.generations[-1].logger.plot_cohort()
 
 if __name__ == '__main__':

### 2016-01-22 python/simulation_logger.py

@@ -13,6 +13,8 @@ class SimulationLogger (object):
             "std_exploration": [],
             "mean_aggro": [],
             "std_aggro": [],
+            "mean_maturation":[],
+            "std_maturation":[],
             "num_winners": []
         }
 
@@ -23,11 +25,15 @@ class SimulationLogger (object):
 
         aggro_traits = map(lambda x: x.aggro, winning_males)
 
+        maturations = map(lambda x: x.maturation_time, winning_males)
+
         self.data["generation"].append(gen.id)
         self.data["mean_exploration"].append(np.mean(exploration_traits))
         self.data["std_exploration"].append(np.std(exploration_traits))
         self.data["mean_aggro"].append(np.mean(aggro_traits))
         self.data["std_aggro"].append(np.std(aggro_traits))
+        self.data["mean_maturation"].append(np.mean(maturations))
+        self.data["std_maturation"].append(np.std(maturations))
         self.data["num_winners"].append(len(winning_males))
 
     # TODO
@@ -35,7 +41,7 @@ class SimulationLogger (object):
         pass
 
     # TODO I need some way to match the paramters with a plot
-    def plot(self):
+    def plot(self, savefig = False):
         gens = self.data["generation"]
         
         plt.plot(gens, self.data["mean_exploration"], label = "mean exploration")
@@ -43,11 +49,18 @@ class SimulationLogger (object):
         plt.plot(gens, self.data["mean_aggro"], label = "mean aggro")
         plt.plot(gens, self.data["std_aggro"], label = "std aggro")
 
+        plt.plot(gens, self.data["mean_maturation"], label = "mean maturationtime")
+        plt.plot(gens, self.data["std_maturation"], label = "std maturation_time")
+
         plt.legend(loc = 2)
         plt.title("Trait values through Generations")
         
         plt.xlabel("generation")
         plt.ylabel("value")
 
+        if savefig:
+            plt.savefig("plots/gen_simulation_trait_values.png" % self.generation.id)
+            plt.close()
+            return
         plt.show()
 

### 2016-01-23 ms.tex
maturation trait
some runs

@@
         \end{itemize}
+    \item maturation time, the time at which to mature is inherited as well but with a constant level of noise
 \end{enumerate}
@@
 $c_n \cdot dt = $ the average probability that a particular combination of $R_n$, reactant molecules will react accordingly in the next infinitesimal time interval $dt$\\
+s
+\begin{verbatim}
+I have the probability of encountering a   
+\end{verbatim}
 
@@
 \includegraphics[width=15cm,height=15cm,keepaspectratio]{figures/area_searched}
-\caption{the area searched my a male in interval $dt$}
+\caption{the area searched by a male in interval $dt$}
 \label{fig:area searched}

### 2016-01-23 R/load_datasets.R

@@ -1,5 +1,5 @@
 library(jsonlite)
-library(plyr)
+library(ggplot2)
 
 # I am guessing that you are in the top dir
 
@@ -13,17 +13,20 @@ df <- lapply(
   }
 )
 
-# select these columns for analysis
 master <- data.frame()
 
-# this works but feels pretty gross and not R like
-# will please teach me the R way
+# this works but feels pretty gross
+ 
+# Will please teach me the R way
+
 
 for (simulation in df){
     simulation <- data.frame(
+      # select these columns for analysis
       simulation$traits.aggression,
       simulation$parameters.patch_area
     )
+    # rename them
     colnames(simulation) <- c(
       "aggression",
       "patch_area"
@@ -33,5 +36,10 @@ for (simulation in df){
 
 setwd("../..")
 
+ggplot(master,aes(x = patch_area, y = aggression, colour = aggression)) + geom_point() + scale_color_gradient(low="blue",high="red")
+#ggplot(master,aes(x = patch_area, y = aggression)) + geom_density2d()
+
+#qplot(y = master$aggression, x = master$patch_area )
+.
 
 

### 2016-01-23 python/simulation.py

@@ -45,7 +45,8 @@ def main():
     sim = Simulation(params)
     for i in range(params["generations"]):
         print ""
-        print "gen", i
+        print "gen =\t\t", i
+        print "patch area = \t", params["patch_area"]
         sim.step()
         if i % params["save_every"] == 0:
             if params["save_scatter_pngs"]:
@@ -55,7 +56,7 @@ def main():
             if params["save_time_series_pngs"]:
                 sim.generations[-1].logger.plot_time_series(True)
 
-        print "num winners: ",len(sim.generations[-1].winners)
+        print "num winners =\t",len(sim.generations[-1].winners)
 
     if params["final_plot"]:
         sim.logger.plot(params["save_sim_trait_pngs"])

### 2016-01-25 ms.tex
fixed remake issue in importer, added some event hoppping theory stuff

@@
 
-The area that a male $m$ with speed trait $v$ and radius trait $r$ can search in time step $dt$ is
+The area that a male $m$ with speed trait $v$ and radius trait $r$ can search in one time step is
 \begin{equation}
-    A_{searched} = 2r \cdot v \cdot dt
+    2r \cdot v
 \end{equation}
@@
 
-The probability of a male that covers $A_{searched}$ area in time period $dt$ encountering any nest is:
+The probability of a male that covers $A_{searched}$ in one time step encountering any nest is:
 \begin{equation}
-    P(\rm {discovery}) = \frac {N \cdot A_{searched}} {A_{total}}
+    P(\rm {discovery}) = \frac {N \cdot 2r\cdot v} {A_{total}}
 \end{equation}
+Let $t_{mat}$ be the time when a male matures.
+Now the maximum time possible time that a male can spend searching is $t_{max} = t_{females mature} - t_{mat}$
+
+The expected number of events is $$ P( \rm {discovery}) \cdot t_{max}$$.
+We then use this as the $\lambda$ parameter to draw the number of nest discovery events for this individual $n$ from the poisson distribution.
+
+We then pull $n$ numbers from the uniform distrobution with lower bound $t_{mat}$ and upper bound $t_{females mature}$ as the times of these discovery events.
+
+%I think I might use a priority queue for this...
+
 \citep{Gillespie-1976}

### 2016-01-25 R/load_datasets.R

@@ -1,26 +1,19 @@
 library(jsonlite)
 library(ggplot2)
 
-# I am guessing that you are in the top dir
-
-setwd("python/data")
-df <- lapply(
-  dir(), 
-  function (filename){ 
-    as.data.frame(
-      fromJSON(filename)
-    )
-  }
-)
-
-master <- data.frame()
-
-# this works but feels pretty gross
- 
-# Will please teach me the R way
-
-
-for (simulation in df){
+load_dataset  <- function (){
+  df <- lapply(
+    dir(), 
+    function (filename){ 
+      as.data.frame(
+        fromJSON(filename)
+      )
+    }
+  )
+  # this doesnt feel like the right way to do this
+  # I cant do a plyr join as they dont have the same number of columns
+  master <- data.frame()
+  for (simulation in df){
     simulation <- data.frame(
       # select these columns for analysis
       simulation$traits.aggression,
@@ -32,14 +25,14 @@ for (simulation in df){
       "patch_area"
     )
     master <- rbind(master, simulation)
+  }
+  return(master)
 }
 
-setwd("../..")
-
-ggplot(master,aes(x = patch_area, y = aggression, colour = aggression)) + geom_point() + scale_color_gradient(low="blue",high="red")
-#ggplot(master,aes(x = patch_area, y = aggression)) + geom_density2d()
-
-#qplot(y = master$aggression, x = master$patch_area )
-.
+make_plot <- function (master){
+  ggplot(master,aes(x = patch_area, y = aggression, colour = aggression)) + geom_point() + scale_color_gradient(low="blue",high="red")
+  ggplot(master,aes(x = patch_area, y = aggression)) + geom_density2d()
+  
+}
 
 

### 2016-01-27 ms.tex
searching theory revised

--- 2016-01-23 ms.tex
@@
 
-We let the total area enclosing all males and all nests be a parameter, $A_{total}$.
+We let the total area enclosing all males and all nests be a parameter, $A_{total}$. We assume spatially random nests.
 
-The probability of a male that covers $A_{searched}$ area in time period $dt$ encountering any nest is:
-\begin{equation}
-    P(\rm {discovery}) = \frac {N \cdot A_{searched}} {A_{total}}
+The mean encounter rate $\varepsilon$ is:
+\begin{equation} \label{eq:encounter rate}
+    \varepsilon = \frac {N \cdot 2r\cdot v} {A_{total}}
 \end{equation}
-\citep{Gillespie-1976}
+\citep{Gurarie2012}
+
+We then use this as the rate parameter $lambda$ to repeatedly draw the time of the next discovery event for this individual from the exponential distribution, we do this for each male from when they mature up to when the females matrure.
+
+When an event occurs we select a nest randomly and then the male will either occupy it or contest another male for it.
+
+%I think I might use a priority queue for this...
+
 The energy used by a male with speed $v$ and radius $r$ is 

### 2016-01-27 python/simulation.py

@@ -65,7 +65,7 @@ def main():
     if params["sim_plot"]:
         sim.logger.plot(params["save_sim_trait_pngs"])
 
-    sim.logger.log_traits_to_JSON_file()
+    #sim.logger.log_traits_to_JSON_file()
 
 if __name__ == '__main__':
     main()

### 2016-01-29 R/load_datasets.R
early exit, aggro scaler

@@ -2,15 +2,16 @@ library(jsonlite)
 library(ggplot2)
 load_dataset  <- function (){
   df <- lapply(
-    dir(), 
+    dir()[grep("*.json", dir())], 
     function (filename){ 
       as.data.frame(
         fromJSON(filename)
       )
     }
   )
+  
   # this doesnt feel like the right way to do this
-  # I cant do a plyr join as they might not have the same number of columns
+  # I cant do a plyr join as they might not have the same number of columns as I modify
   master <- data.frame()
   for (simulation in df){
     simulation <- data.frame(
@@ -29,10 +30,11 @@ load_dataset  <- function (){
 }
 
 make_plot <- function (master){
-  print(ggplot(master, aes(x = patch_area, y = aggression, colour = aggression)) + geom_point() + scale_color_gradient(low="blue",high="red"))
-  
   print(ggplot(master, aes(x = patch_area, y = aggression)) + geom_density2d())
   
+  print(ggplot(master, aes(x = patch_area, y = aggression, colour = aggression)) + geom_point() + scale_color_gradient(low="blue",high="red"))
+  
+  print(qplot(df$aggression, fill = as.factor(df$patch_area)))
 }
 
 

### 2016-01-29 python/male.py

@@ -18,7 +18,7 @@ class Male(object):
             self.radius = abs(norm.rvs(params["radius_mean"], params["radius_sd"]))
             self.speed = abs(norm.rvs(params["speed_mean"], params["speed_sd"]))
             # pull the aggression from the normal distribution
-            self.aggro = params["aggression_max"] * uniform.rvs()
+            self.aggro = abs(norm.rvs(params["aggression_mean"], params["aggression_sd"]))
             self.maturation_time = abs(logistic.rvs(
                 params["maturation_center"],
                 params["maturation_width"]))
@@ -112,7 +112,6 @@ class Male(object):
     # populates a list of discovery events for this male
     # from the given time to the end 
     def fill_events(self, time = -1):
-        
         time = self.maturation_time if time == -1 else time
         time_left = self.params["time_female_maturity"] - time
         expected_num_events = time_left * self.exploration_rate

### 2016-01-29 python/nest.py

@@ -5,7 +5,6 @@ from male import Male
 """ a nest has a RR and female genetics and a male occupier"""
 
 """ TODO
-    abandoning nests
     check winner is alive after contest
 """
 class Nest(object):
@@ -15,6 +14,7 @@ class Nest(object):
         self.rr = params["rr_mean"]
         self.fight_cost = params["fight_cost"]
         self.display_1_cost = params["display_1_cost"]
+        self.aggression_scaler = params["aggression_scaler"]
         self.occupier = None
         self.debug = params["debug"]
 
@@ -46,7 +46,6 @@ class Nest(object):
         attacker.logger.inc_contest_energy(costs)
         defender.logger.inc_contest_energy(costs)
 
-
         if loser.id == defender.id:
             self.eject()
             self.occupy(attacker)
@@ -61,7 +60,7 @@ class Nest(object):
             print "attacker mass = %s, attacker aggro = %s" % (attacker.mass, attacker.aggro)
             print "defender mass = %s, defender aggro = %s" % (defender.mass, defender.aggro)
         
-        atk_escalation = logistic.cdf(attacker.mass * attacker.aggro - defender.mass)
+        atk_escalation = logistic.cdf((1.0/self.aggression_scaler) * (attacker.mass * attacker.aggro - defender.mass))
         rng = uniform.rvs()
         
         if rng > atk_escalation :
@@ -75,7 +74,7 @@ class Nest(object):
                 print "%s <= %s" % (rng, atk_escalation)
                 print "\tattacker escalates"
         
-        def_escalation = logistic.cdf(defender.mass * defender.aggro - attacker.mass)
+        def_escalation = logistic.cdf((1.0/self.aggression_scaler) * defender.mass * defender.aggro - attacker.mass)
         rng = uniform.rvs()
         if rng > def_escalation:
             if self.debug:

### 2016-01-29 python/run_sim.R

@@ -1,0 +1,3 @@
+source("../R/run_simulation.R")
+
+run_simulation("patch_area", 1000, 10000, 500, 8)

### 2016-01-29 ms.tex
updating contest equations

@@
 For a defender $x_{def}$ vs an attacker $x_{atk}$, there are 3 (currently 2) stages:
-each individual has an aggression trait $a$ and a mass $m$.
-\begin{equation}\label{eq:logit}
-    logit^{-1}(\alpha) = \frac{exp(\alpha)}{1 + exp(\alpha)}
+each individual $i$ has an aggression trait $a$ and a mass $m$. At any point in time it also has an amount of energy $E_{current}$.  
+The aggression of the individual is a function of $E_{current}$:
+
+%not sure what's going on with this one
+%\begin{equation}\label{eq:logit}
+ %   logit^{-1}(\alpha) = \frac{exp(\alpha)}{1 + exp(\alpha)}
+%\end{equation}
+
+\begin{equation}\label{eq:aggresstion}
+f(E_{current})=\frac{L}{1+e^{(-K(E_{current}-E_{o})}}
 \end{equation}
 
-$m$ is mass
-$a$ is aggression trait
-\begin{enumerate}
-    \item Display 1
-        Outcomes:
-        \begin{itemize}
-            \item $x_{atk}$ escalates: 
-                $Pr(\rm{escalation}) =  logit^{-1}(a_{atk} \cdot m_{atk} - m_{def})$
-                or leaves.
-            \item $x_{def}$ escalates:
-                $Pr(\rm{escalation}) =  logit^{-1}(a_{def} \cdot m_{def} - m_{atk})$ 
-                or leaves.
-            \item if the attacker leaves the defender keeps the nest
-            \item if the defender leaves the attacker keeps the nest
-            \item the attacker is evaluated first
-            \item if they both escalate then the next stage occurs
-        \end{itemize}
-    \item Display 2:
-        similar to stage 1, this could involve the $R$ value of the nest, or include the oponents aggression levels or energy reseves(?).
-    \item Fight
-        \begin{itemize}
-            \item $x_{def}$ wins:
-            $Pr(\rm{def}) = logit^{-1}(m_{def} - m_{atk})$
-        \end{itemize}
-\end{enumerate}
-The cost of the fighting stage is much greater than the cost of the display stage.
+The contests may last are escalated to different degrees with increasing energy costs for both participants.
+Either participant may abandon the contest at any moment.
+We assume that there are energetic costs, but that there is no time cost to contests.
+Traits $K$ and $E_{o}$ are heritable and potentially under selection.
 
-After the contest Energy costs are deducted from both the winner and the loser.
+The probability of the attacker escalating is:
 
-The winner also has a chance to abandon the RR (based off a trait).\\
+\begin{equation}\label{eq:attacker_escalating}
+Pr(\rm{escalation}) =  logit^{-1}(a_{atk} \cdot m_{atk} - m_{def})
+\end{equation}
 
-If a death occurs (Energy(x) less than 0) then x is removed from the individual pool.
+Similarly the probability of the defender escalating is:
+
+\begin{equation}\label{eq:defender_escalating}
+          Pr(\rm{escalation}) =  logit^{-1}(a_{def} \cdot m_{def} - m_{atk})
+\end{equation}
+
+The first stage is the display phase.  The second phase is the fight phase.  The cost of escalating for both is:
+
+\begin{equation}\label{eq:defender_escalating}
+E_{spent}=equation you guys figured out with all the \lambda 's
+\end{equation}
+
+The cost of the fighting stage is much greater than the cost of the display stage. Add a figure here to show this.
+
+If both individuals escalate to the end of the fight, the winner is chosen.  $x_{def}$ wins with the probability $P_{def} = logit^{-1}(m_{def} - m_{atk})$
+After the contest energy costs are deducted from both the winner and the loser.
+If a death occurs ($E \leq 0$) then x is removed from the individual pool.
 

### 2016-01-30 ms.tex
ms.tex clean up and exponential event jumping (no more urning)

@@
 The population is one with non-overlapping generations, and within each generation there is a predefined annual cycle.
-In spring, up to $K$individuals hatch from eggs, grow throughout spring and summer to increase size, and then within a short period, the entire population mates, females lay eggs and everyone dies.
+In spring, up to $K$ individuals hatch from eggs, grow throughout spring and summer to increase size, and then within a short period, the entire population mates, females lay eggs and everyone dies.
 The short duration of mating period is such that each male and female only mates once.
@@
 \subsection{Immature phase}
+% TODO update this to reflect what happens in the model
+% ie pullign the maturation times from the logistic cdf with location and parameters given
 
@@
 \end{equation}
+
 while the time $\tau$ taken for a male to grow from $m(0) \rightarrow m$ is then
@@
 \subsection{Searching Males}
-When males mature they begin searching for a females nest
+When males mature they begin searching for a females nest.
 
@@
 \begin{equation}
-    2r \cdot v
+    2 \cdot r \cdot v
 \end{equation}
@@
 
-The mean encounter rate $\varepsilon$ is:
+The encounter rate $\varepsilon$ for an individual is:
 \begin{equation} \label{eq:encounter rate}
-    \varepsilon = \frac {N \cdot 2r\cdot v} {A_{total}}
+    \varepsilon = \frac {N \cdot 2 \cdot r \cdot v} {A_{total}}
 \end{equation}
@@
 We then use this as the rate parameter $lambda$ to repeatedly draw the time of the next discovery event for this individual from the exponential distribution, we do this for each male from when they mature up to when the females matrure.
-
 When an event occurs we select a nest randomly and then the male will either occupy it or contest another male for it.
 
-%I think I might use a priority queue for this...
+When a male is kicked out of a nest we discard the old events and generate new ones.
 
@@
 
-\subsection{Occupying males}
-The probability of a male abandoning a nest $j$, is a function of the  the reproductive value of the nest $R_j$ and the abandon trait of the individual $\psi_i$:
+\clearpage
+\subsection{Contest Algorithm}
+See Fig. \ref{fig:contests}.
+For a defender $x_{def}$ vs an attacker $x_{atk}$, there are 3 (currently 2) stages:
+The first stage is the display stage, the second phase is the fight stage.
+each individual $i$ has mass $m$. At the time of the contest it also has an amount of energy $E_{current}$.  
+The aggression of the individual is a function of $E_{current}$:
 
-assumption: $1 < R_j$,  $1 < \psi_i$
+\begin{equation}\label{eq:aggression}
+a_i = f(E_{current})=\frac{L}{1+e^{(-K(E_{current}-E_{o})}}
+\end{equation}
 
-\begin{equation} \label{eq:abandon}
-    Pr({\rm abandon} | R_j ) = \frac{1}{1 + \psi_i \cdot R_j}
+$L$ is a scaling parameter.
+
+Contests can be escalated to different degrees with increasing energy costs for both participants. 
+Either participant may abandon the contest at any moment, they do this by choosing not to escalate.
+We assume that there are energetic costs, but that there is no time cost to contests ie they occur instantanously.
+
+Traits $K$ and $E_{o}$ are heritable and potentially under selection.
+
+The probability of the attacker escalating to  is:
+
+\begin{equation}\label{eq:attacker_escalating}
+Pr(\rm{escalation}) =  logit^{-1}(a_{atk} \cdot m_{atk} - m_{def})
 \end{equation}
+
+Similarly the probability of the defender escalating is:
+
+\begin{equation}\label{eq:defender_escalating}
+          Pr(\rm{escalation}) =  logit^{-1}(a_{def} \cdot m_{def} - m_{atk})
+\end{equation}
+
+The energetic cost of escalating is:
+\begin{equation}\label{eq:defender_escalating}
+E_{spent}=equation you guys figured out with all the \lambda 's
+\end{equation}
+
+The cost of the fighting stage is much greater than the cost of the display stage. Add a figure here to show this.
+
+%\begin{figure}[h!]
+%\centering
+%\includegraphics[width=10cm,height=10cm,keepaspectratio]{figures/example_costs}
+%\caption{Example of escalating energy costs,}
+%\label{fig:contests costs}
+%\end{figure}
+
+If both individuals escalate to the end of the fight, the winner is chosen.  $x_{def}$ wins with the probability $P_{def} = logit^{-1}(m_{def} - m_{atk})$
+After the contest energy costs are deducted from both the winner and the loser.
+If a death occurs ($E \leq 0$) then x is removed from the individual pool.
+
+\clearpage
 
@@
 
-    \item Aggressiveness, influences escalation in contests
+    \item Aggressiveness, determined by two inheritable traits $k$, $E_0$
     Trade offs:
@@
 \end{enumerate}
-
-\subsection{Algorithms}
-\subsection{Event Algorithms}
-So far we are considering two algorithms for stepping through events one is a stochastic simulation algorithm and a simpler (although more computationally expensive) naive stepping algorithm.
-\subsection{SSA}
-Simulating the population dynamics requires accounting for four types of events:
-
-An efficient method for stepping the population is to step between successive events.
-So rather than stepping the system over a given time interval and considering any transitions that may occur in that time period, the system is stepped between successive events by drawing the next event and the time at which the event occurs from appropriate distributions.
-We can identify time to next event using Gillespie minimal process method \citep{Gillespie-1976}, which indicates that with a given event rate, the distribution of waiting times follows an exponential distribution.
-
-In the SSA algorithm \citep{Gillespie-1976} there are 6 different states
-\begin{enumerate}
-    \item $S_1 = $ immature male
-    \item $S_2 = $ mature searching male
-    \item $S_3 = $ unoccupied nest
-    \item $S_4 = $ occupied nest
-    \item $S_5 = $ recently defeated male
-    \item $S_6 = $ recently changed nest
-\end{enumerate}
-
-these states have $R_\mu$ different interactions
-\begin{table}[h!]
-    \caption{state changes}
-    \centering
-    \begin{tabular}{c | c | c | c | l }
-        \hline
-            Event Number & reactants  & & products & Description\\
-        \hline
-        \hline
-            $R_1$       &   $S_1$       & $--c_1-->$      &   $ S_2$      & maturation\\
-            $R_2$       &   $S_2 + S_3$ & $--c_2-->$      &   $S_4 $      & occupation\\
-            $R_3$       &   $S_2 + S_4$ & $--c_3-->$      &   $S_4 + S_5$ & contest\\
-            $R_4$       &   $S_5$       & $--c_4-->$      &   $S_2$       & recovery\\
-            $R_5$       &   $S_2 + S_4$ & $--c_5-->$      &   $S_4$       & contest with death\\
-            $R_6$       &   $S_4$       & $--c_6-->$      &   $S_2 + S_3$ & abandon event\\
-            $R_7$       &   $S_1$       & $--c_7-->$      &   $0$         & metabolic / predation event\\
-            $R_8$       &   $S_2$       & $--c_8-->$      &   $0$         & predation event\\
-            $R_9$       &   $S_4$       & $--c_9-->$      &   $S_3$       & metabolic / predation event\\
-            $R_{10}$    &   $S_6$       & $--c_{10}-->$   &   $S_4$       & nest handover\\
-        \hline
-    \end{tabular}
-
-\end{table}
-
-$c_n$ is called the stochastic reaction constant\\
-$c_n \cdot dt = $ the average probability that a particular combination of $R_n$, reactant molecules will react accordingly in the next infinitesimal time interval $dt$\\
-s
-\begin{verbatim}
-I have the probability of encountering a   
-\end{verbatim}
-
-\citep{Gillespie-1976}
-
-\clearpage
-\subsection{Naive Event Algorithm}
-we have 3 lists
-\begin{verbatim}
-    immature males
-    searching males
-    nests
-\end{verbatim}
-
-\begin{verbatim}
-time = 0
-while time < time when the females mature:
-    check which males have matured
-        add them to the list of searching males
-    for each occupied nest:
-        deduct metabolic costs from the occupier
-        remove dead males
-
-    for each searching male m:
-        deduct metabolic costs from m
-        remove dead males
-        reduce time until next event for m
-        if time until next event <= 0:
-            reset the time until next event
-            if random > exploration_prob:
-                choose a nest at random
-                if it is occupied:
-                    contest the occupying male for control of the nest
-                    if m has won the contest:
-                        m occupy's n
-                        the loser is added to the list of searching males
-                else:
-                    occupy the nest
-    t += delta_time
-\end{verbatim}
-
-\clearpage
-\subsection{Contest Algorithm}
-See Fig. \ref{fig:contests}.
-For a defender $x_{def}$ vs an attacker $x_{atk}$, there are 3 (currently 2) stages:
-each individual $i$ has an aggression trait $a$ and a mass $m$. At any point in time it also has an amount of energy $E_{current}$.  
-The aggression of the individual is a function of $E_{current}$:
-
-%not sure what's going on with this one
-%\begin{equation}\label{eq:logit}
- %   logit^{-1}(\alpha) = \frac{exp(\alpha)}{1 + exp(\alpha)}
-%\end{equation}
-
-\begin{equation}\label{eq:aggresstion}
-f(E_{current})=\frac{L}{1+e^{(-K(E_{current}-E_{o})}}
-\end{equation}
-
-The contests may last are escalated to different degrees with increasing energy costs for both participants.
-Either participant may abandon the contest at any moment.
-We assume that there are energetic costs, but that there is no time cost to contests.
-Traits $K$ and $E_{o}$ are heritable and potentially under selection.
-
-The probability of the attacker escalating is:
-
-\begin{equation}\label{eq:attacker_escalating}
-Pr(\rm{escalation}) =  logit^{-1}(a_{atk} \cdot m_{atk} - m_{def})
-\end{equation}
-
-Similarly the probability of the defender escalating is:
-
-\begin{equation}\label{eq:defender_escalating}
-          Pr(\rm{escalation}) =  logit^{-1}(a_{def} \cdot m_{def} - m_{atk})
-\end{equation}
-
-The first stage is the display phase.  The second phase is the fight phase.  The cost of escalating for both is:
-
-\begin{equation}\label{eq:defender_escalating}
-E_{spent}=equation you guys figured out with all the \lambda 's
-\end{equation}
-
-The cost of the fighting stage is much greater than the cost of the display stage. Add a figure here to show this.
-
-If both individuals escalate to the end of the fight, the winner is chosen.  $x_{def}$ wins with the probability $P_{def} = logit^{-1}(m_{def} - m_{atk})$
-After the contest energy costs are deducted from both the winner and the loser.
-If a death occurs ($E \leq 0$) then x is removed from the individual pool.
 

### 2016-01-30 R/load_datasets.R

@@ -5,7 +5,8 @@ load_dataset  <- function (){
     dir()[grep("*.json", dir())], 
     function (filename){ 
       as.data.frame(
-        fromJSON(filename)
+        # ignore the trait history portion
+        fromJSON(filename)[1:2]
       )
     }
   )
@@ -17,12 +18,18 @@ load_dataset  <- function (){
     simulation <- data.frame(
       # select these columns for analysis
       simulation$traits.aggression,
-      simulation$parameters.patch_area
+      simulation$parameters.patch_area,
+      simulation$traits.energy_at_female_maturity,
+      simulation$traits.radius,
+      simulation$traits.speed
     )
     # rename them
     colnames(simulation) <- c(
       "aggression",
-      "patch_area"
+      "patch_area",
+      "energy_at_female_maturity",
+      "radius",
+      "speed"
     )
     master <- rbind(master, simulation)
   }

### 2016-01-30 python/generation.py

@@ -27,21 +27,18 @@ class Generation (object):
         self.nests = [Nest(params, i) for i in range(params["N"])]
 
         self.searching = []
-
-
-        if not prev_gen: # no genetics
+        
+        # first generation no genetics
+        if not prev_gen: 
             # create males
             self.searching = [Male(params, self.logger, i) for i in range(params["K"])]
-            
-        else: # previous generation
-            self.searching = []
+        
+        # or create new males based off of the winners in the previous generation    
+        else: 
             id_start = 0
             for p in prev_gen.winners:
                 self.searching += p.get_offspring(self.params, self.logger, id_start)
                 id_start = 1 + self.searching[-1].id
-
-            if self.debug:
-                print "next generation has %s individuals" % len(self.immature)
 
         # remove those that died in the immature phase
         self.searching = filter(lambda m: m.is_alive(), self.searching)
@@ -52,11 +49,15 @@ class Generation (object):
 
         # sort males by the time of their first event
         self.searching.sort(key = lambda x : x.events[0])
+
+        # step through events 
         self.run()
+
         if self.params["generation_plot"]:
             self.logger.plot_cohort()
 
-        # used to grow the next generation, a list of nests that are occupied
+        # used to grow the next generation, 
+        # self.winners is a list of nests that are occupied by males when the females mature
         self.winners = [n for n in self.nests if n.occupied()]
         self.winners = [n for n in self.winners if n.occupier.is_alive()]
 
@@ -86,7 +87,7 @@ class Generation (object):
             else:
                 self.occupation(chosen, dt)
             
-            # re sort the list
+            # re sort the searching list by the first element in the males event list 
             self.searching.sort(key = lambda x : x.events[0])
             
             self.logger.log_cohort()
@@ -96,14 +97,11 @@ class Generation (object):
             if n.occupied():
                 n.occupier.occupy(f_time)
 
-    # iterate over nests subtracting metabolic costs from occupying
-    # males
+    # chooses a nest and sees if a contest takes place
     def occupation(self, chosen, dt):
         # select a nest a random 
         index = int(uniform.rvs() * len(self.nests))
         nest = self.nests[index]
-
-        # deduct metabolic costs from the occupier
 
         if self.debug:
             print "male %s has discovered nest %s at %s" % (
@@ -113,19 +111,23 @@ class Generation (object):
 
         if nest.occupied():
             self.logger.inc_contests()
+            # deduct metabolic costs from the occupier from their last event until now
[…]

### 2016-01-30 python/generation_logger.py

@@ -70,8 +70,8 @@ class GenerationLogger(object):
 
         winners_matured = np.mean([m.maturation_time for m in occupying_males])
                 
-        plt.title("Gen %s Various generation metrics through time. dt = %s" 
-                % ( self.generation.id, self.generation.params["time_step"]))
+        plt.title("Gen %s Various generation metrics through time." 
+                % ( self.generation.id))
         plt.xlabel("time steps")
 
         plt.plot(times, searching, label = "searching males")
@@ -98,8 +98,8 @@ class GenerationLogger(object):
         contest_energy = self.data["contest_energy"]
         occupying_energy = self.data["occupying_energy"]
 
-        plt.title("Gen %s: total Energy expenditure through time. delta = %s" 
-                % (self.generation.id, self.generation.params["time_step"]))
+        plt.title("Gen %s: total Energy expenditure through time" 
+                % (self.generation.id))
         plt.xlabel("Time")
         plt.ylabel("Energy expenditure (J)")
 
@@ -112,7 +112,7 @@ class GenerationLogger(object):
         p1 = plt.Rectangle((0, 0), 1, 1, fc="red")
         p2 = plt.Rectangle((0, 0), 1, 1, fc="green")
         p3 = plt.Rectangle((0, 0), 1, 1, fc="blue")
-        plt.legend([p1, p2, p3], ["search", "occupying", "contest"])
+        plt.legend([p3, p2, p1], ["search", "occupying", "contest"])
 
         if savefig:
             plt.savefig("plots/gen_%03d_total_energy_expenditure.png" % self.generation.id)
@@ -153,7 +153,6 @@ class GenerationLogger(object):
         ]
 
         plt.plot(occupying_exploration_trait, occupying_aggro_trait, 'ro',)
-        plt.axis([0,20,0,10])
         plt.title("gen %s: trait values of %s winners" %( self.generation.id, len(occupying_exploration_trait)))
         
         plt.xlabel("2 * r * v")

### 2016-01-30 python/male.py

@@ -1,7 +1,7 @@
 #!/usr/bin/python
 
 import math
-from scipy.stats import norm, logistic, uniform, poisson
+from scipy.stats import norm, logistic, uniform, expon
 
 # a male individual in the population
 class Male(object):
@@ -9,7 +9,7 @@ class Male(object):
     def __init__(self, params, logger, id, mom = None, dad = None):
         self.id = id
         self.params = params
-        self.events = [1000000]
+        self.events = []
 
         self.logger = logger
 
@@ -75,7 +75,7 @@ class Male(object):
         b = params["growth_param_b"]
         self.mass = pow(params["initial_mass"], 1.0 - b) 
         self.mass += self.maturation_time * a * ( 1.0 - b )
-        self.mass = pow(self.mass, 1.0/(1.0 - b))
+        self.mass = pow(self.mass, 1.0 / (1.0 - b))
     
     # a male explores its surroundings
     # consumes energy from the time of the last event to now
@@ -113,17 +113,22 @@ class Male(object):
     # from the given time to the end 
     def fill_events(self, time = -1):
         time = self.maturation_time if time == -1 else time
-        time_left = self.params["time_female_maturity"] - time
-        expected_num_events = time_left * self.exploration_rate
-        num_events = poisson.rvs(expected_num_events)
-        # pull the time of each event from 
-        for i in range(num_events):
-            self.events.append( uniform.rvs(time,
-                                self.params["time_female_maturity"]))
-        self.events.sort()
+        
+        self.events = []
+        
+        # pull the wait time from the exponential distrobution with rate parameter = exporation rate
+        # in scipy.stats the rate parameter is scale = 1 / rate
+        next_event = time + expon.rvs(scale = 1.0 / self.exploration_rate)
+        
+        while next_event < self.params["time_female_maturity"]:
+            self.events.append(next_event)
+            next_event += expon.rvs(scale = 1.0 / self.exploration_rate)
 
-    def empty_events(self):
-        self.events = [1000000]
+        # for if the first event is after the females mature
+        self.events.append(1000000)
+
+
+        
 
     def to_string(self):
         return "%s: explor=%s\tmat=%s\tM=%s\tE=%s\t" % (

### 2016-01-30 python/simulation_logger.py

@@ -73,7 +73,8 @@ class SimulationLogger (object):
             m["energy_at_female_maturity"] = male.energy
 
             traits.append(m)
-        # the json that we will write to file
+
+        # the json that we will write to file to read later in R
         jsons = {
             "parameters":self.params,
             "traits":traits,
@@ -86,8 +87,8 @@ class SimulationLogger (object):
         f.write(json.dumps(jsons))
         f.close()
 
-    # returns true if the early exit conditions have been met
 
+    # returns true if the early exit conditions have been met
     def get_early_exit(self):
         last_aggro = self.data["std_aggro"][-1]
         last_exploration = self.data["std_exploration"][-1]

### 2016-01-30 python/male.py
aggression functions, but not integrated

@@ -127,8 +127,9 @@ class Male(object):
         # for if the first event is after the females mature
         self.events.append(1000000)
 
-
-        
+    """ returns the aggression level of the male """
+    def get_aggression(self):
+        return self.params["L"] * logistic.cdf(self.k * (self.energy - self.e_0))
 
     def to_string(self):
         return "%s: explor=%s\tmat=%s\tM=%s\tE=%s\t" % (

### 2016-01-30 python/nest.py

@@ -2,12 +2,15 @@
 import sys
 from scipy.stats import logistic, uniform
 from male import Male
-""" a nest has a RR and female genetics and a male occupier"""
 
 """ TODO
-    check winner is alive after contest
+    
+    make sure that if the winner is dead at the end of the contest 
+    and the loser is alive then the loser becomes the winner
+
 """
 class Nest(object):
+""" a nest can have a male occupier or be empty"""
 
     def __init__(self, params, id):
         self.id = id
@@ -22,6 +25,7 @@ class Nest(object):
         return bool(self.occupier)
 
     def contest(self, attacker):
+    """ a contest between males, the winner gets to occupy the nest"""
         defender = self.occupier
 
         # energy costs associated with this contest

### 2016-01-30 python/run_sim.R

@@ -1,3 +1,3 @@
 source("../R/run_simulation.R")
-
+# this script is called from pbs
 run_simulation("patch_area", 1000, 10000, 500, 8)

### 2016-02-02 scrap/aggression.Rmd
agression and costs change and some .Rmd's

---
title: "Aggression Exploration"
author: "John"
date: "1 February 2016"
output: html_document
---

This is for exploring aggression levels with the two inheritable traits "E_0" and "K"
Currently E_curr is the actual current energy of the male, I think that it would be better if it were a % of max?

$$ a_i = f( E_{current} ) = \frac {L}{1 + exp(-K(E_{current} - E_0))}  $$

```{R}
library(rasterVis)

k_min <- 0
k_max <- 2
e_0_min <- -2
e_0_max <- 2

L <- 2

get_level_plotable <- function (e_current){
  old_aggression <- function(L, k, e_0){
    L / (1 + exp(-k*(e_current - e_0)))
  }
  k_val <- seq(k_min, k_max, by = 0.01)
  e_0_val <- seq(e_0_min, e_0_max, by = 0.01)
  r <- raster(outer(k_val, e_0_val, function(a, b){ old_aggression(L, a, b)}), 
              xmn = k_min,
              xmx = k_max,
              ymn = e_0_min,
              ymx = e_0_max)
  return(r)
}
my_theme <- BuRdTheme()

my_at <- seq(0, L, length.out=length(my_theme$regions$col) - 1)
my_ckey <- list(at=my_at, col=my_theme$regions$col)

```

Lets see what this looks like for an individual with max energy (e_current = 1)

```{R}

levelplot(get_level_plotable(1), margin=FALSE, xlab = "K", ylab = "E_0", main = "Aggression", at = my_at, colorkey = my_ckey, par.settings=my_theme)

```

half energy (e_current = 0.5)

```{R}
levelplot(get_level_plotable(0.5), margin=FALSE, xlab = "K", ylab = "E_0", main = "Aggression", at = my_at, colorkey = my_ckey, par.settings=my_theme)
```

close to 0 energy (e_current = 0.1)
```{R}

levelplot(get_level_plotable(0.1), margin=FALSE, xlab = "K", ylab = "E_0", main = "Aggression", at = my_at, colorkey = my_ckey, par.settings=my_theme)

```

### 2016-02-02 scrap/costs.Rmd
agression and costs change and some .Rmd's

---
title: "Costs"
author: "John"
date: "1 February 2016"
output: html_document
---

Costs of fighting

I was thinking something like

$$ E_{costs} = \frac {E_{loser max}} {1 + exp(-(\frac{m_{winner} - m_{loser}}{k} - m_0)}$$

Where the energy cost of fighting is deducted from both sides.
$E_0$ describes the center, ie the cost to two evently matched opponents with $E_0 = 0$ is 50%

```{r}
library(rasterVis)
loser_masses <- seq(1, 100, by = 0.1)
winner_masses <- seq(1, 100, by = 0.1)
e_0 <- 0 
E_loser_max <- 100
cost_scaling <- 25

costs <- function( m_loser, m_winner){
    E_loser_max / (1 + exp(-((m_winner - m_loser)/cost_scaling - e_0)))
}

r <- raster(outer(loser_masses, winner_masses, costs), 
              xmn = 0,
              xmx = 100,
              ymn = 0,
              ymx = 100)
levelplot(r, margin = FALSE, 
          main = "Cost of Fighting", 
          xlab = "loser's mass",
          ylab = "winner's mass"
)
```

### 2016-02-02 python/generation.py

@@ -53,13 +53,14 @@ class Generation (object):
         # step through events 
         self.run()
 
-        if self.params["generation_plot"]:
-            self.logger.plot_cohort()
 
         # used to grow the next generation, 
         # self.winners is a list of nests that are occupied by males when the females mature
         self.winners = [n for n in self.nests if n.occupied()]
         self.winners = [n for n in self.winners if n.occupier.is_alive()]
+
+        if self.params["generation_plot"]:
+            self.logger.plot_cohort()
 
     def run(self):
         f_time = self.params["time_female_maturity"] 

### 2016-02-02 python/generation_logger.py

@@ -125,38 +125,43 @@ class GenerationLogger(object):
         if len(occupying_males) == 0:
             print "extintion event"
 
-        occupying_exploration_trait = [
+        winner_exploration_trait = [
             m.exploration
             for m in occupying_males
         ]
-        occupying_aggro_trait = [
-            m.aggro
+        winner_e_0_trait = [
+            m.e_0
+            for m in occupying_males
+        ]
+        winner_k_trait = [
+            m.k
             for m in occupying_males
         ]
 
-        plt.boxplot([occupying_aggro_trait, occupying_exploration_trait])
-        labels = ('aggression', 'exploration')
-        plt.xticks([1,2],labels)
+        plt.boxplot([winner_k_trait,winner_e_0_trait, winner_exploration_trait])
+        labels = ('k aggression', 'e_0 aggression', 'exploration (2 * r * v)')
+        plt.xticks([1,2,3],labels)
         plt.legend()
         plt.title("gen %s: trait distribution of winners" % self.generation.id)
         plt.show()
 
+    """ plots the aggression traits """
     def plot_trait_scatter(self, savefig = False):
-        occupying_males = self.get_occupying_males()
-        occupying_exploration_trait = [
-            m.exploration
-            for m in occupying_males
+        winners = self.get_occupying_males()
+        winners_e_0_traits = [
+            m.e_0
+            for m in winners
         ]
-        occupying_aggro_trait = [
-            m.aggro
-            for m in occupying_males
+        winners_k_traits = [
+            m.k
+            for m in winners
         ]
 
-        plt.plot(occupying_exploration_trait, occupying_aggro_trait, 'ro',)
-        plt.title("gen %s: trait values of %s winners" %( self.generation.id, len(occupying_exploration_trait)))
+        plt.plot(winners_k_traits, winners_e_0_traits, 'ro',)
+        plt.title("gen %s: aggrssion trait values of %s winners" %( self.generation.id, len(winners_k_traits)))
         
-        plt.xlabel("2 * r * v")
-        plt.ylabel("aggression")
+        plt.xlabel("k trait")
+        plt.ylabel("e_0 trait")
         if savefig:
             plt.savefig("plots/gen_%03d_winner_trait_values.png" % self.generation.id)
             plt.close()
@@ -199,6 +204,5 @@ class GenerationLogger(object):
     def get_occupying_males(self):
         return [
             n.occupier 
-            for n in self.generation.nests 
-            if n.occupied()
+            for n in self.generation.winners
         ]

### 2016-02-02 python/male.py

@@ -18,7 +18,9 @@ class Male(object):
             self.radius = abs(norm.rvs(params["radius_mean"], params["radius_sd"]))
             self.speed = abs(norm.rvs(params["speed_mean"], params["speed_sd"]))
             # pull the aggression from the normal distribution
-            self.aggro = abs(norm.rvs(params["aggression_mean"], params["aggression_sd"]))
+            self.k = abs(norm.rvs(params["k_mean"], params["k_sd"]))
+            self.e_0 = abs(norm.rvs(params["e_0_mean"], params["e_0_sd"]))
+            
             self.maturation_time = abs(logistic.rvs(
                 params["maturation_center"],
                 params["maturation_width"]))
@@ -28,7 +30,8 @@ class Male(object):
             
             self.radius = dad.radius
             self.speed = dad.speed
-            self.aggro = dad.aggro
+            self.k = dad.k
+            self.e_0 = dad.e_0
 
             if params["inherit_maturation_time"]:
                 self.maturation_time = dad.maturation_time
@@ -46,9 +49,12 @@ class Male(object):
                 self.speed += norm.rvs(0, mutation_sd)
 
             if uniform.rvs() < mutation_rate:
-                self.aggro += norm.rvs(0, mutation_sd)
+                self.e_0 += norm.rvs(0, mutation_sd)
 
-            self.aggro = 0 if self.aggro < 0 else self.aggro
+            if uniform.rvs() < mutation_rate:
+                self.k += norm.rvs(0, mutation_sd)
+
+            self.k = 0 if self.k < 0 else self.k
             self.speed = 0 if self.speed < 0 else self.speed
             self.radius = 0 if self.radius < 0 else self.radius
             self.maturation_time = 0 if self.maturation_time < 0 else self.maturation_time
@@ -58,6 +64,7 @@ class Male(object):
 
         # energy at maturation
         self.energy = self.mass * params["mass_to_energy"]
+        self.energy_max = self.energy
         # see if this male lives to maturity
         self.time_last_event = self.maturation_time
 
@@ -103,6 +110,7 @@ class Male(object):
     def is_alive(self):
         return self.energy >= 0
 
+    """ deducts the energy cost of occupying from the male"""
     def occupy(self, time):
         spent = (time - self.time_last_event) * self.metabolic_cost_occupy
         self.time_last_event = time
@@ -129,7 +137,11 @@ class Male(object):
 
[…]

### 2016-02-02 python/nest.py

@@ -10,14 +10,13 @@ from male import Male
 
 """
 class Nest(object):
-""" a nest can have a male occupier or be empty"""
+    """ a nest can have a male occupier or be empty"""
 
     def __init__(self, params, id):
         self.id = id
         self.rr = params["rr_mean"]
         self.fight_cost = params["fight_cost"]
         self.display_1_cost = params["display_1_cost"]
-        self.aggression_scaler = params["aggression_scaler"]
         self.occupier = None
         self.debug = params["debug"]
 
@@ -25,7 +24,7 @@ class Nest(object):
         return bool(self.occupier)
 
     def contest(self, attacker):
-    """ a contest between males, the winner gets to occupy the nest"""
+        """ a contest between males, the winner gets to occupy the nest"""
         defender = self.occupier
 
         # energy costs associated with this contest
@@ -42,7 +41,8 @@ class Nest(object):
         # fight
         if not loser:
             loser = self.fight(attacker, defender)
-            costs += self.fight_cost
+            costs += loser.get_fight_cost(attacker if loser == defender else defender)
+
         
         attacker.energy -= costs
         defender.energy -= costs
@@ -50,21 +50,28 @@ class Nest(object):
         attacker.logger.inc_contest_energy(costs)
         defender.logger.inc_contest_energy(costs)
 
+        if not defender.is_alive():
+            loser = defender
+
         if loser.id == defender.id:
             self.eject()
             self.occupy(attacker)
+
         return loser
 
     # the first display phase, the loser is returned or None
     def display_1(self, attacker, defender):
         ## attacker goes first
         # prob the attacker will escalate
+        atk_aggression = attacker.get_aggression()
+        def_aggression = defender.get_aggression()
+
         if self.debug:
             print "fight!"
-            print "attacker mass = %s, attacker aggro = %s" % (attacker.mass, attacker.aggro)
-            print "defender mass = %s, defender aggro = %s" % (defender.mass, defender.aggro)
+            print "attacker mass = %s, attacker aggro = %s" % (attacker.mass, atk_aggression)
+            print "defender mass = %s, defender aggro = %s" % (defender.mass, def_aggression)
         
-        atk_escalation = logistic.cdf((1.0/self.aggression_scaler) * (attacker.mass * attacker.aggro - defender.mass))
+        atk_escalation = logistic.cdf((attacker.mass * atk_aggression - defender.mass))
         rng = uniform.rvs()
         
[…]

### 2016-02-02 python/run_sim.R

@@ -1,3 +1,3 @@
+# this script is called from the pbs script
 source("../R/run_simulation.R")
-# this script is called from pbs
-run_simulation("patch_area", 1000, 10000, 500, 8)
+run_simulation("patch_area", 1000, 30000, 500, 8)

### 2016-02-02 python/simulation_logger.py

@@ -6,6 +6,8 @@ import json
 
 """
     This class will hold summary information about many generations in a simulation
+
+   TODO Fix early exit 
 """
 class SimulationLogger (object):
     def __init__(self, simulation, params):
@@ -15,8 +17,10 @@ class SimulationLogger (object):
             "generation": [],
             "mean_exploration": [],
             "std_exploration": [],
-            "mean_aggro": [],
-            "std_aggro": [],
+            "mean_e_0": [],
+            "std_e_0": [],
+            "mean_k": [],
+            "std_k": [],
             "mean_maturation":[],
             "std_maturation":[],
             "contest_energy":[],
@@ -30,15 +34,20 @@ class SimulationLogger (object):
 
         exploration_traits = map(lambda x: x.exploration, winning_males)
 
-        aggro_traits = map(lambda x: x.aggro, winning_males)
+        aggression_e_0_traits = map(lambda x: x.e_0, winning_males)
+        aggression_k_traits = map(lambda x: x.k, winning_males)
 
         maturations = map(lambda x: x.maturation_time, winning_males)
 
         self.data["generation"].append(gen.id)
         self.data["mean_exploration"].append(np.mean(exploration_traits))
         self.data["std_exploration"].append(np.std(exploration_traits))
-        self.data["mean_aggro"].append(np.mean(aggro_traits))
-        self.data["std_aggro"].append(np.std(aggro_traits))
+        
+        self.data["mean_e_0"].append(np.mean(aggression_e_0_traits))
+        self.data["std_e_0"].append(np.std(aggression_e_0_traits))
+        self.data["mean_k"].append(np.mean(aggression_k_traits))
+        self.data["std_k"].append(np.std(aggression_k_traits))
+        
         self.data["mean_maturation"].append(np.mean(maturations))
         self.data["std_maturation"].append(np.std(maturations))
         self.data["num_winners"].append(len(winning_males))
@@ -65,7 +74,8 @@ class SimulationLogger (object):
 
         for male in winners:
             m = {}
-            m["aggression"] = male.aggro
+            m["e_0"] = male.e_0
+            m["k"] = male.k
             m["speed"] = male.speed
             m["radius"] = male.radius
             m["maturation_time"] = male.maturation_time
@@ -90,6 +100,7 @@ class SimulationLogger (object):
 
     # returns true if the early exit conditions have been met
     def get_early_exit(self):
+        return False
         last_aggro = self.data["std_aggro"][-1]
[…]

### 2016-02-02 python/run_sim.R
small changes and clean up

@@ -1,3 +1,4 @@
 # this script is called from the pbs script
 source("../R/run_simulation.R")
-run_simulation("patch_area", 1000, 30000, 500, 8)
+
+run_simulation("patch_area", 1000, 30000, 250, 16)

### 2016-02-02 python/simulation_logger.py

@@ -100,12 +100,12 @@ class SimulationLogger (object):
 
     # returns true if the early exit conditions have been met
     def get_early_exit(self):
-        return False
-        last_aggro = self.data["std_aggro"][-1]
+        last_e_0 = self.data["std_e_0"][-1]
+        last_k = self.data["std_k"][-1]
         last_exploration = self.data["std_exploration"][-1]
-        print "std's of traits exp %s, aggression %s" % (last_exploration, last_aggro)
+        print "std's of traits exp %s, k %s, e_0 %s" % (last_exploration, last_k, last_e_0)
         cutoff = self.params["early_exit_sd_cutoff"]
-        if last_aggro < cutoff and last_exploration < cutoff:
+        if last_k < cutoff and last_e_0 < cutoffs and last_exploration < cutoff:
             if self.data["num_winners"][-1] > self.params["min_winners_cutoff"]:
                 return True
         return False

### 2016-02-05 ms.tex
integrated results into remake workflow

@@
 
-%\begin{figure}[h!]
-%\centering
-%\includegraphics[width=10cm,height=10cm,keepaspectratio]{figures/example_costs}
-%\caption{Example of escalating energy costs,}
-%\label{fig:contests costs}
-%\end{figure}
-
 If both individuals escalate to the end of the fight, the winner is chosen.  $x_{def}$ wins with the probability $P_{def} = logit^{-1}(m_{def} - m_{atk})$
@@
 \clearpage
+
+\section{Results} % (fold)
+\label{sec:results}
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=15cm,height=10cm,keepaspectratio]{figures/energy_pc_plot.pdf}
+    \caption{Percent energy expediture of cohort at generation 499}
+    \label{fig:energy_pc}
+\end{figure}
+
+As patch area increases the number of fight oppurtunities decreases, 
+This can probably explain the strong decrease in energy spent fighting. but I have no idea, early days yet.
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=15cm,height=12cm,keepaspectratio]{figures/energy_total_plot.pdf}
+    \caption{Total energy expeniture of cohort at generation 499}
+    \label{fig:energy_total}
+\end{figure}
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=15cm,height=10cm,keepaspectratio]{figures/aggression_density.pdf}
+    \caption{Number of successful males at different patch areas, gaps are because it took longer than 12 hours to run}
+    \label{fig:num_winners}
+\end{figure}
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=15cm,height=10cm,keepaspectratio]{figures/num_winners.pdf}
+    \caption{Number of successful males at different patch areas, gaps are because it took longer than 12 hours to run}
+    \label{fig:num_winners}
+\end{figure}
+
+\clearpage
+% section results (end)results
 
@@
 \caption{the area searched by a male in interval $dt$}
-\label{fig:area searched}
+\label{fig:area searched figure}
 \end{figure}

### 2016-02-05 scrap/aggression.Rmd
integrated results into remake workflow

@@
 
-k_min <- 0
-k_max <- 2
-e_0_min <- -2
-e_0_max <- 2
+k_min <- -5
+k_max <- 5
+e_0_min <- -5
+e_0_max <- 5
 
@@
   old_aggression <- function(L, k, e_0){
-    L / (1 + exp(-k*(e_current - e_0)))
+    L / (1 + exp(k*e_current + e_0))
   }
@@
 
-levelplot(get_level_plotable(1), margin=FALSE, xlab = "K", ylab = "E_0", main = "Aggression", at = my_at, colorkey = my_ckey, par.settings=my_theme)
+levelplot(get_level_plotable(1), margin=FALSE, xlab = "K", ylab = "E_0", main = "Aggression, energy = 100%", at = my_at, colorkey = my_ckey, par.settings=my_theme)
 
@@
 ```{R}
-levelplot(get_level_plotable(0.5), margin=FALSE, xlab = "K", ylab = "E_0", main = "Aggression", at = my_at, colorkey = my_ckey, par.settings=my_theme)
+levelplot(get_level_plotable(0.5), margin=FALSE, xlab = "K", ylab = "E_0", main = "Aggression, energy = 50%", at = my_at, colorkey = my_ckey, par.settings=my_theme)
 ```
@@
 
-levelplot(get_level_plotable(0.1), margin=FALSE, xlab = "K", ylab = "E_0", main = "Aggression", at = my_at, colorkey = my_ckey, par.settings=my_theme)
+levelplot(get_level_plotable(0.1), margin=FALSE, xlab = "K", ylab = "E_0", main = "Aggression, energy = 10%", at = my_at, colorkey = my_ckey, par.settings=my_theme)
 

### 2016-02-05 scrap/costs.Rmd
integrated results into remake workflow

@@
 E_loser_max <- 100
-cost_scaling <- 25
 
-costs <- function( m_loser, m_winner){
-    E_loser_max / (1 + exp(-((m_winner - m_loser)/cost_scaling - e_0)))
+costs <- function( m_loser, m_winner, k){
+    E_loser_max / (1 + exp(-((m_winner - m_loser) / k - e_0)))
 }
-
-r <- raster(outer(loser_masses, winner_masses, costs), 
+```
+costs_scaling <- 25:
+```{R}
+cost_scaling <- 25 # k 
+r <- raster(outer(loser_masses, winner_masses, function(a,b){costs(a, b, cost_scaling)}), 
               xmn = 0,
@@
               ymx = 100)
-levelplot(r, margin = FALSE, 
-          main = "Cost of Fighting", 
+
+plot(r, 
+  main = "Cost of Fighting k = 25", 
   xlab = "loser's mass",
-          ylab = "winner's mass"
-)
+  ylab = "winner's mass")
 ```
+
+costs_scaling <- 50:
+```{R}
+costs_scaling <- 1
+r <- raster(outer(loser_masses, winner_masses, function(a, b){costs(a, b, cost_scaling)}),
+              xmn = 0,
+              xmx = 100,
+              ymn = 0,
+              ymx = 100)
+
+plot(r,
+  main = "Cost of Fighting k = 1", 
+  xlab = "loser's mass",
+  ylab = "winner's mass")
+
+```

### 2016-02-05 R/download_data.R

@@ -1,0 +1,9 @@
+# downloads the data from google docs
+# not sure this is needed tbqh
+
+get_data <- function(){
+  library(httr)
+  url <- "https://drive.google.com/uc?export=download&id=0B3rkdRF2R-5rbVNlQzRFU3ZwLVU"
+  # I think there might be an issue with a prompt, also where to save the data?
+  GET(url)
+}

### 2016-02-05 R/load_datasets.R

@@ -1,6 +1,7 @@
-library(jsonlite)
-library(ggplot2)
-load_dataset  <- function (){
+# loads the dataset of trait values and parameters
+# returns a data.frame
+load_population  <- function (){
+  setwd("python/data")
   df <- lapply(
     dir()[grep("*.json", dir())], 
     function (filename){ 
@@ -15,33 +16,39 @@ load_dataset  <- function (){
   # I cant do a plyr join as they might not have the same number of columns as I modify
   master <- data.frame()
   for (simulation in df){
-    simulation <- data.frame(
-      # select these columns for analysis
-      simulation$traits.aggression,
-      simulation$parameters.patch_area,
-      simulation$traits.energy_at_female_maturity,
-      simulation$traits.radius,
-      simulation$traits.speed
-    )
+    simulation <- data.frame(simulation)
     # rename them
-    colnames(simulation) <- c(
-      "aggression",
-      "patch_area",
-      "energy_at_female_maturity",
-      "radius",
-      "speed"
-    )
+    colnames(simulation) <- lapply(colnames(simulation), 
+                                  function(x) {sub("^[^.]*.","",  x)})
     master <- rbind(master, simulation)
   }
+  setwd("../..")
   return(master)
 }
 
-make_plot <- function (master){
-  print(ggplot(master, aes(x = patch_area, y = aggression)) + geom_density2d())
-  
-  print(ggplot(master, aes(x = patch_area, y = aggression, colour = aggression)) + geom_point() + scale_color_gradient(low="blue",high="red"))
-  
-  print(qplot(df$aggression, fill = as.factor(df$patch_area)))
+# loads the trait history
+# and energy
+# returns a data.frame
+
+load_trait_history  <- function (){
+  setwd("python/data")
+  traits <- lapply(
+    dir()[grep("*.json", dir())], 
+    function (filename){ 
+      as.data.frame(
+        # ignore the trait history portion
+        fromJSON(filename)[2:3]
+      )
+    }
+  )
+  master <- data.frame()
+  for (simulation in traits){
+    simulation <- data.frame(simulation)
+    # rename them
+    colnames(simulation) <- lapply(colnames(simulation), 
+                                   function(x) {sub("^[^.]*.","",  x)})
+    master <- rbind(master, simulation)
+  }
+  setwd("../..")
+  return(master)
 }
-
-

### 2016-02-05 R/make_plots.R

@@ -1,0 +1,131 @@
+# what is what we are plotting the density of
+# returns a ggplot2 object
+density_plot <- function (data, what, title) {
+  return(
+    ggplot(data, aes_string(x = "patch_area", y = what)) + 
+    stat_density2d(aes(fill = ..density.. ), geom = "tile", contour = FALSE) + 
+    ggtitle(title)
+  )
+}
+
+# plots the number of winners
+winners_plot <- function (df){
+  plot( table( df$patch_area), 
+        ylab = "Number of Winning Males", 
+        xlab = "Patch Size",
+        main = "Winning males"
+  )
+}
+
+# plots k parameter 
+k_density_plot <- function(df) {
+  density_plot(df, "k", "Aggression parameter k density")
+}
+
+# plots e_0 density against area
+e_0_density_plot <- function(df) {
+  density_plot(df, "e_0", "Aggression parameter e_0 density")
+}
+
+# plots exploration trait density
+exploration_density_plot <- function (df){
+  df$exploration <- 2 * df$speed * df$radius
+  density_plot(df, "exploration", "Exploration Trait density")
+}
+
+
+# plots the multiplot of aggression densitys
+aggression_density_plots <- function(df) {
+
+  aggression <- function(e_current, k, e_0){
+    2 / (1 + exp(-k * ( e_current + e_0)))
+  }
+  
+  df$aggression_at_max <- aggression(1, df$k, df$e_0)
+  df$aggression_at_half <- aggression(0.5, df$k, df$e_0)
+  df$aggression_at_10pc <- aggression(0.1, df$k, df$e_0)
+
+  density_plot(df, "aggression_at_half", "Aggression at half of max energy")
+  density_plot(df, "aggression_at_10pc", "Aggression at 10% of max energy")
+  # todo make it return multiple plots
+  return(
+      density_plot(df, "aggression_at_max", "Aggression at max energy")
+  )
+
+}
+
+# select the last generation 
+# and adds some different variables
+get_generation <- function(m, gen){
+  final <- m[m$generation == gen,]
+  final$total_e <- final$contest_energy + final$occupying_energy + final$search_energy
+  final$contest_as_pc <- final$contest_energy / final$total_e
+  final$occupying_as_pc <- final$occupying_energy / final$total_e
+  final$search_as_pc <- final$search_energy / final$total_e
+  final$e_per_winner <- final$total_e / final$num_winners
+  return(final)
+}
+
[…]

### 2016-02-05 python/simulation_logger.py
small bugfix

@@ -105,7 +105,7 @@ class SimulationLogger (object):
         last_exploration = self.data["std_exploration"][-1]
         print "std's of traits exp %s, k %s, e_0 %s" % (last_exploration, last_k, last_e_0)
         cutoff = self.params["early_exit_sd_cutoff"]
-        if last_k < cutoff and last_e_0 < cutoffs and last_exploration < cutoff:
+        if last_k < cutoff and last_e_0 < cutoff and last_exploration < cutoff:
             if self.data["num_winners"][-1] > self.params["min_winners_cutoff"]:
                 return True
         return False

### 2016-02-06 ms.tex
more integration into remake

--- 2016-01-30 ms.tex
@@
 % TODO update this to reflect what happens in the model
-% ie pullign the maturation times from the logistic cdf with location and parameters given
+% ie pulling the maturation times from the logistic cdf with location and parameters given
 
@@
 
-We then use this as the rate parameter $lambda$ to repeatedly draw the time of the next discovery event for this individual from the exponential distribution, we do this for each male from when they mature up to when the females matrure.
+We then use this as the rate parameter $lambda$ to repeatedly draw the time of the next discovery event for this individual from the exponential distribution, we do this for each male from when they mature up to when the females mature.
 When an event occurs we select a nest randomly and then the male will either occupy it or contest another male for it.
@@
 Either participant may abandon the contest at any moment, they do this by choosing not to escalate.
-We assume that there are energetic costs, but that there is no time cost to contests ie they occur instantanously.
+We assume that there are energetic costs, but that there is no time cost to contests ie they occur instantaneously.
 
@@
 The cost of the fighting stage is much greater than the cost of the display stage. Add a figure here to show this.
-
-%\begin{figure}[h!]
-%\centering
-%\includegraphics[width=10cm,height=10cm,keepaspectratio]{figures/example_costs}
-%\caption{Example of escalating energy costs,}
-%\label{fig:contests costs}
-%\end{figure}
 
@@
 
+\section{Results} % (fold)
+\label{sec:results}
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=8cm,height=10cm,keepaspectratio]{figures/energy_pc_plot.pdf}
+    \caption{Percent energy expenditure of cohort at generation 499}
+    \label{fig:energy pc}
+
+    \centering
+    \includegraphics[width=8cm,height=10cm,keepaspectratio]{figures/energy_total_plot.pdf}
+    \caption{Total energy expenditure of cohort at generation 499}
+    \label{fig:energy total}
+\end{figure}
+As patch area increases the number of fight importunities decreases, 
+This can probably explain the strong decrease in energy spent fighting. but I have no idea, early days yet.
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=10cm,height=10cm,keepaspectratio]{figures/k_density.pdf}
+    \caption{a density plot of the aggression parameter k at max energy}
+    \label{fig:k density}
+
+    \includegraphics[width=10cm,height=10cm,keepaspectratio]{figures/e_0_density.pdf}
+    \caption{a density plot of the aggression parameter $e_0$ at max energy}
+    \label{fig:e_0 density}
+\end{figure}
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=15cm,height=8cm,keepaspectratio]{figures/aggression_density.pdf}
+    \caption{a density plot of aggression at max energy. I am trying to get it to be multiple plots}
+    \label{fig:aggression density}
+\end{figure}
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=15cm,height=10cm,keepaspectratio]{figures/num_winners.pdf}
+    \caption{Number of successful males at different patch areas, gaps are because it took longer than 12 hours to run}
+    \label{fig:num winners}
+\end{figure}
+
+\clearpage
+% section results (end)results
+
 I am not sure if this is the best way to do this, the tables in the plant paper are more aesthetic.\\
@@
         $N$ & & number of nests \\
-        $R_i$ & & repoductive value of nest of nest $i$\\
-        $\bar{R}$ & & average repoductive value of nest\\
+        $R_i$ & & reproductive value of nest of nest $i$\\
+        $\bar{R}$ & & average reproductive value of nest\\
         \hline
@@
 \centering
-\includegraphics[width=10cm,height=10cm,keepaspectratio]{figures/event_algorithm}
-\caption{Algorithm for modelling dynamics within the ``mature'' male population.}
-\label{fig:events}
-\end{figure}
-\clearpage
-
-\begin{figure}[h!]
-\centering
-\includegraphics[width=15cm,height=15cm,keepaspectratio]{figures/contest_algorithm}
-\caption{Algorithm for modelling contests behaviour within the ``mature'' male population.}
-\label{fig:contests}
-\end{figure}
-\clearpage
-
-\begin{figure}[h!]
-\centering
 \includegraphics[width=15cm,height=15cm,keepaspectratio]{figures/area_searched}
 \caption{the area searched by a male in interval $dt$}
-\label{fig:area searched}
+\label{fig:area searched figure}
 \end{figure}

### 2016-02-06 scrap/costs.Rmd
more integration into remake

@@
 winner_masses <- seq(1, 100, by = 0.1)
-e_0 <- 0 
+e_0 <- -0.5
 E_loser_max <- 100
@@
 ```{R}
-cost_scaling <- 25 # k 
+cost_scaling <- 100 # k 
 r <- raster(outer(loser_masses, winner_masses, function(a,b){costs(a, b, cost_scaling)}), 

### 2016-02-06 R/make_plots.R

@@ -86,7 +86,7 @@ as_stackable_total <- function (m){
 # returns a ggplot2 object
 history_plot <- function(df, what, bounds,  title) {
   return(
-    ggplot(data = last_gen, aes(patch_area)) +
+    ggplot(data = df, aes(patch_area)) +
       geom_ribbon(
         aes_string(
           ymin = paste(what, "-", bounds),
@@ -111,18 +111,23 @@ energy_stack_plot <- function (df, title){
 # wrapper for a % energy plot
 energy_pc_plot <- function (m){
   energy_stack_plot(as_stackable_pc(m), 
-    "Energy spending accross population")
+    "Energy spending across population")
 }
 
 # wrapper for a less "full plot that shows the decline in energy"
 energy_total_plot <- function(m){
   energy_stack_plot(as_stackable_total(m), 
-      "Energy spending accross population")
+      "Energy spending across population")
 }
 
-# takes 1 generation
-aggression_history_plot <- function (m){
+# takes 1 generation of multiple patch areas
+# ie all generation 300
+k_history_plot <- function (m){
   history_plot(m, "mean_k", "std_k", "Aggression, parameter K")
+}
+
+e_0_history_plot <- function (m){
+  history_plot(m, "mean_e_0", "std_e_0", "Aggression, parameter e_0")
 }
 
 # takes 1 generation

### 2016-02-06 python/generation.py
fight costs as a flat % fixed energy_reporting bug

@@ -58,7 +58,7 @@ class Generation (object):
         # self.winners is a list of nests that are occupied by males when the females mature
         self.winners = [n for n in self.nests if n.occupied()]
         self.winners = [n for n in self.winners if n.occupier.is_alive()]
-
+        
         if self.params["generation_plot"]:
             self.logger.plot_cohort()
 
@@ -94,9 +94,16 @@ class Generation (object):
             self.logger.log_cohort()
             self.time += dt
 
+        # final deductions for occupiers
         for n in self.nests:
             if n.occupied():
                 n.occupier.occupy(f_time)
+
+        # final deduction for the searchers
+        for m in self.searching:
+            m.search(f_time)
+            
+        self.logger.log_cohort()
 
     # chooses a nest and sees if a contest takes place
     def occupation(self, chosen, dt):

### 2016-02-06 python/male.py

@@ -10,7 +10,6 @@ class Male(object):
         self.id = id
         self.params = params
         self.events = []
-
         self.logger = logger
 
         # trait values
@@ -91,9 +90,14 @@ class Male(object):
         spent = (event_time - self.time_last_event ) * self.metabolic_cost_search
         self.events.pop(0)
 
-        self.energy -= spent
-        self.time_last_event = event_time
-        self.logger.inc_search_energy(spent)
+        # if the male has died searching
+        if spent >= self.energy:
+            self.logger.inc_search_energy(self.energy)
+            self.energy -= (spent + 1)
+        else:
+            self.time_last_event = event_time
+            self.energy -= spent
+            self.logger.inc_search_energy(spent)
 
     # generates a random number to see if the male survived to maturity
     # based off maturation time
@@ -139,9 +143,9 @@ class Male(object):
     def get_aggression(self):
         return self.params["L"] * logistic.cdf(self.k * ((self.energy / self.energy_max)- self.e_0))
 
+    # changed to 50% of total
     def get_fight_cost(self, winner):
-        return self.energy_max * logistic.cdf((winner.mass - self.mass)/ self.params["cost_scaler"] 
-            - self.params["cost_center"])
+        return self.energy_max * self.params["cost_scaler"]
 
     def to_string(self):
         return "%s: explor=%s\tmat=%s\tM=%s\tE=%s\t" % (

### 2016-02-06 python/nest.py

@@ -15,7 +15,7 @@ class Nest(object):
     def __init__(self, params, id):
         self.id = id
         self.rr = params["rr_mean"]
-        self.fight_cost = params["fight_cost"]
+
         self.display_1_cost = params["display_1_cost"]
         self.occupier = None
         self.debug = params["debug"]
@@ -99,6 +99,7 @@ class Nest(object):
 
         return None
 
+    # returns the loser of the fight
     def fight(self, attacker, defender):
         if self.debug:
             print "fight"

### 2016-02-07 ms.tex
comparison between energy budgets and cost functions

@@
     \includegraphics[width=8cm,height=10cm,keepaspectratio]{figures/energy_pc_plot.pdf}
-    \caption{Percent energy expenditure of cohort at generation 499}
+    \caption{Flat costs. Percent energy expenditure of cohort at generation 499}
     \label{fig:energy pc}
@@
     \includegraphics[width=8cm,height=10cm,keepaspectratio]{figures/energy_total_plot.pdf}
-    \caption{Total energy expenditure of cohort at generation 499}
+    \caption{ Flat costs Total energy expenditure of cohort at generation 499}
     \label{fig:energy total}
@@
 As patch area increases the number of fight importunities decreases,
-This can probably explain the strong decrease in energy spent fighting. but I have no idea, early days yet.
+This can probably explain the strong decrease in energy spent fighting.but I have no idea, early days yet.
+\clearpage
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=8cm,height=10cm,keepaspectratio]{figures/energy_pc_plot_var.pdf}
+    \caption{Variable costs. Percent energy expenditure of cohort at generation 499}
+    \label{fig:var energy pc}
+    \centering
+    \includegraphics[width=8cm,height=10cm,keepaspectratio]{figures/energy_total_plot_var.pdf}
+    \caption{Variable costs, Total energy expenditure of cohort at generation 499}
+    \label{fig:var energy total}
+\end{figure}
+
+The results for variable costs are different, why? from here on all graphs are from the flat cost data...
 
@@
     \label{fig:e_0 density}
+\end{figure}
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=10cm,height=10cm,keepaspectratio]{figures/k_history.pdf}
+    \caption{a line(ish) plot of the aggression parameter $k$ at max energy}
+    \label{fig:k history}
+
+    \includegraphics[width=10cm,height=10cm,keepaspectratio]{figures/e_0_history.pdf}
+    \caption{a line(ish) plot of the aggression parameter $e_0$  k at max energy}
+    \label{fig:e_0 history}
 \end{figure}

### 2016-02-07 R/load_datasets.R

@@ -1,7 +1,11 @@
 # loads the dataset of trait values and parameters
 # returns a data.frame
-load_population  <- function (){
-  setwd("python/data")
+# args:
+# directory: the folder in which the collection of jsons will be read from
+# levels: the number of levels to change directory back to
+# ie python/data -> 2
+load_population  <- function (directory  = "", levels = 0){
+  setwd(directory)
   df <- lapply(
     dir()[grep("*.json", dir())], 
     function (filename){ 
@@ -22,7 +26,7 @@ load_population  <- function (){
                                   function(x) {sub("^[^.]*.","",  x)})
     master <- rbind(master, simulation)
   }
-  setwd("../..")
+  setwd(Reduce(function(...) {paste(..., sep = "")}, rep("../", levels)))
   return(master)
 }
 
@@ -30,13 +34,13 @@ load_population  <- function (){
 # and energy
 # returns a data.frame
 
-load_trait_history  <- function (){
-  setwd("python/data")
+load_trait_history  <- function (directory  = "", levels = 0){
+  setwd(directory)
   traits <- lapply(
     dir()[grep("*.json", dir())], 
     function (filename){ 
       as.data.frame(
-        # ignore the trait history portion
+        # read the trait history portion
         fromJSON(filename)[2:3]
       )
     }
@@ -49,6 +53,6 @@ load_trait_history  <- function (){
                                    function(x) {sub("^[^.]*.","",  x)})
     master <- rbind(master, simulation)
   }
-  setwd("../..")
+  setwd(Reduce(function(...) {paste(..., sep = "")}, rep("../", levels)))
   return(master)
 }

### 2016-02-07 ms.tex
tex update: costs and maturationtime added

@@
 % TODO update this to reflect what happens in the model
-% ie pulling the maturation times from the logistic cdf with location and parameters given
 
@@
 
-while the time $\tau$ taken for a male to grow from $m(0) \rightarrow m$ is then
-\begin{equation} \label{eq:tau}
-\tau(m_0, m) = \frac1{a(1-b)}\left(m ^{1-b} - m_0 ^{1-b}\right).
-\end{equation}
+We Pull the maturation time from the logistic distribution with a  location and width parameters.
 
@@
 
+We are looking at two \'cost functions':
+\begin{enumerate}
+    \item A flat percentage of the losers maximum energy or...
+    \item $$ E_{costs} = \frac {E_{loser max}} {1 + exp(-(\frac{m_{winner} - m_{loser}}{k} - m_0)}$$ Where the fight cost is a function of the losers mass and the winners mass, and a couple of other parameters.
+\end{enumerate}
+
 \clearpage
@@
     \includegraphics[width=10cm,height=10cm,keepaspectratio]{figures/k_history.pdf}
-    \caption{a line(ish) plot of the aggression parameter $k$ at max energy}
+    \caption{a line(ish) plot of the aggression parameter $k$ at max energy, the grey bounds are the standard deviation}
     \label{fig:k history}
@@
     \includegraphics[width=10cm,height=10cm,keepaspectratio]{figures/e_0_history.pdf}
-    \caption{a line(ish) plot of the aggression parameter $e_0$  k at max energy}
+    \caption{a line(ish) plot of the aggression parameter $e_0$  k at max energy the grey bounds are the standard deviation}
     \label{fig:e_0 history}

### 2016-02-11 ms.tex
cts contests

@@
 
-    \item Aggressiveness, determined by two inheritable traits $k$, $E_0$
-    Trade offs:
-    \begin{itemize}
-        \item more likely to engage in fights with a low probability of winning
-    \end{itemize}
+    \item aggression 0 ($\beta$) called $e_0$
+    \item aggression 1 ($\alpha$)
 
-    \item search radius, the radius of a searching male (can be thought of as line of sight), See Fig. \ref{fig:area searched}.
+    \item Search radius, the radius of a searching male (can be thought of as line of sight), See Fig. \ref{fig:area searched}.
     Trade offs:

### 2016-02-11 scrap/aggression.Rmd
cts contests

--- 2016-02-02 scrap/aggression.Rmd
@@
 This is for exploring aggression levels with the two inheritable traits "E_0" and "K"
-Currently E_curr is the actual current energy of the male, I think that it would be better if it were a % of max?
+Currently E_curr is the actual current energy of the male?
 
@@
 
-k_min <- 0
-k_max <- 2
-e_0_min <- -2
-e_0_max <- 2
-
+k_min <- -5
+k_max <- 5
+e_0_min <- -5
+e_0_max <- 5
 L <- 2
+k_val <- seq(k_min, k_max, by = 0.01)
+e_0_val <- seq(e_0_min, e_0_max, by = 0.01)
 
@@
   }
-  k_val <- seq(k_min, k_max, by = 0.01)
-  e_0_val <- seq(e_0_min, e_0_max, by = 0.01)
-  r <- raster(outer(k_val, e_0_val, function(a, b){ old_aggression(L, a, b)}), 
+  return(raster(outer(k_val, e_0_val, function(a, b){ old_aggression(L, a, b)}), 
               xmn = k_min,
@@
               ymx = e_0_max)
-  return(r)
+  )
 }
@@
 
-levelplot(get_level_plotable(1), margin=FALSE, xlab = "K", ylab = "E_0", main = "Aggression", at = my_at, colorkey = my_ckey, par.settings=my_theme)
+levelplot(get_level_plotable(1), margin = FALSE, 
+          xlab = "K", 
+          ylab = "E_0",
+          main = "Aggression, energy = 100%", 
+          at = my_at, 
+          colorkey = my_ckey,
+          par.settings = my_theme)
 
@@
 
-half energy (e_current = 0.5)
+This kind of scaling is limited by a upper and lower being $0$ and $L$.
 
+Does this makes sense though? Why?? it seems like an arbitrary constraint and another parameter that the model maybe sensitive to.
+I also think that  having it be a funciton of current energy might be too strong an assumption.
+
+Do spiders really have a sense of how much energy they have "left"?
+
+Maybe Aggression should instead be the function inside of the logit ie
+$$ P(escalation) = logit^{-1}(f(male_i, male_j)$$
+
+the aggression of male $i$ against $j$ is now a function $f$.
+
+Some low order polynomial maybe:
+I think maybe of the difference in normalised mass:
+
+$$\bar{M_i} = \frac {M_i} {M_i + M_j}$$
+
+$$\bar{M_j} = \frac {M_j} {M_i + M_j}$$
+
+$$\Delta \bar{M} = \bar{M_i} - \bar{M_j}$$
+
+$$f(male_i, male_j) = g(\Delta M) = \beta + \alpha \cdot \Delta M $$
+
+Where the trait values:
+$$\alpha , \beta$$
+are inerited and selected for
+
+'
 ```{R}
-levelplot(get_level_plotable(0.5), margin=FALSE, xlab = "K", ylab = "E_0", main = "Aggression", at = my_at, colorkey = my_ckey, par.settings=my_theme)
+poly_1 <- function (k, e_0, delta_m) {
+  e_0 + k*delta_m
+}
+m <- outer(k_val, e_0_val, function (k, e_0) {poly_1 (k, e_0, 0)})
+levelplot(raster(m,
+                 xmn = k_min,
+                 xmx = k_max,
+                 ymn = e_0_min,
+                 ymx = e_0_max), 
+          main = "linear aggression",
+          xlab = "K", 
+          ylab = "E_0")
 ```
-
-close to 0 energy (e_current = 0.1)
-```{R}
-
-levelplot(get_level_plotable(0.1), margin=FALSE, xlab = "K", ylab = "E_0", main = "Aggression", at = my_at, colorkey = my_ckey, par.settings=my_theme)
-
-```

### 2016-02-11 scrap/continuous_escalation.Rmd
cts contests

---
title: "A step towards continuous Contests"
author: "John"
date: "9 February 2016"
output: pdf_document
---

I think we may have started to explore the concept of a 'continuous escalation' with escalating costs,
An assumption that greatly simplify the modelling of the IBM is that contests occur instantaneously.

I think more realistically, that contests escalate continuously along with the costs of escalating.
This would be nice as it would take care of several things in the model:
 - The concepts of Aggression and Escalation and Fighting costs are confusing and are 
 - These concepts are also pretty sensitive to arbitrary parameters
 - It removes discrete contest stages

We will assume an escaltion curve $kappa$ such that:
$$\kappa : Commitment \rightarrow Intensity$$
where $Intensity$ is $\frac{Energy}{Commitment}$

This example has a quadratic escaltion curve.

```{r, echo=FALSE}
library(ggplot2)
commitment <- seq(0,5, by = 0.01)

escalation_curve <- function (t){
  t^2
}

df <- data.frame("Commitment" = commitment, "Intensity" = escalation_curve(commitment))
ggplot(df,  aes( x = Commitment, y = Intensity )) +
  geom_line() +
  ggtitle("Example escalation curve") +
  geom_vline(xintercept = 3) + 
  geom_vline(xintercept = 4) +
  annotate("text",x = 2.8, y = 10, label = "c_ i") +
  annotate("text",x = 3.8, y = 20, label = "c_ j")
```

At the start of the contest each male chooses how much  energy they are willing to commit.
They do this by reporting their level of "commitment" $c_i$.

The male that chooses to commit less energy (above it is $c_i$) to the contest will be the the loser.

The energy that this male chose to commit will then be deducted from both males with the loser continuing to search for mates

The costs of fighting:
------

To get the cost of the fight we take the intergral of $\kappa$:

$$Energy Spent = \int_0^{c_i}{\kappa(t)}dt$$

if the $\kappa$ is $commitment^2$ then:
$$EnergySpent = \frac{c_i^3}{3}$$
Where $c_i$ is the losing male

How does the male decide how much to commit? 
-------

With a commitment funciton!
This commitment function will have the same structure for each male unless we want to venture into some genetic programming where we would grow an expression tree, but that raises more complexity also it would make the results hard to interpret.

An example of a commitment fucntion;
-------

A male can make an estimate about both their own and their opponents mass $M_i, M_j$ respectivly.

The error in this estimate is assumed to be close to zero but we can introduce some later ( wew Boltero)

We define normalised mass difference as so:
$$\bar{M_i} = \frac {M_i} {M_i + M_j}$$

$$\bar{M_j} = \frac {M_j} {M_i + M_j}$$

$$\Delta \bar{M} = \bar{M_i} - \bar{M_j}$$

The commitment that a male chooses is then a funciton:

$$c_i(\Delta \bar{M}) = \beta + \alpha \cdot \Delta \bar{M} $$

Where the trait values:
$$\alpha , \beta$$
Are under selection.

Introducing a stochastic element to these fights?
------

Where? in the males estimation of mass maybe?

We could say that probability that the winner is the one who commits more is equal to the logistic of the difference between their commitment values..

Example:
$$ c_i = 3 $$
$$ c_j = 2 $$

$$P(male_i wins) = logit^(c_i - c_j)$$

The lower energy cost is deducted from both sides

### 2016-02-11 python/generation.py

@@ -5,6 +5,7 @@ from generation_logger import GenerationLogger
 
 from scipy.stats import uniform
 import numpy as np
+import itertools
 
 # a generation in the model
 # aggregates males
@@ -53,7 +54,6 @@ class Generation (object):
         # step through events 
         self.run()
 
-
         # used to grow the next generation, 
         # self.winners is a list of nests that are occupied by males when the females mature
         self.winners = [n for n in self.nests if n.occupied()]
@@ -80,12 +80,12 @@ class Generation (object):
             # deduct metabolic costs from the chosen male
             chosen.search(self.time + dt)
             
-            if not chosen.is_alive():
+            if not chosen.is_alive(): # either we have died searching or we...
                 if self.debug:
                     print "male %s has died at %s" % (chosen.id, self.time)
                 self.searching.remove(chosen)
                 self.logger.inc_killed()
-            else:
+            else: # attempy to occupy a nest
                 self.occupation(chosen, dt)
             
             # re sort the searching list by the first element in the males event list 
@@ -94,14 +94,14 @@ class Generation (object):
             self.logger.log_cohort()
             self.time += dt
 
-        # final deductions for occupiers
+        # final energy deductions for occupiers
         for n in self.nests:
             if n.occupied():
                 n.occupier.occupy(f_time)
-
-        # final deduction for the searchers
+ 
+        # final energy deduction for the searchers
         for m in self.searching:
-            m.search(f_time)
+            m.search(f_time, True)
             
         self.logger.log_cohort()
 
@@ -131,11 +131,17 @@ class Generation (object):
 
             # if chosen has won the contest
             # we deal with the loser
-            if loser != chosen:
+
+            
+            if loser != chosen and loser.is_alive():
                 self.searching.insert(0, loser)
                 self.searching.remove(chosen)
                 self.logger.inc_take_overs()
                 loser.fill_events(self.time + dt)
+            elif not loser.is_alive():
+                self.logger.inc_killed()
+                if loser == chosen:
+                    self.searching.remove(loser)
 
             if self.debug:
                 print "\tnest %s is occupied by %s" % (

### 2016-02-11 python/male.py

@@ -6,14 +6,14 @@ from scipy.stats import norm, logistic, uniform, expon
 # a male individual in the population
 class Male(object):
     # constructor
-    def __init__(self, params, logger, id, mom = None, dad = None):
+    def __init__(self, params, logger, id, dad = None):
         self.id = id
         self.params = params
         self.events = []
         self.logger = logger
 
         # trait values
-        if not mom and not dad: # first gen: no breeding
+        if not dad: # first gen: no breeding
             self.radius = abs(norm.rvs(params["radius_mean"], params["radius_sd"]))
             self.speed = abs(norm.rvs(params["speed_mean"], params["speed_sd"]))
             # pull the aggression from the normal distribution
@@ -23,7 +23,7 @@ class Male(object):
             self.maturation_time = abs(logistic.rvs(
                 params["maturation_center"],
                 params["maturation_width"]))
-        elif not mom: # genetic inheritance from the dads
+        else: # genetic inheritance from the dads
             mutation_rate = params["mutation_rate"]
             mutation_sd = params["mutation_sd"]
             
@@ -53,7 +53,6 @@ class Male(object):
             if uniform.rvs() < mutation_rate:
                 self.k += norm.rvs(0, mutation_sd)
 
-            self.k = 0 if self.k < 0 else self.k
             self.speed = 0 if self.speed < 0 else self.speed
             self.radius = 0 if self.radius < 0 else self.radius
             self.maturation_time = 0 if self.maturation_time < 0 else self.maturation_time
@@ -86,9 +85,8 @@ class Male(object):
     # a male explores its surroundings
     # consumes energy from the time of the last event to now
     # removes the event from the list of events
-    def search(self, event_time):
+    def search(self, event_time, final = False):
         spent = (event_time - self.time_last_event ) * self.metabolic_cost_search
-        self.events.pop(0)
 
         # if the male has died searching
         if spent >= self.energy:
@@ -98,6 +96,9 @@ class Male(object):
             self.time_last_event = event_time
             self.energy -= spent
             self.logger.inc_search_energy(spent)
+        
+        if not final:
+            self.events.pop(0)
 
     # generates a random number to see if the male survived to maturity
     # based off maturation time
@@ -139,13 +140,11 @@ class Male(object):
         # for if the first event is after the females mature
         self.events.append(1000000)
 
-    """ returns the aggression level of the male """
[…]

### 2016-02-11 python/nest.py

@@ -14,11 +14,8 @@ class Nest(object):
 
     def __init__(self, params, id):
         self.id = id
-        self.rr = params["rr_mean"]
-
-        self.display_1_cost = params["display_1_cost"]
+        self.params = params
         self.occupier = None
-        self.debug = params["debug"]
 
     def occupied(self):
         return bool(self.occupier)
@@ -27,94 +24,60 @@ class Nest(object):
         """ a contest between males, the winner gets to occupy the nest"""
         defender = self.occupier
 
-        # energy costs associated with this contest
-        costs = 0
-        loser = None
-
         if not defender.is_alive():
+            self.eject()
+            self.occupy(attacker)
+            return defender
+
+
+        defender_commitment = defender.get_commitment(attacker)
+        attacker_commitment = attacker.get_commitment(defender)
+
+        # the defender wins
+        if defender_commitment >= attacker_commitment:
+            winner = defender
+            loser = attacker
+            delta_commitment = defender_commitment - attacker_commitment
+            cost = self.energy_spent(attacker_commitment)
+        else: # the attacker wins
+            winner = attacker
             loser = defender
+            delta_commitment = attacker_commitment - defender_commitment
+            cost = self.energy_spent(defender_commitment)
 
-        if not loser:
-            loser = self.display_1(attacker, defender)
-            costs += self.display_1_cost
+        # the winner wins with probability equal to the difference in commitment
+        # the chance of an upset happening is :
+        if logistic.cdf(delta_commitment) > uniform.rvs():
+            temp = winner
+            winner = loser
+            loser = temp
 
-        # fight
-        if not loser:
-            loser = self.fight(attacker, defender)
-            costs += loser.get_fight_cost(attacker if loser == defender else defender)
-
+        # make sure no more than the max energy of either male can be deducted
+        cost = min([winner.energy, loser.energy, cost])
         
-        attacker.energy -= costs
-        defender.energy -= costs
+        # the costs are deducted
+        attacker.energy -= cost
+        defender.energy -= cost
+        if self.params["debug"]:
+            print "\tfight cost =", cost
+            print "winner = %s , loser = %s" % (winner.id, loser.id)
 
-        attacker.logger.inc_contest_energy(costs)
-        defender.logger.inc_contest_energy(costs)
+        # the costs logged
+        attacker.logger.inc_contest_energy(cost)
+        defender.logger.inc_contest_energy(cost)
 
         if not defender.is_alive():
[…]

### 2016-02-11 python/simulation.py

@@ -7,6 +7,13 @@ from generation import Generation
 from simulation_logger import SimulationLogger
 
 # the simulation class aggregates generation objects
+"""
+    TODO
+
+    rename k -> a_alpha
+    and e_0  -> a_beta
+    across whole document
+"""
 class Simulation(object):
 
     # constructor

### 2016-02-11 python/simulation_logger.py

@@ -116,8 +116,10 @@ class SimulationLogger (object):
         
         plt.plot(gens, self.data["mean_exploration"], label = "mean exploration")
         plt.plot(gens, self.data["std_exploration"], label = "std exploration")
-        plt.plot(gens, self.data["mean_aggro"], label = "mean aggro")
-        plt.plot(gens, self.data["std_aggro"], label = "std aggro")
+        plt.plot(gens, self.data["mean_e_0"], label = "mean e_0")
+        plt.plot(gens, self.data["std_e_0"], label = "std e_0")
+        plt.plot(gens, self.data["mean_k"], label = "mean k")
+        plt.plot(gens, self.data["std_k"], label = "std k")
 
         plt.plot(gens, self.data["mean_maturation"], label = "mean maturationtime")
         plt.plot(gens, self.data["std_maturation"], label = "std maturation_time")

### 2016-02-13 ms.tex
fixed bug with cts contests

@@
 
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=8cm,height=10cm,keepaspectratio]{figures/energy_pc_plot_cts.pdf}
+    \caption{New cts contests. Percent energy expenditure of cohort at generation 499}
+    \label{fig:var energy pc}
+    \centering
+    \includegraphics[width=8cm,height=10cm,keepaspectratio]{figures/energy_total_plot_cts.pdf}
+    \caption{New cts contests, Total energy expenditure of cohort at generation 499}
+    \label{fig:var energy total}
+\end{figure}
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=10cm,height=10cm,keepaspectratio]{figures/cts_k_density.pdf}
+    \caption{New cts contests a density plot of the aggression parameter k at max energy, ie $\alpha$}
+    \label{fig:var k density}
+
+    \includegraphics[width=10cm,height=10cm,keepaspectratio]{figures/cts_e_0_density.pdf}
+    \caption{ New cts contests a density plot of the aggression parameter $e_0$ at max energy, ie $\beta$,
+    commitment = $\alpha \Delta \bar{M} + \beta$ }
+    \label{fig:var e_0 density}
+\end{figure}
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=10cm,height=10cm,keepaspectratio]{figures/cts_k_history.pdf}
+    \caption{New cts contests, k ie $\alpha$}
+    \label{fig:k history}
+
+    \includegraphics[width=10cm,height=10cm,keepaspectratio]{figures/cts_e_0_history.pdf}
+    \caption{New cts contests,  e 0 ie $\beta$}
+    \label{fig:e_0 history}
+\end{figure}
+
 \clearpage

### 2016-02-13 python/generation_logger.py

@@ -159,7 +159,7 @@ class GenerationLogger(object):
 
         plt.plot(winners_k_traits, winners_e_0_traits, 'ro',)
         plt.title("gen %s: aggrssion trait values of %s winners" %( self.generation.id, len(winners_k_traits)))
-        
+        plt.axis([-20,20,-20,20])
         plt.xlabel("k trait")
         plt.ylabel("e_0 trait")
         if savefig:

### 2016-02-13 python/male.py

@@ -143,8 +143,12 @@ class Male(object):
 
     # the commitment value for the male 
     def get_commitment(self, opponent):
-        n_mass_difference = (self.mass - opponent.mass)/(self.mass + opponent.mass)
-        return self.e_0 + (n_mass_difference * self.k)
+        if not self.params["exp_commitment"]:
+            n_mass_difference = (self.mass - opponent.mass)/(self.mass + opponent.mass)
+            return self.e_0 + (n_mass_difference * self.k)
+            
+        mass_ratio = (self.mass / opponent.mass)
+        return math.exp(self.e_0) * mass_ratio ** self.k
 
     def to_string(self):
         return "%s: explor=%s\tmat=%s\tM=%s\tE=%s\t" % (

### 2016-02-13 python/nest.py

@@ -20,7 +20,7 @@ class Nest(object):
     def occupied(self):
         return bool(self.occupier)
 
-    def contest(self, attacker):
+    def contest(self, attacker, log = False):
         """ a contest between males, the winner gets to occupy the nest"""
         defender = self.occupier
 
@@ -47,20 +47,24 @@ class Nest(object):
 
         # the winner wins with probability equal to the difference in commitment
         # the chance of an upset happening is :
-        if logistic.cdf(delta_commitment) > uniform.rvs():
+        prob_upset = logistic.cdf(self.params["aggression_logit_scaler"] * delta_commitment)
+        if prob_upset < uniform.rvs():
             temp = winner
             winner = loser
             loser = temp
 
+        if log:
+            print "gen, defender_mass, attacker_mass,",
+            print "defender_commitment, attacker_commitment,",
+            print "fight_cost,defender_energy, attacker_energy, prob_upset, defener_wins"
+            print "%s,%s,%s,%s,%s,%s,%s,%s,%s, %s" % (defender.logger.generation.id, defender.mass, attacker.mass, defender_commitment,
+                attacker_commitment, cost, defender.energy, attacker.energy, prob_upset, str(winner == defender))
         # make sure no more than the max energy of either male can be deducted
         cost = min([winner.energy, loser.energy, cost])
         
         # the costs are deducted
         attacker.energy -= cost
         defender.energy -= cost
-        if self.params["debug"]:
-            print "\tfight cost =", cost
-            print "winner = %s , loser = %s" % (winner.id, loser.id)
 
         # the costs logged
         attacker.logger.inc_contest_energy(cost)

### 2016-02-13 python/run_sim.R

@@ -1,4 +1,6 @@
 # this script is called from the pbs script
-source("../R/run_simulation.R")
+args <- commandArgs(trailingOnly = TRUE)
+args[2:5] <- as.numeric(args[2:5])
 
-run_simulation("patch_area", 1000, 30000, 250, 16)
+source("../R/run_simulation.R")
+run_simulation(args[1], args[2], args[3], args[4], args[5])

### 2016-02-13 python/simulation.py

@@ -59,9 +59,10 @@ def main():
     
     sim = Simulation(params)
     for i in range(params["generations"]):
-        print ""
-        print "gen =\t\t", i
-        print "patch area = \t", params["patch_area"]
+        if params["debug"]:
+            print ""
+            print "gen =\t\t", i
+            print "patch area = \t", params["patch_area"]
         sim.step()
         if i % params["save_every"] == 0:
             if params["save_scatter_pngs"]:

### 2016-02-13 python/simulation_logger.py

@@ -64,6 +64,11 @@ class SimulationLogger (object):
             gen.logger.data["occupying_energy"][-1]
         )
 
+        if self.params["debug"]:
+            print "contest_energy:", gen.logger.data["contest_energy"][-1]
+            print "search_energy: ", gen.logger.data["search_energy"][-1]
+            print "occupying_energy:", gen.logger.data["occupying_energy"][-1]
+
 
     def  log_traits_to_JSON_file(self):
         # get the winners

### 2016-02-13 python/run_sim.R
small run_sim bug

@@ -1,6 +1,8 @@
 # this script is called from the pbs script
 args <- commandArgs(trailingOnly = TRUE)
-args[2:5] <- as.numeric(args[2:5])
-
 source("../R/run_simulation.R")
-run_simulation(args[1], args[2], args[3], args[4], args[5])
+run_simulation(args[1], 
+    as.numeric(args[2]), 
+    as.numeric(args[3]), 
+    as.numeric(args[4]), 
+    as.numeric(args[5]))

### 2016-02-16 ms.tex
cts contest theory merged into ms, plots of cts in ms

--- 2016-02-11 ms.tex
@@
 \begin{equation} \label{eq:growth}
-\frac{{\rm d} m}{{\rm d} t} = a \, m^ b
+\frac{{\mathrm d} m}{{\mathrm d} t} = a \, m^ b
 \end{equation}
@@
 \begin{equation}\label{eq:allocation}
-Pr({\rm mature} \, | m) = \frac1{1 + \exp\left(\alpha \left(1 - m / m_{\rm mat}\right)\right)},
+Pr({\mathrm mature} \, | m) = \frac1{1 + \exp\left(\alpha \left(1 - m / m_{\mathrm mat}\right)\right)},
 \end{equation}
-where $m_{\rm mat}$ is the size at which 50\% of the population is mature and $\alpha$ determines the width of the curve.
+where $m_{\mathrm mat}$ is the size at which 50\% of the population is mature and $\alpha$ determines the width of the curve.
 
@@
 \subsection{Contest Algorithm}
-See Fig. \ref{fig:contests}.
-For a defender $x_{def}$ vs an attacker $x_{atk}$, there are 3 (currently 2) stages:
-The first stage is the display stage, the second phase is the fight stage.
-each individual $i$ has mass $m$. At the time of the contest it also has an amount of energy $E_{current}$.  
-The aggression of the individual is a function of $E_{current}$:
+Contests 
+When a searching male encounters a nest that is occupied a contest over possession of that nest occurs. Contests occur instantaneously.
 
-\begin{equation}\label{eq:aggression}
-a_i = f(E_{current})=\frac{L}{1+e^{(-K(E_{current}-E_{o})}}
-\end{equation}
+The costs of a contest escalates with more commitment.
 
-$L$ is a scaling parameter.
+We will assume an escalation curve $kappa$ such that:
+$$\kappa : Commitment \rightarrow Intensity$$
+where $Intensity$ is $\frac{Energy}{Commitment}$
 
-Contests can be escalated to different degrees with increasing energy costs for both participants. 
-Either participant may abandon the contest at any moment, they do this by choosing not to escalate.
-We assume that there are energetic costs, but that there is no time cost to contests ie they occur instantaneously.
+We use the escalation curve $$ \kappa (c) = c^2 $$
 
-Traits $K$ and $E_{o}$ are heritable and potentially under selection.
+At the start of the contest each male chooses how much  energy they are willing to commit.
+They do this by reporting their level of ``commitment'' $c_i$.
 
-The probability of the attacker escalating to  is:
+The male that chooses to commit more of its energy to the contest will win with probability $= logit^(k(c_i - c_j))$. Where k is is a parameter (I reduced it to 10 (also what )).
 
-\begin{equation}\label{eq:attacker_escalating}
-Pr(\rm{escalation}) =  logit^{-1}(a_{atk} \cdot m_{atk} - m_{def})
-\end{equation}
+The energy that this male chose to commit will then be deducted from both males with the loser continuing to search for mates
 
-Similarly the probability of the defender escalating is:
+To get the cost of the fight we take the integral of $\kappa$:
 
-\begin{equation}\label{eq:defender_escalating}
-          Pr(\rm{escalation}) =  logit^{-1}(a_{def} \cdot m_{def} - m_{atk})
-\end{equation}
+$$Energy Spent = \int_0^{c_i}{\kappa(t)}dt$$
 
-The energetic cost of escalating is:
-\begin{equation}\label{eq:defender_escalating}
-E_{spent}=equation you guys figured out with all the \lambda 's
-\end{equation}
+if the $\kappa$ is $commitment^2$ then
+$$EnergySpent = \frac{c_i^3}{3}$$
+Where $c_i$ is the losing male.
 
-The cost of the fighting stage is much greater than the cost of the display stage. Add a figure here to show this.
+The male chooses how much energy to commit based off of a commitment function, we examine two different commitment functions, for a male $i$ against a male $j$
 
-If both individuals escalate to the end of the fight, the winner is chosen.  $x_{def}$ wins with the probability $P_{def} = logit^{-1}(m_{def} - m_{atk})$
-After the contest energy costs are deducted from both the winner and the loser.
-If a death occurs ($E \leq 0$) then x is removed from the individual pool.
+An exponential one:
+$$\mathrm{commitment} = e^\beta \cdot (\frac{m_i}{m_j})^\alpha $$
 
-We are looking at two \'cost functions':
-\begin{enumerate}
-    \item A flat percentage of the losers maximum energy or...
-    \item $$ E_{costs} = \frac {E_{loser max}} {1 + exp(-(\frac{m_{winner} - m_{loser}}{k} - m_0)}$$ Where the fight cost is a function of the losers mass and the winners mass, and a couple of other parameters.
-\end{enumerate}
+And a linear one:
+
+$$\mathrm{commitment} = \beta + \alpha \cdot (\frac{m_i - m_j}{m_i + m_j})$$
+
+Where the trait values:
+$$\alpha , \beta$$
+Are under selection.
 
@@
     \item aggression 0 ($\beta$) called $e_0$
-    \item aggression 1 ($\alpha$)
+    \item aggression 1 ($\alpha$) called $k$
 
@@
 \label{sec:results}
-
+% energy expenditure plots
 \begin{figure}[h!]
     \centering
-    \includegraphics[width=8cm,height=10cm,keepaspectratio]{figures/energy_pc_plot.pdf}
-    \caption{Flat costs. Percent energy expenditure of cohort at generation 499}
+    \includegraphics[width=10cm,height=10cm,keepaspectratio]{figures/energy_pc_plot_exp_f10.pdf}
+    \caption{Exponential commitment function, ftime = 10. Percent energy expenditure of cohort at generation 499}
     \label{fig:energy pc}
@@
     \centering
-    \includegraphics[width=8cm,height=10cm,keepaspectratio]{figures/energy_total_plot.pdf}
-    \caption{ Flat costs Total energy expenditure of cohort at generation 499}
+    \includegraphics[width=10cm,height=10cm,keepaspectratio]{figures/energy_total_plot_exp_f10.pdf}
+    \caption{Exponential commitment function, ftime = 10. Total energy expenditure of cohort at generation 499}
     \label{fig:energy total}
 \end{figure}
-As patch area increases the number of fight importunities decreases,
-This can probably explain the strong decrease in energy spent fighting.but I have no idea, early days yet.
 \clearpage
@@
     \centering
-    \includegraphics[width=8cm,height=10cm,keepaspectratio]{figures/energy_pc_plot_var.pdf}
-    \caption{Variable costs. Percent energy expenditure of cohort at generation 499}
+    \includegraphics[width=10cm,height=10cm,keepaspectratio]{figures/energy_pc_plot_exp_f15.pdf}
+    \caption{Exponential commitment function, ftime = 15. Percent energy expenditure of cohort at generation 499}
     \label{fig:var energy pc}
     \centering
-    \includegraphics[width=8cm,height=10cm,keepaspectratio]{figures/energy_total_plot_var.pdf}
-    \caption{Variable costs, Total energy expenditure of cohort at generation 499}
+
+    \includegraphics[width=10cm,height=10cm,keepaspectratio]{figures/energy_total_plot_exp_f15.pdf}
+    \caption{Exponential commitment function, ftime = 15. Total energy expenditure of cohort at generation 499}
     \label{fig:var energy total}
@@
 
-The results for variable costs are different, why? from here on all graphs are from the flat cost data...
+% parameter plots density  plots
 
+% exp 10
 \begin{figure}[h!]
     \centering
-    \includegraphics[width=10cm,height=10cm,keepaspectratio]{figures/k_density.pdf}
-    \caption{a density plot of the aggression parameter k at max energy}
+    \includegraphics[width=10cm,height=10cm,keepaspectratio]{figures/exp_f10_k_density.pdf}
+    \caption{Exponential commitment function, ftime = 10. a density plot of the aggression parameter k}
     \label{fig:k density}
 
-    \includegraphics[width=10cm,height=10cm,keepaspectratio]{figures/e_0_density.pdf}
-    \caption{a density plot of the aggression parameter $e_0$ at max energy}
+    \includegraphics[width=10cm,height=10cm,keepaspectratio]{figures/exp_f10_e_0_density.pdf}
+    \caption{Exponential commitment function, ftime = 10. a density plot of the aggression parameter $e_0$}
     \label{fig:e_0 density}
@@
     \centering
-    \includegraphics[width=10cm,height=10cm,keepaspectratio]{figures/k_history.pdf}
-    \caption{a line(ish) plot of the aggression parameter $k$ at max energy, the grey bounds are the standard deviation}
+    \includegraphics[width=10cm,height=10cm,keepaspectratio]{figures/exp_f10_k_history.pdf}
+    \caption{Exponential commitment function, ftime = 10. $k$ the grey bounds are the standard deviation}
     \label{fig:k history}
 
-    \includegraphics[width=10cm,height=10cm,keepaspectratio]{figures/e_0_history.pdf}
-    \caption{a line(ish) plot of the aggression parameter $e_0$  k at max energy the grey bounds are the standard deviation}
+    \includegraphics[width=10cm,height=10cm,keepaspectratio]{figures/exp_f10_e_0_history.pdf}
+    \caption{ Exponential commitment function, ftime = 10. $e_0$ the grey bounds are the standard deviation}
     \label{fig:e_0 history}
+\end{figure}
+% exp 15
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=10cm,height=10cm,keepaspectratio]{figures/exp_f15_k_density.pdf}
+    \caption{Exponential commitment function, ftime = 15. a density plot of the aggression parameter k}
+    \label{fig:k density}
+
+    \includegraphics[width=10cm,height=10cm,keepaspectratio]{figures/exp_f15_e_0_density.pdf}
+    \caption{Exponential commitment function, ftime = 15. a density plot of the aggression parameter $e_0$.}
+    \label{fig:e_0 density}
 \end{figure}
@@
     \centering
-    \includegraphics[width=15cm,height=8cm,keepaspectratio]{figures/aggression_density.pdf}
-    \caption{a density plot of aggression at max energy. I am trying to get it to be multiple plots}
-    \label{fig:aggression density}
+    \includegraphics[width=10cm,height=10cm,keepaspectratio]{figures/exp_f15_k_history.pdf}
+    \caption{Exponential commitment function, ftime = 15. $k$ the grey bounds are the standard deviation}
+
+    \includegraphics[width=10cm,height=10cm,keepaspectratio]{figures/exp_f15_e_0_history.pdf}
+    \caption{Exponential commitment function, ftime = 15. $e_0$ the grey bounds are the standard deviation}
 \end{figure}
+
+%% WINNERS
 
@@
     \centering
-    \includegraphics[width=15cm,height=10cm,keepaspectratio]{figures/num_winners.pdf}
-    \caption{Number of successful males at different patch areas, gaps are because it took longer than 12 hours to run}
+    \includegraphics[width=15cm,height=10cm,keepaspectratio]{figures/exp_f10_num_winners.pdf}
+    \caption{exp ftime = 10, exp commitment Number of successful males at different patch areas}
+    \label{fig:num winners expf10}
+
+    \includegraphics[width=15cm,height=10cm,keepaspectratio]{figures/exp_f15_num_winners.pdf}
+    \caption{exp ftime = 15, exp commitment. Number of successful males at different patch areas, gaps are extinctions}
     \label{fig:num winners}

### 2016-02-16 scrap/continuous_escalation.Rmd
fixed bug with cts contests
cts contest theory merged into ms, plots of cts in ms

@@
 
-if the $\kappa$ is $commitment^2$ then:
+if the $\kappa$ is $commitment^2$ the:n
 $$EnergySpent = \frac{c_i^3}{3}$$
@@
 We define normalised mass difference as so:
-$$\bar{M_i} = \frac {M_i} {M_i + M_j}$$
-
-$$\bar{M_j} = \frac {M_j} {M_i + M_j}$$
+$$\bar{M_i} = \frac {M_i} {M_i + M_j}, \bar{M_j} = \frac {M_j} {M_i + M_j}$$
 

### 2016-02-16 python/nest.py

@@ -29,7 +29,6 @@ class Nest(object):
             self.occupy(attacker)
             return defender
 
-
         defender_commitment = defender.get_commitment(attacker)
         attacker_commitment = attacker.get_commitment(defender)
 

### 2016-02-16 scrap/contests.R

@@ -1,0 +1,29 @@
+contests_exp <- read.csv("contest_logs.csv", stringsAsFactors = FALSE)
+
+contests_lin <- read.csv("contest_logs_lin.csv", stringsAsFactors = FALSE)
+
+contests$gen <- as.numeric(contests$gen)
+contests$defender_mass <- as.numeric(contests$defender_mass) 
+contests$attacker_mass <- as.numeric(contests$attacker_mass)
+
+contests_lin$gen <- as.numeric(contests_lin$gen)
+contests_lin$defender_mass <- as.numeric(contests_lin$defender_mass) 
+contests_lin$attacker_mass <- as.numeric(contests_lin$attacker_mass)
+
+
+ggplot(contests_exp, aes( x = factor(gen), y = fight_cost)) +
+  geom_boxplot() +
+  ggtitle("exp")
+
+ggplot(contests_lin, aes( x = factor(gen), y = fight_cost)) +
+  geom_boxplot() +
+  ggtitle("lin")
+
+ggplot(contests_exp, aes( x = factor(gen), y = attacker_commitment)) +
+  geom_boxplot() +
+  ggtitle("exp")
+
+ggplot(contests_lin, aes( x = factor(gen), y = attacker_commitment)) +
+  geom_boxplot() +
+  ggtitle("lin")
+  

### 2016-02-16 python/simulation_logger.py
prep for data cleaning

@@ -96,7 +96,7 @@ class SimulationLogger (object):
             "history":self.data
         }
 
-        file_name = "data/" + strftime("%m_%d_%H_%M_%S__", gmtime()) + str(uuid.uuid4()) + ".json"
+        file_name = "../data/" + strftime("%m_%d_%H_%M_%S__", gmtime()) + str(uuid.uuid4()) + ".json"
         
         f = open(file_name,"w")
         f.write(json.dumps(jsons))

### 2016-02-17 R/load_datasets.R
lin commitment

@@ -2,57 +2,28 @@
 # returns a data.frame
 # args:
 # directory: the folder in which the collection of jsons will be read from
-# levels: the number of levels to change directory back to
-# ie python/data -> 2
-load_population  <- function (directory  = "", levels = 0){
-  setwd(directory)
-  df <- lapply(
-    dir()[grep("*.json", dir())], 
-    function (filename){ 
-      as.data.frame(
-        # ignore the trait history portion
-        fromJSON(filename)[1:2]
-      )
-    }
-  )
-  
-  # this doesnt feel like the right way to do this
-  # I cant do a plyr join as they might not have the same number of columns as I modify
-  master <- data.frame()
-  for (simulation in df){
-    simulation <- data.frame(simulation)
-    # rename them
-    colnames(simulation) <- lapply(colnames(simulation), 
-                                  function(x) {sub("^[^.]*.","",  x)})
-    master <- rbind(master, simulation)
-  }
-  setwd(Reduce(function(...) {paste(..., sep = "")}, rep("../", levels)))
+# plyr <3 <3 <3 <3 <3
+
+load_population  <- function (directory  = ""){
+  master <- ldply(list.files(directory, full.names = TRUE),                   
+                  function (filename) {
+                    data.frame(
+                      fromJSON(filename)[1:2]
+                    )})
+  colnames(master) <- sub("^[^.]*.","", colnames(master))
   return(master)
 }
 
 # loads the trait history
 # and energy
 # returns a data.frame
-
-load_trait_history  <- function (directory  = "", levels = 0){
-  setwd(directory)
-  traits <- lapply(
-    dir()[grep("*.json", dir())], 
-    function (filename){ 
-      as.data.frame(
-        # read the trait history portion
-        fromJSON(filename)[2:3]
-      )
-    }
-  )
-  master <- data.frame()
-  for (simulation in traits){
-    simulation <- data.frame(simulation)
-    # rename them
-    colnames(simulation) <- lapply(colnames(simulation), 
-                                   function(x) {sub("^[^.]*.","",  x)})
-    master <- rbind(master, simulation)
-  }
-  setwd(Reduce(function(...) {paste(..., sep = "")}, rep("../", levels)))
+# plyr <3 <3 <3
+load_trait_history  <- function (directory  = ""){
+  master <- ldply(list.files(directory, full.names = TRUE),                   
+                  function (filename) {
+                    data.frame(
+                      fromJSON(filename)[2:3]
[…]

### 2016-02-18 scrap/evolution.Rmd
new plots and evolution rmd

---
title: "Evolution plots"
author: "John Wilshire"
date: "18 February 2016"
output:
  pdf_document:
    fig_crop: no
    fig_height: 4
  html_document: default
---

```{r, echo=FALSE}
library(jsonlite)
library(plyr)
library(ggplot2)
source("../R/load_datasets.R")
source("../R/make_plots.R")
```

These are roughly equivilant to the density plots but I feel the density plot is a little bit too messy...

```{R}
expf10_pop <- load_population("../data/exp_f10/")
expf15_pop <- load_population("../data/exp_f15/")
linf10_pop <- load_population("../data/lin_f10/")
population_plots <- function (what){
  print(pop_values(expf10_pop, what, paste("expf10_pop", what)))
  print(pop_values(expf15_pop, what, paste("expf15_pop", what)))
  print(pop_values(linf10_pop, what, paste("linf10_pop", what)))
}

population_plots("e_0")
population_plots("k")
population_plots("2 * radius * speed")

```

These plots trace how parameters evolve through generations with coloured by patch area...

```{r}
expf10 <- load_trait_history("../data/exp_f10/")
expf15 <- load_trait_history("../data/exp_f15/")
linf10 <- load_trait_history("../data/lin_f10/")

evolution_plots <- function (what){
  print(trait_evolution_plot(expf10, what, paste("expf10", what)))
  print(trait_evolution_plot(expf15, what, paste("expf15", what)))
  print(trait_evolution_plot(linf10, what, paste("linf10", what)))
}

evolution_plots("mean_k")
evolution_plots("std_k")
evolution_plots("std_e_0")
evolution_plots("num_winners")
evolution_plots("mean_maturation")
evolution_plots("contest_energy")
evolution_plots("search_energy")
evolution_plots("occupying_energy")
evolution_plots("mean_exploration")
evolution_plots("std_exploration")
```

### 2016-02-18 R/download_data.R

@@ -1,9 +1,0 @@
-# downloads the data from google docs
-# not sure this is needed tbqh
-
-get_data <- function(){
-  library(httr)
-  url <- "https://drive.google.com/uc?export=download&id=0B3rkdRF2R-5rbVNlQzRFU3ZwLVU"
-  # I think there might be an issue with a prompt, also where to save the data?
-  GET(url)
-}

### 2016-02-18 R/make_plots.R

@@ -33,27 +33,6 @@ exploration_density_plot <- function (df){
   density_plot(df, "exploration", "Exploration Trait density")
 }
 
-
-# plots the multiplot of aggression densitys
-aggression_density_plots <- function(df) {
-
-  aggression <- function(e_current, k, e_0){
-    2 / (1 + exp(-k * ( e_current + e_0)))
-  }
-  
-  df$aggression_at_max <- aggression(1, df$k, df$e_0)
-  df$aggression_at_half <- aggression(0.5, df$k, df$e_0)
-  df$aggression_at_10pc <- aggression(0.1, df$k, df$e_0)
-
-  density_plot(df, "aggression_at_half", "Aggression at half of max energy")
-  density_plot(df, "aggression_at_10pc", "Aggression at 10% of max energy")
-  # todo make it return multiple plots
-  return(
-      density_plot(df, "aggression_at_max", "Aggression at max energy")
-  )
-
-}
-
 # select the last generation 
 # and adds some different variables
 get_generation <- function(m, gen){
@@ -134,3 +113,24 @@ e_0_history_plot <- function (m){
 exploration_history_plot <- function (m){
   history_plot(m, "mean_exploration", "std_exploration", "exploration traits")
 }
+
+trait_evolution_plot <- function(df, what, title){
+  return(
+    ggplot(df, aes_string(x = "generation", y = what, group = "patch_area" )) +
+      geom_line(aes(col = patch_area)) +
+      ggtitle(title)
+  )
+}
+
+
+# plots the given trait against patch area
+# takes a population (not trait history)
+# like a clearer density plot...
+pop_values <- function (df, what, title){
+  ggplot(df, aes_string(x = "patch_area", y = what)) + 
+    geom_point(aes_string(col = what)) +
+    geom_smooth(method=lm) +
+    ggtitle(title)
+}
+
+

### 2016-02-18 scrap/evolution.Rmd
added nest density plots

@@
 linf10_pop <- load_population("../data/lin_f10/")
-population_plots <- function (what){
-  print(pop_values(expf10_pop, what, paste("expf10_pop", what)))
-  print(pop_values(expf15_pop, what, paste("expf15_pop", what)))
-  print(pop_values(linf10_pop, what, paste("linf10_pop", what)))
+population_plots <- function (y, x){
+  print(pop_values(expf10_pop, x, y, paste("expf10_pop", y)))
+  print(pop_values(expf15_pop, x, y, paste("expf15_pop", y)))
+  print(pop_values(linf10_pop, x, y, paste("linf10_pop", y)))
 }
 
-population_plots("e_0")
-population_plots("k")
-population_plots("2 * radius * speed")
+population_plots("e_0", "patch_area")
+population_plots("k", "patch_area")
+
+```
+These plots are against the log of the nest density ie log(nests / patch_area)
+```{R}
+population_plots("e_0", "log(N / patch_area)")
+population_plots("k", "log(N / patch_area)")
 

### 2016-02-18 R/make_plots.R

@@ -126,9 +126,9 @@ trait_evolution_plot <- function(df, what, title){
 # plots the given trait against patch area
 # takes a population (not trait history)
 # like a clearer density plot...
-pop_values <- function (df, what, title){
-  ggplot(df, aes_string(x = "patch_area", y = what)) + 
-    geom_point(aes_string(col = what)) +
+pop_values <- function (df, x, y, title){
+  ggplot(df, aes_string(x = x, y = y)) + 
+    geom_point() +
     geom_smooth(method=lm) +
     ggtitle(title)
 }

### 2016-02-24 ms.tex
lin commitment
ms cleanup

@@
 Filled nests can only be obtained by entering into and winning a contest.
-Once occupied, males can choose to occupy a nest or continue searching.
-At the end of each season, each male has a fitness defined by the reproductive value of the nest it holds, $R_i$.
+At the end of each season, each male has a fitness defined by the fecundity of the nest it holds.
 
@@
 
-Denote $N$ to be the total number of available nests, $R_i$ to be the reproductive value of nest $i$ (number of offspring per nest per year), $\bar{R}$ to be the average reproductive value of all nests (number of offspring per nest per year), and $p(R)$ to be the probability density distribution of nest quality with respect to $R$.
-The distribution of nest qualities is then set according to the distribution
-\begin{equation} \label{eq:pdf_R}
-    Pr(R_i = x) =\bar{R} \, p(R).
-\end{equation}
+Denote $N$ to be the total number of available nests, The fecundity (number of offspring per nest per year) is uniform across the population of nests (females) and is given the symbol $F$.
+
 The total number of male offspring produced in each generation is then
-\begin{equation} \label{eq:pdf_R}
-    K = \bar{R} N.
+\begin{equation} \label{eq:pdf_F}
+    K = F N
 \end{equation}
@@
 \subsection{Immature phase}
-% TODO update this to reflect what happens in the model
 
@@
 \begin{equation} \label{eq:growth}
-\frac{{\mathrm d} m}{{\mathrm d} t} = a \, m^ b
+    \frac{{\mathrm d} m}{{\mathrm d} t} = a m^ b
 \end{equation}
+
 where $a$ is mass-based rate of growth. Integrating eq. \ref{eq:growth}, the size of individuals at time $t$, having started growing at time $0$ is given by
@@
 
-We Pull the maturation time from the logistic distribution with a  location and width parameters.
+We Pull the maturation time of a male randomly from the logistic distribution with location and width parameters.
 
-If we assume a constant mortality rate $k_i$ during the immature phase, the probability of males surviving to time $t$ is given by
+If we assume a constant mortality rate $k_{im}$ during the immature phase, the probability of a male surviving to time $t$ from $0$ is given by
 \begin{equation} \label{eq:surv_immature}
-S(t) = \exp(-k_i \, t).
+    S(t) = \exp(-k_i t).
 \end{equation}
 
-Finally, we assume males transition from immature $\rightarrow$ mature according to logistic probability function
-\begin{equation}\label{eq:allocation}
-Pr({\mathrm mature} \, | m) = \frac1{1 + \exp\left(\alpha \left(1 - m / m_{\mathrm mat}\right)\right)},
-\end{equation}
-where $m_{\mathrm mat}$ is the size at which 50\% of the population is mature and $\alpha$ determines the width of the curve.
+When a male matures it is given an energy budget linearly proportional to its mass such that: 
+$$ E = k_em$$\\
 
 \subsection{Mature phase}
-When a male matures it is given an energy budget linearly proportional to its mass, this is a parameter into the model.\\
-
 \subsection{Searching Males}
@@
 \clearpage
+
 \subsection{Contest Algorithm}
@@
 
-The male that chooses to commit more of its energy to the contest will win with probability $= logit^(k(c_i - c_j))$. Where k is is a parameter (I reduced it to 10 (also what )).
+The male that chooses to commit more of its energy to the contest will win with probability $= logit^{-1}(k(c_i - c_j))$.
+Where k is is a parameter that shapes how sensitive the contest is to small differences in commitment values.
 
@@
 A new individual will inherit the fathers genes,
-There is a global mutation rate parameter which is the chance that a trait will not be inherited perfectly but instead will be mutated.
+There is a mutation rate parameter which is the chance that a trait will not be inherited perfectly but instead will be mutated.
+
 This means adding some normally distributed noise with $\mu = 0$ and $\sigma$ varying from trait to trait.
@@
 \begin{enumerate}
-    \item Abandon Rate, influences RR abandonment rate (not yet implemented)
-    Trade offs:
-    \begin{itemize}
-        \item more likely to abandon a nest and not discover another before the females mature
-    \end{itemize}
-
     \item aggression 0 ($\beta$) called $e_0$
@@
 
-    \item Search radius, the radius of a searching male (can be thought of as line of sight), See Fig. \ref{fig:area searched}.
-    Trade offs:
-    \begin{itemize}
-        \item higher metabolic cost of search
-    \end{itemize}
-
-    \item search speed, the distance that a searching male will cover in one time step, see Fig. \ref{fig:area searched}.
-    Trade offs:
-        \begin{itemize}
-            \item higher metabolic cost of search
-            \item higher predation rate
-        \end{itemize}
-    \item maturation time, the time at which to mature is inherited as well but with a constant level of noise
+    \item r, the search radius of a searching male, See Fig. \ref{fig:area searched}.
+    \item v, the search \"velocity" of a male.
 \end{enumerate}
@@
 %% WINNERS
-
 \begin{figure}[h!]
@@
 
+%%% Linear
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=10cm,height=10cm,keepaspectratio]{figures/energy_pc_plot_exp_f15.pdf}
+    \caption{Linear commitment function, ftime = 10. Percent energy expenditure of cohort at generation 499}
+    \label{fig:var energy pc}
+    \centering
+
+    \includegraphics[width=10cm,height=10cm,keepaspectratio]{figures/energy_total_plot_exp_f15.pdf}
+    \caption{Linear commitment function, ftime = 10. Total energy expenditure of cohort at generation 499}
+    \label{fig:var energy total}
+\end{figure}
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=10cm,height=10cm,keepaspectratio]{figures/exp_f15_k_density.pdf}
+    \caption{Linear commitment function, ftime = 10. a density plot of the aggression parameter k}
+    \label{fig:k density}
+
+    \includegraphics[width=10cm,height=10cm,keepaspectratio]{figures/exp_f15_e_0_density.pdf}
+    \caption{Linear commitment function, ftime = 10. a density plot of the aggression parameter $e_0$.}
+    \label{fig:e_0 density}
+\end{figure}
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=10cm,height=10cm,keepaspectratio]{figures/exp_f15_k_history.pdf}
+    \caption{Linear commitment function, ftime = 10. $k$ the grey bounds are the standard deviation}
+
+    \includegraphics[width=10cm,height=10cm,keepaspectratio]{figures/exp_f15_e_0_history.pdf}
+    \caption{Linear commitment function, ftime = 10 $e_0$ the grey bounds are the standard deviation}
+\end{figure}
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=15cm,height=10cm,keepaspectratio]{figures/lin_f10_num_winners.pdf}
+    \caption{Linear commitment function, ftime = 10 Number of successful males at different patch areas}
+    \label{fig:num winners expf10}
+\end{figure}
+
 \clearpage
@@
     \centering
-    \begin{tabular}{c | c | l }
+    \begin{tabular}{c | l }
         \hline
-        Symbol & unit & Description\\
+        Symbol & Description\\
         \hline
         \hline
-        $K$ & & number of individual males \\
-        $N$ & & number of nests \\
-        $R_i$ & & reproductive value of nest of nest $i$\\
-        $\bar{R}$ & & average reproductive value of nest\\
+        $N$ &  The number of nests \\
+        $R$ &  The reproductive value of nest of a nest\\
+        $t_m$  & The time at which the females to mature\\
+        $A_{total}$ & The area which encloses the nests\\
         \hline
-        $t_m$ & & the time at which the females begin to mature\\
-        $S(t)$ = exp($-k_it$) & & The probability of an individual surviving to time t.\\
+        $k_{im}$ & Immature mortality coefficient \\
+        $k_e$ & The mass to energy coefficient \\
         \hline
-        $m$ & $g$ & mass\\
-        $m_mat$ & $g$ & The size at which 50\% of the population is mature.\\
-        $E$ & kJ & kiloJoules\\
+        Individual & \\
+        \hline
+        $m$ & mass\\
+        $E$ & The current energy of an individual\\
+        $r$ & An individuals search radius\\
+        $v$ & An individuals speed\\ 
         \hline
@@
 \end{table}
+
+% "patch_area":10000,
+
+% "maturation_center":4,
+% "maturation_width":1,
+ % "growth_param_a":0.5,
+% "growth_param_b":0.1,
+% "initial_mass":5,
+% "mass_to_energy":10.0,
+% "search_energy_coef":5,
+% "mutation_rate":0.1,
+% "mutation_sd":0.1
+% "stability_cutoff": 0.01,
+% "stability_lag":50
 

### 2016-02-26 ms.tex
further ms clean up, removed some plots

@@
 \end{equation}
+The nests are randomly distributed.
 
@@
 \begin{equation} \label{eq:growth}
-    \frac{{\mathrm d} m}{{\mathrm d} t} = a m^ b
+    \frac{dm}{dt} = a m^ b
 \end{equation}
@@
 \begin{equation} \label{eq:surv_immature}
-    S(t) = \exp(-k_i t).
+    S(t) = \exp(-k_{im} t).
 \end{equation}
@@
 When a male matures it is given an energy budget linearly proportional to its mass such that: 
-$$ E = k_em$$\\
+$$ E = k_em$$
+After a male matures it cannot gain more energy.
 
@@
 
-The area that a male $m$ with speed trait $v$ and radius trait $r$ can search in one time step is
+The area that a male with speed trait $v$ and radius trait $r$ can search in one time step is
 \begin{equation}
@@
 
-We then use this as the rate parameter $lambda$ to repeatedly draw the time of the next discovery event for this individual from the exponential distribution, we do this for each male from when they mature up to when the females mature.
-When an event occurs we select a nest randomly and then the male will either occupy it or contest another male for it.
+We then use this as the rate parameter $lambda$ to repeatedly draw the time of the next event for this individual from the exponential distribution, we do this for each male from when they mature up to when the females mature.
+When an event occurs we select a nest randomly and then the male will either occupy the nest or contest another male for control of the nest.
 
-When a male is kicked out of a nest we discard the old events and generate new ones.
+When a male loses a contest it is kicked out of the nest, we discard the old events and generate a list of new ones.
 
-The energy used by a male with speed $v$ and radius $r$ is 
+The energy used by a mature male male with speed $v$ and radius $r$ is 
 \begin{equation}
@@
 
-To get the cost of the fight we take the integral of $\kappa$:
+To get the cost of the fight we take the integral of the commitment function $\kappa$:
 
@@
 
-And a linear one:
-
-$$\mathrm{commitment} = \beta + \alpha \cdot (\frac{m_i - m_j}{m_i + m_j})$$
+%And a linear one:
+% $$\mathrm{commitment} = \beta + \alpha \cdot (\frac{m_i - m_j}{m_i + m_j})$$
 
@@
 % energy expenditure plots
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=10cm,height=10cm,keepaspectratio]{figures/energy_pc_plot_exp_f10.pdf}
-    \caption{Exponential commitment function, ftime = 10. Percent energy expenditure of cohort at generation 499}
-    \label{fig:energy pc}
-
-    \centering
-    \includegraphics[width=10cm,height=10cm,keepaspectratio]{figures/energy_total_plot_exp_f10.pdf}
-    \caption{Exponential commitment function, ftime = 10. Total energy expenditure of cohort at generation 499}
-    \label{fig:energy total}
-\end{figure}
-\clearpage
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=10cm,height=10cm,keepaspectratio]{figures/energy_pc_plot_exp_f15.pdf}
-    \caption{Exponential commitment function, ftime = 15. Percent energy expenditure of cohort at generation 499}
-    \label{fig:var energy pc}
-    \centering
-
-    \includegraphics[width=10cm,height=10cm,keepaspectratio]{figures/energy_total_plot_exp_f15.pdf}
-    \caption{Exponential commitment function, ftime = 15. Total energy expenditure of cohort at generation 499}
-    \label{fig:var energy total}
-\end{figure}
-
-% parameter plots density  plots
-
-% exp 10
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=10cm,height=10cm,keepaspectratio]{figures/exp_f10_k_density.pdf}
-    \caption{Exponential commitment function, ftime = 10. a density plot of the aggression parameter k}
-    \label{fig:k density}
-
-    \includegraphics[width=10cm,height=10cm,keepaspectratio]{figures/exp_f10_e_0_density.pdf}
-    \caption{Exponential commitment function, ftime = 10. a density plot of the aggression parameter $e_0$}
-    \label{fig:e_0 density}
-\end{figure}
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=10cm,height=10cm,keepaspectratio]{figures/exp_f10_k_history.pdf}
-    \caption{Exponential commitment function, ftime = 10. $k$ the grey bounds are the standard deviation}
-    \label{fig:k history}
-
-    \includegraphics[width=10cm,height=10cm,keepaspectratio]{figures/exp_f10_e_0_history.pdf}
-    \caption{ Exponential commitment function, ftime = 10. $e_0$ the grey bounds are the standard deviation}
-    \label{fig:e_0 history}
-\end{figure}
-% exp 15
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=10cm,height=10cm,keepaspectratio]{figures/exp_f15_k_density.pdf}
-    \caption{Exponential commitment function, ftime = 15. a density plot of the aggression parameter k}
-    \label{fig:k density}
-
-    \includegraphics[width=10cm,height=10cm,keepaspectratio]{figures/exp_f15_e_0_density.pdf}
-    \caption{Exponential commitment function, ftime = 15. a density plot of the aggression parameter $e_0$.}
-    \label{fig:e_0 density}
-\end{figure}
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=10cm,height=10cm,keepaspectratio]{figures/exp_f15_k_history.pdf}
-    \caption{Exponential commitment function, ftime = 15. $k$ the grey bounds are the standard deviation}
-
-    \includegraphics[width=10cm,height=10cm,keepaspectratio]{figures/exp_f15_e_0_history.pdf}
-    \caption{Exponential commitment function, ftime = 15. $e_0$ the grey bounds are the standard deviation}
-\end{figure}
-
-%% WINNERS
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=15cm,height=10cm,keepaspectratio]{figures/exp_f10_num_winners.pdf}
-    \caption{exp ftime = 10, exp commitment Number of successful males at different patch areas}
-    \label{fig:num winners expf10}
-
-    \includegraphics[width=15cm,height=10cm,keepaspectratio]{figures/exp_f15_num_winners.pdf}
-    \caption{exp ftime = 15, exp commitment. Number of successful males at different patch areas, gaps are extinctions}
-    \label{fig:num winners}
-\end{figure}
-
-%%% Linear
-
 \begin{figure}[h!]
@@
 % section results (end)results
-
-I am not sure if this is the best way to do this, the tables in the plant paper are more aesthetic.\\
-
 \begin{table}[h!]
@@
         $k_{im}$ & Immature mortality coefficient \\
-        $k_e$ & The mass to energy coefficient \\
+        $k_e$ & The initial mass to energy coefficient \\
+        $k_{search}$ & The metabolic cost of searching \\
         \hline
@@
         \hline
+        mutation rate & The chance of a trait being mutated\\
+        mutation sd & The standard deviation of  these mutations\\
+        \hline
     \end{tabular}
 \end{table}
-
-% "patch_area":10000,
 

### 2016-02-29 scrap/evolution.Rmd
old data cleanup and scrap changes

@@
 output:
+  html_document: default
   pdf_document:
@@
     fig_height: 4
-  html_document: default
 ---
@@
 linf10_pop <- load_population("../data/lin_f10/")
+exp10_1000_pop <- load_population("../data/exp_f10_1000/")
 population_plots <- function (y, x){

### 2016-02-29 scrap/generation_comparison.Rmd
old data cleanup and scrap changes

---
title: "A Comparison of Different Generation lengths"
author: "John"
date: "19 February 2016"
output: pdf_document
---
I ran one simulation stepping by 100 for 500 generations.
and another stepping by 250 for 2000 generations.

```{r}
library(jsonlite)
library(plyr)
library(ggplot2)
source("~/Documents/contests_model/R/load_datasets.R")
source("~/Documents/contests_model/R/make_plots.R")
cm_100 <- load_population("~/Documents/contests_model/data/cm_100/")
cm_2k <- load_population("~/Documents/contests_model/data/cm_2k/")

# get upper and lower bounds of trait values
get_limits <- function (what, dfs){
  return(c(min(dfs[what]), max(dfs[what])))
}

population_plots <- function (x, y, limits){
  print(
    pop_values(cm_100, x = x, y = y, paste("cm_100",x, y)) + 
      scale_y_continuous(limits = limits)
  )
  print(
    pop_values(cm_2k, x = x, y = y, paste("cm_2k",x, y)) + 
      scale_y_continuous(limits = limits)
  )
}
e_0_limits <- get_limits("e_0", rbind(cm_2k, cm_100))
k_limits <- get_limits("k", rbind(cm_2k, cm_100))
```

There seems to be coexistence of strategies after a certain density is reached.
Where to next?

Test different parameters maybe?
I am running it with lower a lower logit coeff for the winner swap...

How to limit the aggression drift when training on a lot of generations?
Different commitment functions?
Different escalation curves?

```{r}
population_plots("log(N / patch_area)", "e_0", e_0_limits)
population_plots("log(N / patch_area)", "k", k_limits)
population_plots("patch_area", "e_0", e_0_limits)
population_plots("patch_area", "k", k_limits)
```

### 2016-02-29 R/make_plots.R

@@ -1,8 +1,8 @@
 # what is what we are plotting the density of
 # returns a ggplot2 object
-density_plot <- function (data, what, title) {
+density_plot <- function (data, x = "patch_area", y, title) {
   return(
-    ggplot(data, aes_string(x = "patch_area", y = what)) + 
+    ggplot(data, aes_string(x = x, y = y)) + 
     stat_density2d(aes(fill = ..density.. ), geom = "tile", contour = FALSE) + 
     ggtitle(title)
   )
@@ -129,8 +129,55 @@ trait_evolution_plot <- function(df, what, title){
 pop_values <- function (df, x, y, title){
   ggplot(df, aes_string(x = x, y = y)) + 
     geom_point() +
-    geom_smooth(method=lm) +
     ggtitle(title)
 }
 
+
+
+# visualise how contest traits change with increasing density
+# takes a population (not a history)
+# creates a gif at the filepath given
+# with the given title string
+# if delete pngs it cleans up the contest_gif directory
+trait_scatter_gif <- function(dataset, filepath = "../scrap/contest_gifs/", 
+                              progress = TRUE,
+                              plot_title = "my_gif",
+                              delete_pngs = FALSE){
+  
+  # get the limits
+  k_limits = c(min(dataset$k), max(dataset$k))
+  e_0_limits = c(min(dataset$e_0), max(dataset$e_0))
+  
+  # create the pngs
+  # dlply split dataframe return list of dataframes
+  for (simulation in dlply(dataset, .(patch_area))){
+    filename <- sprintf("%06d_%s.png", simulation$patch_area[1], plot_title)
+    filename <- paste(filepath, filename, sep = "")
+    if(progress) print(filename)
+    
+    png(filename = filename)
+    print(
+      ggplot(simulation, aes(x = k, y = e_0)) +
+        geom_point() +
+        scale_x_continuous(limits=k_limits) +
+        scale_y_continuous(limits=e_0_limits) +
+        ggtitle(paste(plot_title, "Patch Area", simulation$patch_area[1], 
+                      "Density: ", log(simulation$N[1] / simulation$patch_area[1])))
+    )
+    dev.off()
+  }
+  
+  # this makes the gifs
+  # I am not sure if this will work on mac...
+  command <- paste("convert -delay 20 -loop 0 ", filepath,"*.png ", filepath, plot_title,".gif", sep = "")
+  if (progress) print(command)
[…]

### 2016-02-29 python/simulation.py

@@ -55,7 +55,8 @@ def main():
         # type conversion
         params[p] = type(params[p])(sys.argv[2])
 
-    np.random.seed(seed=params["random_seed"])
+    if params["random_seed"] != 0:
+        np.random.seed(seed=params["random_seed"])
     
     sim = Simulation(params)
     for i in range(params["max_gen"]):

### 2016-03-14 ms.tex
old data cleanup and scrap changes
results section fixed

@@
 \newcommand{\smurl}[1]{{\footnotesize\url{#1}}}
+
+\newcommand{\myfigure}[4]{
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=10cm,height=9cm,keepaspectratio]{#1}
+    \caption{#2}
+    \centering
+    \includegraphics[width=10cm,height=9cm,keepaspectratio]{#3}
+    \caption{#4}
+\end{figure}
+}
+
 \usepackage{graphicx}
@@
 
-This means adding some normally distributed noise with $\mu = 0$ and $\sigma$ varying from trait to trait.
+This means a chance to add some normally distributed noise with $\mu = 0$ and small $\sigma$.
 
@@
 \label{sec:results}
-% energy expenditure plots
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=10cm,height=10cm,keepaspectratio]{figures/energy_pc_plot_exp_f15.pdf}
-    \caption{Linear commitment function, ftime = 10. Percent energy expenditure of cohort at generation 499}
-    \label{fig:var energy pc}
-    \centering
 
-    \includegraphics[width=10cm,height=10cm,keepaspectratio]{figures/energy_total_plot_exp_f15.pdf}
-    \caption{Linear commitment function, ftime = 10. Total energy expenditure of cohort at generation 499}
-    \label{fig:var energy total}
-\end{figure}
+\myfigure{figures/energy_pc_plot_cm_100.pdf}{pc energy spent on what}{figures/energy_total_plot_cm_100.pdf}{total energy spent on what}
+
+\myfigure{figures/e_0_nest_density.pdf}{e 0 with patch area}{figures/k_patch_area.pdf}{k with patch area}
+
+\myfigure{figures/e_0_nest_density.pdf}{e 0 with nest density}{figures/k_patch_area.pdf}{k with nest density}
 
@@
     \centering
-    \includegraphics[width=10cm,height=10cm,keepaspectratio]{figures/exp_f15_k_density.pdf}
-    \caption{Linear commitment function, ftime = 10. a density plot of the aggression parameter k}
-    \label{fig:k density}
-
-    \includegraphics[width=10cm,height=10cm,keepaspectratio]{figures/exp_f15_e_0_density.pdf}
-    \caption{Linear commitment function, ftime = 10. a density plot of the aggression parameter $e_0$.}
-    \label{fig:e_0 density}
-\end{figure}
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=10cm,height=10cm,keepaspectratio]{figures/exp_f15_k_history.pdf}
-    \caption{Linear commitment function, ftime = 10. $k$ the grey bounds are the standard deviation}
-
-    \includegraphics[width=10cm,height=10cm,keepaspectratio]{figures/exp_f15_e_0_history.pdf}
-    \caption{Linear commitment function, ftime = 10 $e_0$ the grey bounds are the standard deviation}
-\end{figure}
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=15cm,height=10cm,keepaspectratio]{figures/lin_f10_num_winners.pdf}
-    \caption{Linear commitment function, ftime = 10 Number of successful males at different patch areas}
-    \label{fig:num winners expf10}
+    \includegraphics[width=15cm,height=10cm,keepaspectratio]{figures/cm_100_num_winners.pdf}
+    \caption{Number of successful males at different patch areas}
 \end{figure}
@@
 
-% "maturation_center":4,
-% "maturation_width":1,
-% "growth_param_a":0.5,
-% "growth_param_b":0.1,
-% "initial_mass":5,
-% "mass_to_energy":10.0,
-% "search_energy_coef":5,
-% "mutation_rate":0.1,
-% "mutation_sd":0.1
-% "stability_cutoff": 0.01,
-% "stability_lag":50
-
-\section{Figures}
-
-\begin{figure}[h!]
-\centering
-\includegraphics[width=15cm,height=15cm,keepaspectratio]{figures/area_searched}
-\caption{the area searched by a male in interval $dt$}
-\label{fig:area searched figure}
-\end{figure}
-\clearpage
-
 \bibliography{refs}

### 2016-03-14 R/make_plots.R

@@ -63,6 +63,7 @@ as_stackable_total <- function (m){
 }
 
 # returns a ggplot2 object
+# this plot doesnt show the multiple clusters
 history_plot <- function(df, what, bounds,  title) {
   return(
     ggplot(data = df, aes(patch_area)) +
@@ -77,20 +78,19 @@ history_plot <- function(df, what, bounds,  title) {
   )
 }
 
-
 # returns a ggplot2 object
-energy_stack_plot <- function (df, title){
+energy_stack_plot <- function (df, title, what="patch_area"){
   return(
-    ggplot(df, aes(patch_area, energy)) + 
+    ggplot(df, aes_string(x = what, y = "energy")) + 
       geom_area(aes(fill = type)) + 
       ggtitle(title)
   )
 }
 
 # wrapper for a % energy plot
-energy_pc_plot <- function (m){
+energy_pc_plot <- function (m, what="patch_area"){
   energy_stack_plot(as_stackable_pc(m), 
-    "Energy spending across population")
+    "Energy spending across population", what)
 }
 
 # wrapper for a less "full plot that shows the decline in energy"
@@ -131,7 +131,6 @@ pop_values <- function (df, x, y, title){
     geom_point() +
     ggtitle(title)
 }
-
 
 
 # visualise how contest traits change with increasing density
@@ -177,7 +176,7 @@ trait_scatter_gif <- function(dataset, filepath = "../scrap/contest_gifs/",
   # THIS COULD FUCK UP YOUR SHIT IF YOUR PLOT TITLE HAS A " " in it
   command <- paste("rm ", filepath, "*", plot_title, ".png", sep = "")
   if (progress) print(command)
-  system(command)
+  if(delete_pngs) system(command)
   
 }
 

### 2016-03-15 ms.tex
clustering plots, ms changes

@@
     \centering
-    \includegraphics[width=10cm,height=9cm,keepaspectratio]{#1}
+    \includegraphics[width=13cm,height=10cm]{#1}
     \caption{#2}
     \centering
-    \includegraphics[width=10cm,height=9cm,keepaspectratio]{#3}
+    \includegraphics[width=13cm,height=10cm]{#3}
     \caption{#4}
@@
 
+% whats b? define/ state value
 Within the immature phase, we assume males grow with rate
@@
 
+% needs more detail, need location and width parameter values
+
 We Pull the maturation time of a male randomly from the logistic distribution with location and width parameters.
 
+% state k_im value and example probability
 If we assume a constant mortality rate $k_{im}$ during the immature phase, the probability of a male surviving to time $t$ from $0$ is given by
@@
 $$ E = k_em$$
-After a male matures it cannot gain more energy.
+After a male matures it cannot gain more energy. When a mature male uses its energy allocation it dies.
 
@@
 \end{equation}
+A male is 
 See Fig. \ref{fig:area searched}.
 
-We let the total area enclosing all males and all nests be a parameter, $A_{total}$. We assume spatially random nests.
+The energy used by a searching male with speed $v$ and radius $r$ is 
+\begin{equation}
+    E(r,v) = k_{search} \cdot r \cdot v
+\end{equation}
+where $k_{search}$ is a constant.  % state value
 
-The encounter rate $\varepsilon$ for an individual is:
+We let the total area enclosing all nests be a parameter, $A_{total}$.
+$N$ is the number of nests enclosed. $\frac{N}{A_{total}}$ is the nest density.
+
+The encounter rate $\lambda$ for an individual is:
 \begin{equation} \label{eq:encounter rate}
-    \varepsilon = \frac {N \cdot 2 \cdot r \cdot v} {A_{total}}
+    \lambda = \frac {N \cdot 2 \cdot r \cdot v} {A_{total}}
 \end{equation}
 \citep{Gurarie2012}
-
-We then use this as the rate parameter $lambda$ to repeatedly draw the time of the next event for this individual from the exponential distribution, we do this for each male from when they mature up to when the females mature.
+We assume spatially random nests so nest discovery is a poisson process with the the expected number of encounters in one time step being $\lambda$.
+Because of this we can then use this as the rate parameter to repeatedly draw the time of the next event for this individual from the exponential distribution.
+We do this for each male from when they mature up to when the females mature.
 When an event occurs we select a nest randomly and then the male will either occupy the nest or contest another male for control of the nest.
@@
 
-The energy used by a mature male male with speed $v$ and radius $r$ is 
-\begin{equation}
-    E(r,v) = k_{search} \cdot r \cdot v
-\end{equation}
-where $k_{search}$ is a constant.
-
 \clearpage
 
-\subsection{Contest Algorithm}
-Contests 
-When a searching male encounters a nest that is occupied a contest over possession of that nest occurs. Contests occur instantaneously.
+\subsection{Contests}
+When a searching male encounters a nest that is occupied a contest over possession of that nest occurs. Contests occur instantaneously, taking up .
 
@@
 
-$$Energy Spent = \int_0^{c_i}{\kappa(t)}dt$$
+$$Energy Spent = \int_0^{c_i}{\kappa(t)}dt $$
 
-if the $\kappa$ is $commitment^2$ then
+if the $\kappa(c)$ is $c^2$ then
 $$EnergySpent = \frac{c_i^3}{3}$$
@@
 
-The male chooses how much energy to commit based off of a commitment function, we examine two different commitment functions, for a male $i$ against a male $j$
+The male chooses how much energy to commit based off of a commitment function. for a male $i$ against a male $j$
 
@@
 $$\mathrm{commitment} = e^\beta \cdot (\frac{m_i}{m_j})^\alpha $$
-
-%And a linear one:
-% $$\mathrm{commitment} = \beta + \alpha \cdot (\frac{m_i - m_j}{m_i + m_j})$$
 
@@
 
-\myfigure{figures/energy_pc_plot_cm_100.pdf}{pc energy spent on what}{figures/energy_total_plot_cm_100.pdf}{total energy spent on what}
+\myfigure{figures/energy_pc_plot_cm_100.png}{pc energy spent on what}{figures/energy_total_plot_cm_100.png}{total energy spent on what}
 
-\myfigure{figures/e_0_nest_density.pdf}{e 0 with patch area}{figures/k_patch_area.pdf}{k with patch area}
-
-\myfigure{figures/e_0_nest_density.pdf}{e 0 with nest density}{figures/k_patch_area.pdf}{k with nest density}
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=15cm,height=10cm,keepaspectratio]{figures/cm_100_num_winners.pdf}
-    \caption{Number of successful males at different patch areas}
-\end{figure}
+\myfigure{figures/e_0_patch_area.png}{e 0 with patch area}{figures/clustered_e_0_plot.pdf}{an equivilant plot with k means clustering done, with the k parameter chosen by $pamk$, the number of clusters (coexisting strategies) can be be seen on the left}
 

### 2016-03-15 scrap/clustering.rmd
clustering plots, ms changes

---
title: "Clustering"
author: "John Wilshire"
date: "15 March 2016"
output: pdf_document
---

```{r, echo=FALSE}
library(fpc)

library(jsonlite)
library(plyr)
library(ggplot2)
source("~/Documents/contests_model/R/load_datasets.R")
source("~/Documents/contests_model/R/make_plots.R")
```

the pamk function finds the ideal k for k means.

```{R}

exp_100 <- load_population("~/Documents/contests_model/data/cm_100/")

exp_1000 <- exp_100[exp_100$patch_area %in% seq(1000,30000, by = 500),]

pamk_clusters <- ddply(exp_1000, .variables = .(patch_area), 
          .fun = function (df){
            m <- pamk(df$e_0, krange = 1:3, alpha = 1e-12)
            cluster <- m$pamobject$clustering
            mediod <- m$pamobject$medoids[cluster]
            distance <- m$pamobject$clusinfo[cluster, "diameter"]
            nc <- max(cluster)
            unique(data.frame(cluster, mediod, distance, nc))
          })

ggplot(pamk_clusters, aes(x = patch_area, y = mediod)) +
  geom_errorbar(aes( ymin = mediod - (distance/2), ymax = mediod + (distance/2))) +
  geom_point(aes(col=as.factor(nc))) + 
  ylab("log(beta)") + 
  ggtitle("Clustered using pamk") + 
  scale_colour_discrete(name="Number of Clusters")

qplot(data = exp_1000, maturation_time, 
      fill=as.factor(patch_area), 
      main ="maturation times of winners")

qplot(data = exp_1000, mass, fill=as.factor(patch_area), main ="mass distrobution of winners, mass center of inital population is 4")

qplot(data = exp_1000, x = patch_area, y = mass)

```

### 2016-03-15 R/make_plots.R

@@ -64,6 +64,7 @@ as_stackable_total <- function (m){
 
 # returns a ggplot2 object
 # this plot doesnt show the multiple clusters
+# not very good as it doesnt show co-existence
 history_plot <- function(df, what, bounds,  title) {
   return(
     ggplot(data = df, aes(patch_area)) +
@@ -99,19 +100,27 @@ energy_total_plot <- function(m){
       "Energy spending across population")
 }
 
-# takes 1 generation of multiple patch areas
-# ie all generation 300
-k_history_plot <- function (m){
-  history_plot(m, "mean_k", "std_k", "Aggression, parameter K")
-}
 
-e_0_history_plot <- function (m){
-  history_plot(m, "mean_e_0", "std_e_0", "Aggression, parameter e_0")
-}
-
-# takes 1 generation
-exploration_history_plot <- function (m){
-  history_plot(m, "mean_exploration", "std_exploration", "exploration traits")
+# plyr is so cool
+# split on patch area, apply culstering, reapply
+clustered_e_0_plot <- function(dff, select_every = 200, title ="Coexistance"){
+  pamk_clusters <- ddply(dff[dff$patch_area %in% seq(min(dff$patch_area), max(dff$patch_area), by = select_every), ],
+                         .variables = .(patch_area), 
+                         .fun = function (df){
+                           m <- pamk(df$e_0, krange = 1:3, alpha = 1e-12)
+                           cluster <- m$pamobject$clustering
+                           mediod <- m$pamobject$medoids[cluster]
+                           distance <- m$pamobject$clusinfo[cluster, "diameter"]
+                           nc <- max(cluster)
+                           unique(data.frame(cluster, mediod, distance, nc))
+                         })
+  
+  ggplot(pamk_clusters, aes(x = patch_area, y = mediod)) +
+    geom_errorbar(aes( ymin = mediod - (distance/2), ymax = mediod + (distance/2))) +
+    geom_point(aes(col=as.factor(nc))) + 
+    ylab("beta") + 
+    ggtitle(title) + 
+    scale_colour_discrete(name="Number of Clusters")
 }
 
 trait_evolution_plot <- function(df, what, title){

### 2016-03-18 ms.tex
logging of fights

@@
 
+We use $F = 20$ and $N = 500$
+
 \subsection{Immature phase}
 
-% whats b? define/ state value
 Within the immature phase, we assume males grow with rate
@@
 
-% needs more detail, need location and width parameter values
+We use $a = 0.5$ and $b = 0.1$ and $m(0) = 5$
 
-We Pull the maturation time of a male randomly from the logistic distribution with location and width parameters.
+We Pull the maturation time of a male randomly from the logistic distribution with center $= 4$ and width $= 1$.
 
-% state k_im value and example probability
 If we assume a constant mortality rate $k_{im}$ during the immature phase, the probability of a male surviving to time $t$ from $0$ is given by
@@
 
+We use $k_{im} = 0.2$
+
+For example the probability of a male surviving to maturity from to time = $5$ is $exp(-0.2 \cdot 5)= 0.3678$.
+
 When a male matures it is given an energy budget linearly proportional to its mass such that: 
 $$ E = k_em$$
+
+We use $k_e = 10$
+
 After a male matures it cannot gain more energy. When a mature male uses its energy allocation it dies.
@@
 \end{equation}
-where $k_{search}$ is a constant.  % state value
+where $k_{search} = 5$ is a constant.  % state value
 
@@
 \citep{Gurarie2012}
+
 We assume spatially random nests so nest discovery is a poisson process with the the expected number of encounters in one time step being $\lambda$.
-Because of this we can then use this as the rate parameter to repeatedly draw the time of the next event for this individual from the exponential distribution.
-We do this for each male from when they mature up to when the females mature.
+Because of this we can then use $\lambda$ as the rate parameter to repeatedly draw the time of the next event for this individual from the exponential distribution.
+
+We do this for each male from when they mature up to when the females mature (at $\text{time} = 10$).
 When an event occurs we select a nest randomly and then the male will either occupy the nest or contest another male for control of the nest.
@@
 \subsection{Contests}
-When a searching male encounters a nest that is occupied a contest over possession of that nest occurs. Contests occur instantaneously, taking up .
+When a searching male encounters a nest that is occupied a contest over possession of that nest occurs.
 
-The costs of a contest escalates with more commitment.
+Contests occur instantaneously, taking up no time, this means that another male cannot encounter the nest whilst the contest is taking place, additionally the contest has no intrinsic metabolic cost.
 
-We will assume an escalation curve $kappa$ such that:
-$$\kappa : Commitment \rightarrow Intensity$$
-where $Intensity$ is $\frac{Energy}{Commitment}$
+At the start of the contest each male chooses how much  energy they are willing to commit to this contest, they have full information about their opponents mass.
 
-We use the escalation curve $$ \kappa (c) = c^2 $$
+They do this by reporting their level of commitment as a positive value $c$.
 
-At the start of the contest each male chooses how much  energy they are willing to commit.
-They do this by reporting their level of ``commitment'' $c_i$.
+The intensity of a fight increases with more commitment:
 
-The male that chooses to commit more of its energy to the contest will win with probability $= logit^{-1}(k(c_i - c_j))$.
-Where k is is a parameter that shapes how sensitive the contest is to small differences in commitment values.
+We define $\text{Intensity} = \frac{\text{d}Energy}{\text{d}Commitment}$
 
-The energy that this male chose to commit will then be deducted from both males with the loser continuing to search for mates
+We use an intensity curve $I(c) = c^2$
 
-To get the cost of the fight we take the integral of the commitment function $\kappa$:
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=10cm,height=8cm]{figures/example_escalation.png}
+    \caption{The energy spent in the fight between male i and male j is the pink shaded area.}
+\end{figure}
 
-$$Energy Spent = \int_0^{c_i}{\kappa(t)}dt $$
+To get the cost of the fight we take the integral of the intensity function:
+$$E_\text{cost}(c) = \int_0^{c}{I(t)}dt = \int_0^{c}{t^2}dt $$
 
-if the $\kappa(c)$ is $c^2$ then
-$$EnergySpent = \frac{c_i^3}{3}$$
-Where $c_i$ is the losing male.
+The lower energy cost ( corrosponding to $c_i$ in the diagram) will then be deducted from both males.
 
-The male chooses how much energy to commit based off of a commitment function. for a male $i$ against a male $j$
+Male $i$ will loose the contest with probability 
+$$P(\text{})= logit^{-1}(k(c_j - c_i))$$
 
-An exponential one:
-$$\mathrm{commitment} = e^\beta \cdot (\frac{m_i}{m_j})^\alpha $$
+Where k is is a parameter that shapes how sensitive the contest is to small differences in commitment values (we use $k = 10$).
+can lower k be thought of as perception error (?? mike).
+
+The male chooses how much energy to commit based off of a commitment function. For a male $i$ against a male $j$
+
+$$\text{commitment} = e^\beta \cdot (\frac{m_i}{m_j})^\alpha $$
 
@@
 $$\alpha , \beta$$
-Are under selection.
+are under selection.
 
@@
 \begin{enumerate}
-    \item aggression 0 ($\beta$) called $e_0$
-    \item aggression 1 ($\alpha$) called $k$
-
-    \item r, the search radius of a searching male, See Fig. \ref{fig:area searched}.
-    \item v, the search \"velocity" of a male.
+    \item $\beta$
+    \item $\alpha$
+    \item $r$, the search radius of a searching male, See Fig. \ref{fig:area searched}.
+    \item $v$, the search speed of a male.
 \end{enumerate}
 
-\clearpage
+The initial distribution of $\alpha$ and $\beta$ is $N(0, 3)$.
+
+The initial distribution of $v$ and $r$ is $N(2, 0.5)$.
+
+Where $N(\mu, \sigma)$, is the normal distribution with mean $= \mu$ and standard deviation $\sigma$.
 
@@
 
-\myfigure{figures/e_0_patch_area.png}{e 0 with patch area}{figures/clustered_e_0_plot.pdf}{an equivilant plot with k means clustering done, with the k parameter chosen by $pamk$, the number of clusters (coexisting strategies) can be be seen on the left}
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=15cm,height=10cm]{figures/clustered_e_0_plot.pdf}
+    \caption{A plot of the distribution of the aggression parameter $beta$ over varing $A_{total}$ with k means clustering done, with the k parameter chosen by $pamk$, the number of clusters (coexisting strategies) can be be seen on the right}
+\end{figure}
 

### 2016-03-18 scrap/contest_data.Rmd
logging of fights

---
title: "actual contest data"
author: "John Wilshire"
date: "18 March 2016"
output: html_document
---

```{R, echo = FALSE}
library(jsonlite)
library(plyr)
library(ggplot2)
source("~/Documents/contests_model/R/load_datasets.R")
source("~/Documents/contests_model/R/make_plots.R")

contests <- data.frame(fromJSON("~/Documents/contests_model/data/03_18_05_08_07__c7a73359-a904-4c5c-ba5b-189dff605a90.json")[c("contests", "parameters")])
colnames(contests) <- sub("^[^.]*.","", colnames(contests))
```

Seems pretty reasonable...
```{R} 
ggplot(contests, aes(x = def_mass, y = atk_mass)) + 
  geom_point(aes(col=contests$defence_winner)) + 
  ylab("mass of the attacker") + 
  xlab("mass of the defender") + 
  scale_colour_discrete(name="Defender Wins")
```

```{R}
ggplot(contests, aes(y = atk_start_energy , x = def_start_energy )) + 
  geom_point(aes(col=contests$defence_winner)) + 
  ylab("initial energy of the attacker") + 
  xlab("initial energy of the defender") + 
  scale_colour_discrete(name="Defender Wins")

```

It looks like a lot of the contests end in death
```{R}
ggplot(contests, aes(y = atk_end_energy , x = def_end_energy )) + 
  geom_point(aes(col=contests$defence_winner)) + 
  ylab("final energy of the attacker") + 
  xlab("final energy of the defender") + 
  scale_colour_discrete(name="Defender Wins")

```

The intensity function is $c^2$
```{R}
ggplot(contests, aes(x = def_commit, y = atk_commit)) + 
  geom_point(aes(col=contests$defence_winner)) + 
  ylab("commitment of the attacker") + 
  xlab("commitment of the defender") + 
  scale_colour_discrete(name="Defender Wins")

ggplot(contests, aes(x = def_beta, y = atk_beta)) + 
  geom_point(aes(col=contests$defence_winner)) + 
  ylab("beta trait of the attacker") + 
  xlab("beta trait of the defender") + 
  scale_colour_discrete(name="Defender Wins")

ggplot(contests, aes(x = def_beta, y = atk_beta)) + 
  geom_point(aes(col=contests$defence_winner)) + 
  ylab("beta trait of the attacker") + 
  xlab("beta trait of the defender") + 
  scale_colour_discrete(name="Defender Wins")

ggplot(contests, aes(x = def_alpha, y = atk_alpha)) + 
  geom_point(aes(col=contests$defence_winner)) + 
  ylab("alpha trait of the attacker") + 
  xlab("alpha trait of the defender") + 
  scale_colour_discrete(name="Defender Wins")

qplot(contests$fight_cost, fill= contests$defence_winner)

```

### 2016-03-18 R/make_plots.R

@@ -45,6 +45,30 @@ get_generation <- function(m, gen){
   return(final)
 }
 
+commitment <- seq(0,5, by = 0.01)
+
+escalation_curve <- function (t){
+  t^2
+}
+
+example_escalation <- function (){
+  commitment <- seq(0,5, by = 0.01)
+  escalation_curve <- function (t){
+    t^2
+  }
+  df <- data.frame("Commitment" = commitment, "Intensity" = escalation_curve(commitment))
+  ggplot(df,  aes( x = Commitment, y = Intensity )) +
+    geom_line() +
+    geom_vline(xintercept = 3) + 
+    geom_vline(xintercept = 4) +
+    geom_polygon(data = rbind(c(0,0), subset(df, commitment < 3), c(3, 0)), 
+                 aes(alpha = 0.3, fill = "")) + 
+    guides(fill = FALSE, alpha = FALSE) + 
+    annotate("text", x = 2.8, y = 10, label = "c_ i") +
+    annotate("text", x = 3.8, y = 20, label = "c_ j")
+}
+
+
 # returns a dataframe that can be used to make plots that show the % energy spending
 as_stackable_pc <- function (m){
   contest <- data.frame("patch_area" = m$patch_area, "energy" = m$contest_as_pc, "type" = "contest")
@@ -103,11 +127,11 @@ energy_total_plot <- function(m){
 
 # plyr is so cool
 # split on patch area, apply culstering, reapply
-clustered_e_0_plot <- function(dff, select_every = 200, title ="Coexistance"){
+clustered_plot <- function(dff, select_every = 200, title ="Coexistance", what, ylab){
   pamk_clusters <- ddply(dff[dff$patch_area %in% seq(min(dff$patch_area), max(dff$patch_area), by = select_every), ],
                          .variables = .(patch_area), 
                          .fun = function (df){
-                           m <- pamk(df$e_0, krange = 1:3, alpha = 1e-12)
+                           m <- pamk(df[what], krange = 1:3, alpha = 1e-12)
                            cluster <- m$pamobject$clustering
                            mediod <- m$pamobject$medoids[cluster]
                            distance <- m$pamobject$clusinfo[cluster, "diameter"]
@@ -118,9 +142,9 @@ clustered_e_0_plot <- function(dff, select_every = 200, title ="Coexistance"){
   ggplot(pamk_clusters, aes(x = patch_area, y = mediod)) +
[…]

### 2016-03-18 python/generation_logger.py

@@ -4,9 +4,6 @@ import numpy as np
 import sys
 import matplotlib.pyplot as plt
 
-# TODO rename to generation logger
-# TODO make energy time series use a %
-
 class GenerationLogger(object):
     """docstring for logs"""
     def __init__(self, generation):
@@ -35,7 +32,27 @@ class GenerationLogger(object):
             "occupying_energy":[0]
         }
 
-    # this function should be called each time step
+        # records information about the contests that happen in a generation
+        self.contest_data = {
+            "fight_cost":[],
+            "def_start_energy":[],
+            "atk_start_energy":[],
+            "def_end_energy":[],
+            "atk_end_energy":[],
+            "def_commit":[],
+            "atk_commit":[],
+            "atk_mass":[],
+            "def_mass":[],
+            "prob_upset":[],
+            "defence_winner":[],
+            "atk_alpha":[],
+            "atk_alpha":[],
+            "atk_beta":[],
+            "def_alpha":[],
+            "def_beta":[]
+        }
+
+    # this function should be called each time an event is processed
     def log_cohort(self):
         self.data["time"].append(self.generation.time)
         self.data["searching"].append(self.searching)
@@ -47,6 +64,11 @@ class GenerationLogger(object):
         self.data["search_energy"].append(self.search_energy)      
         self.data["contest_energy"].append(self.contest_energy)        
         self.data["occupying_energy"].append(self.occupying_energy)
+
+    def log_contest(self, contest_info_dict):
+        for key in contest_info_dict:
+            self.contest_data[key].append(contest_info_dict[key])
+
 
     def get_col(self, col_name):
         return [ row[col_name] for row in self.data ]

### 2016-03-18 python/nest.py

@@ -60,6 +60,7 @@ class Nest(object):
                 attacker_commitment, cost, defender.energy, attacker.energy, prob_upset, str(winner == defender))
         # make sure no more than the max energy of either male can be deducted
         cost = min([winner.energy, loser.energy, cost])
+
         
         # the costs are deducted
         attacker.energy -= cost
@@ -76,6 +77,25 @@ class Nest(object):
             self.eject()
             self.occupy(attacker)
 
+        defender.logger.log_contest(
+            {
+                "fight_cost":cost,
+                "def_start_energy":defender.energy + cost,
+                "atk_start_energy":attacker.energy + cost,
+                "def_end_energy":defender.energy,
+                "atk_end_energy":attacker.energy,
+                "def_commit":defender_commitment,
+                "atk_commit":attacker_commitment,
+                "atk_mass":attacker.mass,
+                "def_mass":defender.mass,
+                "prob_upset":prob_upset,
+                "defence_winner": winner == defender,
+                "atk_alpha":attacker.k,
+                "atk_beta":attacker.e_0,
+                "def_alpha":defender.k,
+                "def_beta":defender.e_0
+            }
+        )
         return loser
 
     # the cost of a fight that has escalated from 0 to the commitment given

### 2016-03-18 python/simulation_logger.py

@@ -91,7 +91,8 @@ class SimulationLogger (object):
         jsons = {
             "parameters":self.params,
             "traits":traits,
-            "history":self.data
+            "history":self.data,
+            "contests":self.simulation.generations[-1].logger.contest_data
         }
 
         file_name = "../data/" + strftime("%m_%d_%H_%M_%S__", gmtime()) + str(uuid.uuid4()) + ".json"
@@ -107,20 +108,20 @@ class SimulationLogger (object):
             lag_length = self.params["stability_lag"]
             # finds the max sd of the traits in the last 'lag' generations
 
-            deltas = lambda l : [ abs(l[i] - l[i - 1]) 
-                                    for i in range(len(l) - lag_length - 1, len(l) -1)
-                                ]
+            delta = lambda x: max(x[ -lag_length:]) - min(x[ -lag_length:])
 
-            max_delta_sd = max( 
-                max( deltas( self.data["std_e_0"] )),
-                max( deltas( self.data["std_k"] )),
-                max( deltas( self.data["std_exploration"] ))
+            max_delta = max( 
+                delta(self.data["std_e_0"]),
+                delta(self.data["mean_e_0"]),
+                delta(self.data["std_k"]),
+                delta(self.data["mean_k"]),
+                delta(self.data["mean_exploration"])
             )
 
             if not self.params["debug"]:
-                print "max delta sd: %s" % max_delta_sd
+                print "max delta: %s" % max_delta
 
-            if max_delta_sd < self.params["stability_cutoff"]:
+            if max_delta < self.params["stability_cutoff"]:
                 return True
         return False
         

### 2016-03-23 scrap/clustering.rmd
r and v can be fixed

@@
 
+cm_6k <- load_population("~/Documents/contests_model/data/cm_100/")
+clustered_plot(cm_6k, what = "e_0", ylab = "beta")
+
 exp_1000 <- exp_100[exp_100$patch_area %in% seq(1000,30000, by = 500),]
 
-pamk_clusters <- ddply(exp_1000, .variables = .(patch_area), 
+pamk_clusters <- ddply(exp_100, .variables = .(patch_area), 
           .fun = function (df){
-            m <- pamk(df$e_0, krange = 1:3, alpha = 1e-12)
-            cluster <- m$pamobject$clustering
-            mediod <- m$pamobject$medoids[cluster]
-            distance <- m$pamobject$clusinfo[cluster, "diameter"]
-            nc <- max(cluster)
-            unique(data.frame(cluster, mediod, distance, nc))
+            c1 <- pamk(df$k, krange = 1:3, alpha = 1e-12)
+            k_cluster <- c1$pamobject$clustering
+            k_mediod <- c1$pamobject$medoids[k_cluster]
+            k_distance <- c1$pamobject$clusinfo[k_cluster, "diameter"]
+            k_nc <- max(k_cluster)
+
+            c2 <- pamk(df$e_0, krange = 1:3, alpha = 1e-12)
+            e_0_cluster <- c2$pamobject$clustering
+            e_0_mediod <- c2$pamobject$medoids[e_0_cluster]
+            e_0_distance <- c2$pamobject$clusinfo[e_0_cluster, "diameter"]
+            e_0_nc <- max(e_0_cluster)
+
+            data.frame(k_cluster, k_mediod, k_distance, k_nc, e_0_cluster, e_0_mediod, e_0_distance, e_0_nc)
           })
+
+ggplot(data = pamk_clusters, aes(x = patch_area)) + 
+  geom_point(aes(y = k_nc, col = as.factor(e_0_nc)))
+
+pamk_1000 <- pamk_clusters[pamk_clusters$patch_area == 1000, ]
+
+ggplot(data = exp_100[exp_100$patch_area == 1000, ], aes(x = k)) + 
+  geom_point(aes(y = e_0))
+  geom_polygon
 
@@
   geom_point(aes(col=as.factor(nc))) + 
-  ylab("log(beta)") + 
+  ylab("beta") + 
   ggtitle("Clustered using pamk") + 
@@
 
-qplot(data = exp_1000, mass, fill=as.factor(patch_area), main ="mass distrobution of winners, mass center of inital population is 4")
-
-qplot(data = exp_1000, x = patch_area, y = mass)
+pamk(exp_100[exp_100$patch_area == 4600,]$k, krange = 1:3, alpha = 1e-12)
 

### 2016-03-23 scrap/contests_accross_patch_areas.rmd
r and v can be fixed

---
title: "contests accross patch areas"
author: "John Wilshire"
date: "21 March 2016"
output: pdf_document
---

contests_accross_patch_areas.rmd

6k generations, mutation sd 0.05, rate 0.1

```{R}
library(jsonlite)
library(plyr)
library(fpc)
library(ggplot2)
source("~/Documents/contests_model/R/load_datasets.R")
source("~/Documents/contests_model/R/make_plots.R")

```
```{R}
all_contests <- load_contests("~/Documents/scrap/contest_data/contest_logging/")
winners <- load_population("~/Documents/scrap/contest_data/contest_logging/")

contests_summary <- ddply(all_contests, .(patch_area), summarise,
                          number_of_fights = length(fight_cost),
                          mean_fight_cost = mean(fight_cost),
                          def_win_rate = mean(defence_winner),
                          atk_death_rate = mean( atk_end_energy  <= 0),
                          def_death_rate = mean( def_end_energy  <= 0))

qplot(data = contests_summary, x = patch_area, y = mean_fight_cost, main = "mean fight cost")
qplot(data = contests_summary, x = patch_area, y = def_win_rate, main = "defender win rate")
qplot(data = contests_summary, x = patch_area, y = atk_death_rate, main = "atacker death rate")
qplot(data = contests_summary, x = patch_area, y = def_death_rate, main = "defender death rate")

```
```{R}
# lots of deaths?
qplot(data = contests_summary, x = patch_area, y = def_death_rate + atk_death_rate)
contests_summary$density <- 500 / conteasts_summary$patch_area
ggplot(contests_summary, aes(x = density, y = number_of_fights)) + geom_point() + scale_y_log10() + scale_x_log10() + geom_smooth(method = lm)

summary(lm(contests_summary$number_of_fights~log10(contests_summary$density)))

ggplot(all_contests[all_contests$patch_area %in% seq(1000, 30000, by = 1000),],
       aes(x = factor(patch_area), y = fight_cost)) +
  geom_violin()  
# the bars are the max - min in that cluster ymin = mediod - (max - min)/2
# this allows for negative values where there arent acually any negative values if the max is an outlier

clustered_plot(dff = all_contests, what = "fight_cost", ylab = "fight_cost")

summary(all_contests$fight_cost)
```

```{R}
# one generation with low patch area / high density

low_patch_area <- all_contests[all_contests$patch_area == 2200,]
high_patch_area <- all_contests[all_contests$patch_area == 25000,]

qplot(data = low_patch_area, x = def_mass, y = atk_mass, col = defence_winner, main = "low patch area")

qplot(data = high_patch_area, x = def_mass, y = atk_mass, col = defence_winner, main = "low patch area")

```

Here is some other plots 
```{R}
clustered_plot(dff = winners, what = "e_0", ylab = "beta")

clustered_plot(dff = winners, what = "k", ylab = "alpha")

ggplot(winners[winners$patch_area %in% seq(1000, 30000, by = 1000),],
       aes(x = factor(patch_area), y = e_0)) +
  geom_violin() + 
  ggtitle("survivors beta") + 
  ylab("beta")

ggplot(winners[winners$patch_area %in% seq(1000, 30000, by = 500),],
       aes(x = factor(patch_area), y = k)) +
  geom_violin() + 
  ggtitle("survivors alpha") + 
  ylab("alpha")

ggplot(winners[winners$patch_area %in% seq(1000, 30000, by = 500),],
       aes(x = factor(patch_area), y = mass)) +
  geom_violin() + 
  ggtitle("survivors mass") + 
  ylab("mass")

ggplot(winners[winners$patch_area %in% seq(1000, 30000, by = 1000),],
       aes(x = factor(patch_area), y = 2 * speed * radius)) +
  geom_violin() + 
  ggtitle("exploration trait") + 
  ylab("2 * speed * radius")

```

### 2016-03-23 scrap/last_gen_energy.rmd
r and v can be fixed

---
title: "energy_at_max_gen"
author: "John Wilshire"
date: "18 March 2016"
output: html_document
---

```{r}
library(jsonlite)
library(plyr)
library(ggplot2)
source("~/Documents/contests_model/R/load_datasets.R")
source("~/Documents/contests_model/R/make_plots.R")

ts_6k <- load_trait_history("~/Documents/contests_model/data/6k/")

last_gens <- ddply(ts_6k, .variables = .(patch_area),
                   function(df) {
                     max_gen <- max(df$generation)
                   })

```

### 2016-03-23 python/male.py

@@ -14,8 +14,12 @@ class Male(object):
 
         # trait values
         if not dad: # first gen: no breeding
-            self.radius = abs(norm.rvs(params["radius_mean"], params["radius_sd"]))
-            self.speed = abs(norm.rvs(params["speed_mean"], params["speed_sd"]))
+            if params["fix_r_and_v"]:
+                self.radius = params["fix_r"]
+                self.speed = params["fix_v"]
+            else:
+                self.radius = abs(norm.rvs(params["radius_mean"], params["radius_sd"]))
+                self.speed = abs(norm.rvs(params["speed_mean"], params["speed_sd"]))
             # pull the aggression from the normal distribution
             self.k = abs(norm.rvs(params["k_mean"], params["k_sd"]))
             self.e_0 = abs(norm.rvs(params["e_0_mean"], params["e_0_sd"]))
@@ -41,11 +45,14 @@ class Male(object):
                     params["maturation_width"]))
 
             # mutations
-            if uniform.rvs() < mutation_rate:
-                self.radius += norm.rvs(0, mutation_sd)
 
-            if uniform.rvs() < mutation_rate:
-                self.speed += norm.rvs(0, mutation_sd)
+            # are we mutating r and v?
+            if not params["fix_r_and_v"]:
+                if uniform.rvs() < mutation_rate:
+                    self.radius += norm.rvs(0, mutation_sd)
+
+                if uniform.rvs() < mutation_rate:
+                    self.speed += norm.rvs(0, mutation_sd)
 
             if uniform.rvs() < mutation_rate:
                 self.e_0 += norm.rvs(0, mutation_sd)
@@ -67,7 +74,7 @@ class Male(object):
         self.time_last_event = self.maturation_time
 
         self.exploration = 2 * self.radius * self.speed
-        self.exploration_rate = (params["N"] * self.exploration) / params["patch_area"]
+        self.exploration_rate = (params["N"] * self.exploration) / float(params["patch_area"])
 
         self.metabolic_cost_search = params["search_energy_coef"] * self.radius * self.speed
         self.metabolic_cost_occupy = params["search_energy_coef"] * self.radius * self.speed

### 2016-04-04 scrap/contest_data.Rmd
fixed r v data, integrated into remake

@@
 date: "18 March 2016"
-output: html_document
+output: pdf_document
 ---
+
+These plots are from training a population to generation 20, they are to give you guys an example of what we can plot for the contests in a generation.
 
@@
 
-```{R}
+```{R, echo = FALSE}
 ggplot(contests, aes(y = atk_start_energy , x = def_start_energy )) + 
@@
 
-```
-
-It looks like a lot of the contests end in death
-```{R}
 ggplot(contests, aes(y = atk_end_energy , x = def_end_energy )) + 
@@
 
-```
-
-The intensity function is $c^2$
-```{R}
 ggplot(contests, aes(x = def_commit, y = atk_commit)) + 
@@
   xlab("commitment of the defender") + 
-  scale_colour_discrete(name="Defender Wins")
-
-ggplot(contests, aes(x = def_beta, y = atk_beta)) + 
-  geom_point(aes(col=contests$defence_winner)) + 
-  ylab("beta trait of the attacker") + 
-  xlab("beta trait of the defender") + 
   scale_colour_discrete(name="Defender Wins")
@@
 
+ggplot(contests, aes(x = def_mass, y = def_commit)) + 
+  geom_point() + 
+  ggtitle("commitment vs mass")
+
+ggplot(contests, aes(x = def_mass/atk_mass, y = def_commit)) + 
+  geom_point() + 
+  ggtitle("commitment vs mass-ratio")
+
 ```

### 2016-04-04 scrap/contests_accross_patch_areas.rmd
fixed r v data, integrated into remake

@@
 ---
-title: "contests accross patch areas"
+title: "contests accross patch areas FIXED R_V "
 author: "John Wilshire"
@@
 ---
-
-contests_accross_patch_areas.rmd
+FIXED R_V
 
@@
 ```{R}
-all_contests <- load_contests("~/Documents/scrap/contest_data/contest_logging/")
-winners <- load_population("~/Documents/scrap/contest_data/contest_logging/")
+all_contests <- load_contests("~/Documents/scrap/contest_data/fixed_r_v/")
+winners <- load_population("~/Documents/scrap/contest_data/fixed_r_v/")
 
@@
 qplot(data = contests_summary, x = patch_area, y = def_death_rate + atk_death_rate)
-contests_summary$density <- 500 / conteasts_summary$patch_area
-ggplot(contests_summary, aes(x = density, y = number_of_fights)) + geom_point() + scale_y_log10() + scale_x_log10() + geom_smooth(method = lm)
+contests_summary$density <- 500 / contests_summary$patch_area
+ggplot(contests_summary, aes(x = density, y = number_of_fights)) +
+  geom_point() +
+  scale_y_log10() + scale_x_log10() + geom_smooth(method = lm)
 

### 2016-04-04 scrap/continuous_escalation.Rmd
old data cleanup and scrap changes
fixed r v data, integrated into remake

@@
 ---
-title: "A step towards continuous Contests"
+title: "A Step Towards Continuous Contests"
 author: "John"
@@
   geom_line() +
-  ggtitle("Example escalation curve") +
   geom_vline(xintercept = 3) + 
   geom_vline(xintercept = 4) +
-  annotate("text",x = 2.8, y = 10, label = "c_ i") +
-  annotate("text",x = 3.8, y = 20, label = "c_ j")
+  geom_polygon(data = rbind(c(0,0), subset(df, commitment < 3), c(3, 0)), 
+               aes(alpha = 0.3, fill = "")) + 
+  guides(fill = FALSE, alpha = FALSE) + 
+  annotate("text", x = 2.8, y = 10, label = "c_ i") +
+  annotate("text", x = 3.8, y = 20, label = "c_ j")
+
 ```

### 2016-04-04 scrap/evolution.Rmd
fixed r v data, integrated into remake

--- 2016-02-18 scrap/evolution.Rmd
@@
 output:
-  pdf_document:
-    fig_crop: no
-    fig_height: 4
+  pdf_document: default
   html_document: default
@@
 library(ggplot2)
-source("../R/load_datasets.R")
-source("../R/make_plots.R")
+source("~/Documents/contests_model/R/load_datasets.R")
+source("~/Documents/contests_model/R/make_plots.R")
 ```
@@
 
-```{R}
-expf10_pop <- load_population("../data/exp_f10/")
-expf15_pop <- load_population("../data/exp_f15/")
-linf10_pop <- load_population("../data/lin_f10/")
-population_plots <- function (what){
-  print(pop_values(expf10_pop, what, paste("expf10_pop", what)))
-  print(pop_values(expf15_pop, what, paste("expf15_pop", what)))
-  print(pop_values(linf10_pop, what, paste("linf10_pop", what)))
+```{R,, echo=FALSE}
+
+exp_100 <- load_population("~/Documents/contests_model/data/cm_100/")
+# linf10_pop <- load_population("../data/lin_f10/")
+# exp10_1000_pop <- load_population("../data/exp_f10_1000/")
+population_plots <- function (y, x, title=""){
+    print(pop_values(exp_100, x, y, paste(title,"exp_100", y)))
 }
 
-population_plots("e_0")
-population_plots("k")
-population_plots("2 * radius * speed")
+population_plots("e_0", "patch_area")
+population_plots("k", "patch_area")
 
@@
 
-These plots trace how parameters evolve through generations with coloured by patch area...
+These plots are against the log of the nest density ie log(nests / patch_area)
 
-```{r}
-expf10 <- load_trait_history("../data/exp_f10/")
-expf15 <- load_trait_history("../data/exp_f15/")
-linf10 <- load_trait_history("../data/lin_f10/")
+```{R, , echo=FALSE}
+population_plots("e_0", "log(N / patch_area)","density -> ")
+population_plots("k", "log(N / patch_area)", "density -> ")
 
-evolution_plots <- function (what){
-  print(trait_evolution_plot(expf10, what, paste("expf10", what)))
-  print(trait_evolution_plot(expf15, what, paste("expf15", what)))
-  print(trait_evolution_plot(linf10, what, paste("linf10", what)))
-}
+```
 
-evolution_plots("mean_k")
-evolution_plots("std_k")
-evolution_plots("std_e_0")
-evolution_plots("num_winners")
-evolution_plots("mean_maturation")
-evolution_plots("contest_energy")
-evolution_plots("search_energy")
-evolution_plots("occupying_energy")
-evolution_plots("mean_exploration")
-evolution_plots("std_exploration")
+now the energy plot
+
+```{R, echo=FALSE}
+exp_trait_100 <- load_trait_history("~/Documents/contests_model/data/cm_100/")
+last <- get_generation(exp_trait_100,499)
+energy_pc_plot(last, "patch_area")
+
 ```
+
+This is the with the nest density increasing $\rightarrow$
+
+```{R,, echo=FALSE}
+
+energy_pc_plot(last, "log( 500 / patch_area )")
+
+```
+
+```{r, echo=FALSE}
+# exp_100_trait <- load_trait_history("~/Documents/contests_model/data/cm_100/")
+# 
+# evolution_plots <- function (what){
+#   print(trait_evolution_plot(exp_100_trait, what, paste("exp_100", what)))
+# }
+# 
+# evolution_plots("mean_k")
+# evolution_plots("std_k")
+# evolution_plots("std_e_0")
+# evolution_plots("num_winners")
+# evolution_plots("mean_maturation")
+# evolution_plots("contest_energy")
+# evolution_plots("search_energy")
+# evolution_plots("occupying_energy")
+# evolution_plots("mean_exploration")
+# evolution_plots("std_exploration")
+```

### 2016-04-04 scrap/generation_comparison.Rmd
fixed r v data, integrated into remake

@@
 date: "19 February 2016"
-output: pdf_document
+output: html_document
 ---
@@
 cm_100 <- load_population("~/Documents/contests_model/data/cm_100/")
-cm_2k <- load_population("~/Documents/contests_model/data/cm_2k/")
+cm_2k <- load_population("~/Documents/scrap/contest_data/6k/")
+low_logit <- load_population("~/Documents/contests_model/data/low_logit/")
+low_logit_100 <- load_population("~/Documents/contests_model/data/low_logit_100/")
+repeated <- load_population("~/Documents/scrap/contest_data/repeated/")
 
@@
   print(
-    pop_values(cm_2k, x = x, y = y, paste("cm_2k",x, y)) + 
+    pop_values(cm_2k, x = x, y = y, paste("cm_6k",x, y)) + 
+      scale_y_continuous(limits = limits)
+  )
+    print(
+    pop_values(low_logit_100, x = x, y = y, paste("low_logit_100",x, y)) + 
+      scale_y_continuous(limits = limits)
+  )
+      print(
+    pop_values(low_logit, x = x, y = y, paste("low_logit",x, y)) + 
       scale_y_continuous(limits = limits)
@@
 }
-e_0_limits <- get_limits("e_0", rbind(cm_2k, cm_100))
-k_limits <- get_limits("k", rbind(cm_2k, cm_100))
+e_0_limits <- get_limits("e_0", rbind(cm_2k, cm_100, low_logit_100, low_logit))
+k_limits <- get_limits("k", rbind(cm_2k, cm_100, low_logit_100, low_logit))
 ```
 
-There seems to be coexistence of strategies after a certain density is reached.
-Where to next?
+The low_logit graphs are those that have a the coef in the logistic of swapping = 1 compared to 10 in the cm ones,
 
-Test different parameters maybe?
-I am running it with lower a lower logit coeff for the winner swap...
+some interesting co-existance,
+but the same scattering problem...
 
-How to limit the aggression drift when training on a lot of generations?
-Different commitment functions?
-Different escalation curves?
-
-```{r}
+```{r, dev='png'}
 population_plots("log(N / patch_area)", "e_0", e_0_limits)
 population_plots("log(N / patch_area)", "k", k_limits)
+
 population_plots("patch_area", "e_0", e_0_limits)

### 2016-04-04 scrap/stabilisation.Rmd
fixed r v data, integrated into remake

---
title: "Stabilisation of Traits"
author: "John Wilshire"
date: "29 February 2016"
output: pdf_document
fig.show: hold
---

```{r, tidy=FALSE}
library(jsonlite)
library(plyr)
library(ggplot2)
source("~/Documents/contests_model/R/load_datasets.R")
source("~/Documents/contests_model/R/make_plots.R")

#traits_cm_100 <- load_trait_history("~/Documents/contests_model/data/cm_100/")
traits_cm_2k <- load_trait_history("~/Documents/contests_model/data/cm_2k/")

stable_what_where <- function (df, what, where){
  print(
    trait_evolution_plot(df[df$patch_area %in% where,], what, "title")

  )
}

stable_what_where(traits_cm_2k, "std_e_0", seq(0, 30000, by = 5000))

stable_what_where(traits_cm_2k, "mean_e_0", seq(0, 30000, by = 5000))

stable_what_where(traits_cm_2k, "std_k", seq(0, 30000, by = 5000))

stable_what_where(traits_cm_2k, "mean_k", seq(0, 30000, by = 5000))
```

### 2016-04-04 R/load_datasets.R

@@ -8,7 +8,7 @@ load_population  <- function (directory  = ""){
   master <- ldply(list.files(directory, full.names = TRUE),                   
                   function (filename) {
                     data.frame(
-                      fromJSON(filename)[1:2]
+                      fromJSON(filename)[c("traits", "parameters")]
                     )})
   colnames(master) <- sub("^[^.]*.","", colnames(master))
   return(master)
@@ -22,8 +22,19 @@ load_trait_history  <- function (directory  = ""){
   master <- ldply(list.files(directory, full.names = TRUE),                   
                   function (filename) {
                     data.frame(
-                      fromJSON(filename)[2:3]
+                      fromJSON(filename)[c("history", "parameters")]
                     )})
+  colnames(master) <- sub("^[^.]*.","", colnames(master))
+  return(master)
+}
+
+load_contests  <- function (directory = ""){
+  if (directory == "") directory <- getwd()
+  master <- ldply(list.files(directory, full.names = TRUE),                   
+                  function (filename) {
+                    data.frame(
+                      fromJSON(filename)[c("contests", "parameters")]
+  )})
   colnames(master) <- sub("^[^.]*.","", colnames(master))
   return(master)
 }

### 2016-04-04 R/make_plots.R

@@ -44,11 +44,19 @@ get_generation <- function(m, gen){
   final$e_per_winner <- final$total_e / final$num_winners
   return(final)
 }
-
-commitment <- seq(0,5, by = 0.01)
-
-escalation_curve <- function (t){
-  t^2
+# gets the max generation accross max generations
+# takes a time serie
+get_max_gens <- function (dd){
+  ddply(dd, .(patch_area), function(df){
+    max_gen <- max(df$generation)
+    mutate(df[df$generation == max_gen, ],
+      total_e = contest_energy + occupying_energy + search_energy,
+      contest_as_pc = contest_energy / total_e,
+      occupying_as_pc = occupying_energy / total_e,
+      search_as_pc = search_energy / total_e,
+      e_per_winner = total_e / num_winners
+    )
+  })
 }
 
 example_escalation <- function (){

### 2016-05-04 ms.tex
renamed data folder in remake.yml

@@
 
-\myfigure{figures/energy_pc_plot_cm_100.png}{pc energy spent on what}{figures/energy_total_plot_cm_100.png}{total energy spent on what}
+%\myfigure{figures/energy_pc_plot_fixed_r_v.png}{pc energy spent on what}{figures/energy_total_plot_fixed_r_v.png}{total energy spent on what, fixed r_v}
 
@@
     \includegraphics[width=15cm,height=10cm]{figures/clustered_e_0_plot.pdf}
-    \caption{A plot of the distribution of the aggression parameter $beta$ over varing $A_{total}$ with k means clustering done, with the k parameter chosen by $pamk$, the number of clusters (coexisting strategies) can be be seen on the right}
+    \caption{A plot of the distribution of the aggression parameter $beta$ over varing $A_{total}$ with k means clustering done, with the k parameter chosen by $pamk$, the number of clusters (coexisting strategies) can be be seen on the right, r - v fixed}
 \end{figure}
